---
title: "COMPASS LE Synoptic Porewater Nutrients 2022–2024"
author: "Fausto Machado-Silva & Roberta B. Peixoto"
date: "`r Sys.Date()`"

# output:
#   html_document:
#     toc: true
#     toc_depth: 2
#     number_sections: true
#     theme: readable
# output_dir: "To Be Reviewed/HTML"

output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    theme: flatly

---

\newpage

## Run Information 
```{r run information, include=TRUE}
#set the run date & user name 
  run_code_date <- "03/06/2026"
  sample_year <- "2023"
  sample_month <- "Feb"
  user <- "Fausto Machado-Silva"
  QAQC_by <- "Fausto Machado-Silva"

#identify the files you want to read in 
  # read in as a list to accommodate multiple runs in a month
  data_files <- c("Raw Data/COMPASS_Box7_MonMon_NUTR_noPO4_385-465_Feb2023_JB.csv", 
                  "Raw Data/Redo_COMPASS_Box7_MonMon_PO4.NOx.SO4_385-465_Feb2023_JB.csv"
                  )

  data_list <- lapply(data_files, read.csv)
  data_file <- dplyr::bind_rows(data_list)

  
# Define the file path for QAQC log file - NO Need to change just check year 
  file_path <- "Raw Data/COMPASS_Box7_MonMon_Nutrients_QAQC_Std_curves_log.csv"
  final_path <- "Processed Data/COMPASS_Box7_MonMon_Nutrients.csv"

#record any notes about the run or anything other info here: 
  run_notes <- "Two run dates: (1st) NOx, NO2, NH4, SO4, (2nd) NOx, SO4 and PO4" #Add notes from QAQC
  
#Set up file path for metadata 
  #downloaded metadata csv - downloaded from Google drive as csv for this year
  Raw_Metadata = "Raw Data/COMPASS_LE_Nutrient_SampleLog.csv"
  
#Set up file path for standard curves 
  #downloaded metadata csv - downloaded from Google drive as csv for this year
  st_curve = "Raw Data/COMPASS_LE_Nutrients_St_curve_Box7_NOx_NO2_NH4_SO4.csv"

  cat(run_notes)
```

##Setup
```{r setup, include=FALSE}

#let you know which section you are in 
cat("Setup")

#a link to the Gitbook or whatever protocol you are using for this analysis 


#load/install packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr,
  purrr,
  ggplot2,
  ggpubr,
  tidyr,
  tidyverse,
  lubridate,
  tinytex,
  stringr,
  writexl,
  readr,
  readxl,
  purrr,
  broom, 
  patchwork)

#Coefficients / constants that are needed for calculations 
N_mw <- 14.0067    # molecular weight of N 
P_mw <- 30.973762  # molecular weight of P 
S_mw <- 32.06      # molecular weight of S
Con1 <- 1000       # conversion factor value
Con2 <- 1000000    # conversion factor value 

#Detection limit and top standards for flagging 
####Need to determine detection limits for these analytes for this SEAL####
NOx_dl <- 0.025  #mgL
NH3_dl <- 0.02  #mgL
PO4_dl <- 0.003 #mgL
NO2_dl <- 0
SO4_dl <- 0

NOx_top <- 1    #ug/L
NH3_top <- 2    #ug/L
PO4_top <- 1    #ug/L
NO2_top <- 40   #ug/L
SO4_top <- 200  #ug/L

#Check standard expected concentrations 
##We set 10% variability as acceptable, with values expected to be either the average between the CCB (blank) and the top standard for each variable, or half of the top standard.
NOx_CCV <- NOx_top/2 
NH3_CCV <- NH3_top/2 
PO4_CCV <- PO4_top/2 
SO4_CCV <- NO2_top/2
NO2_CCV <- SO4_top/2

# #Spike Concentrations 
# NOx_Spk <- 0.5 #ppm or mg/L
# NH3_Spk <- 1
# PO4_Spk <- 0.152 

#expected ranges for sample concentrations used for flags 
r2_cutoff = 0.990
chk_flag = 0.25
chk_conc_flag = 10 #cutoff level for percent difference of check stds vs. expected concentration, 10% for OWC LE Root Exclusion experiment
chks_flag = 60
rep_flag = 25 
#^this is a 25% error between samples
#blank_flag is calculated based on samples later in this code

# #pe check Concentrations
# NH3_pe <- 1.034
# NOx_pe <- 1.51
# PO4_pe <- 0.824

```

##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#read in the raw metadata file 
raw_metadata <- read.csv(Raw_Metadata, check.names = TRUE)

head(raw_metadata)
# 1_1_R34_4192024_ch1_trC_cm

##Format Date column 
raw_metadata$Date_formatted <- paste0(
  raw_metadata$Year,
  sprintf("%02d", raw_metadata$Month),
  sprintf("%02d", raw_metadata$Day)
)

#Create NUTR IDs from what was collected for comparison later
raw_metadata <- raw_metadata %>%
  mutate(NUTR_ID = paste(Box,
                         box_trail, 
                         Vial_number, 
                         Date_formatted, 
                         paste0(Site), 
                         paste0(Zone),
                         paste0(Water),
                         paste0(Sampler),
                         paste0(Repl),
                         paste0(Depth, "cm"),
                        sep = "_"), 
                              NA) 

#Take out the columns and rows that are not relevant
nutr_metadata <- raw_metadata %>%
  select(-"NA") %>%
  # only keep specific columns
  filter(!is.na(NUTR_ID) & NUTR_ID != "")  # remove missing/blank NUTR_ID rows

```

\newpage

##Import Data & Clean
```{r Import Data, include=FALSE}
cat("Import Data")

#set file path for data - in run info chunk 
path <- ("file path")

#Read in Raw Data 
data <- data_file

data$Dilution = (data$Man.Dil.Factor*data$Auto.Dil.Factor)
data$Run_number = 1

data$datetime <- parse_date_time(
  data$Date.and.Time,
  orders = c("mdY HM", "Ymd HMS"),
  tz = "America/New_York"
)

data$Run_Date <- as.Date(data$datetime)


# Rename columns
df_all <- data %>%
  select(Sample_Name = Sample.ID,
         Run_Number = Run_number,
         Conc = Results,
         Absorbance = Absorbance,
         Dilution = Dilution,
         Unit = Units,
         Test = Test,
         Run_Date = Run_Date,
         datetime = datetime) 


df_all <- df_all %>%
  mutate(Test2 = case_when(
    grepl("NOx", Test) ~ "NOx",
    grepl("Ammonia", Test) ~ "NH3",
    grepl("PO4", Test) ~ "PO4",
    grepl("Sulfate", Test) ~ "SO4",
    grepl("Nitrite", Test) ~ "NO2",
    TRUE ~ NA_character_
  ))


#Checking column headers
head(df_all)


```

##Assessing standard Curves
```{r Pull out Standard Curves, echo=FALSE}

#### Standard curves — automatic per Run_Date for all analyses ####

# read in the raw metadata file
stds <- read.csv(st_curve)
stds$Run_Date <- as.Date(stds$Run_Date, format = "%m/%d/%Y")

# Making standards dataframe based on the test ---------------------------------

stds_NOx <- stds %>% filter(Test == "Nitrate (NOx)") 
stds_NH3 <- stds %>% filter(Test == "Ammonia (PPS)")
stds_PO4 <- stds %>% filter(Test == "Phosphate")
stds_NO2 <- stds %>% filter(Test == "Nitrite")
stds_SO4 <- stds %>% filter(Test == "Sulfate")

#For R2
options(digits = 10)


# NOx — LINEAR CURVE PER RUN DATE -----------------------------------------------

NOx_runs <- split(stds_NOx, stds_NOx$Run_Date)

std_curve_NOx <- bind_rows(lapply(NOx_runs, function(df) {

  fit <- lm(Conc ~ Absorbance, data = df)

  data.frame(
    Test = "NOx",
    Run_Date = unique(df$Run_Date),
    Intercept = coef(fit)[1],
    Slope = coef(fit)[2],
    R_squared = summary(fit)$r.squared
  )
}))

NOx_labels <- std_curve_NOx %>%
  mutate(
    R2_display = signif(R_squared, 12),
    label = paste0("R² = ", formatC(R2_display, digits = 12, format = "f"))
  )

# NH3 — LINEAR CURVE PER RUN DATE----------------------------------------------

NH3_runs <- split(stds_NH3, stds_NH3$Run_Date)

std_curve_NH3 <- bind_rows(lapply(NH3_runs, function(df) {

  fit <- lm(Conc ~ Absorbance, data = df)

  data.frame(
    Test = "NH3",
    Run_Date = unique(df$Run_Date),
    Intercept = coef(fit)[1],
    Slope = coef(fit)[2],
    R_squared = summary(fit)$r.squared
  )
}))

NH3_labels <- std_curve_NH3 %>%
  mutate(
    R2_display = signif(R_squared, 12),
    label = paste0("R² = ", formatC(R2_display, digits = 12, format = "f"))
  )


# PO4 — LINEAR CURVE PER RUN DATE -----------------------------------------------

PO4_runs <- split(stds_PO4, stds_PO4$Run_Date)

std_curve_PO4 <- bind_rows(lapply(PO4_runs, function(df) {

  fit <- lm(Conc ~ Absorbance, data = df)

  data.frame(
    Test = "PO4",
    Run_Date = unique(df$Run_Date),
    Intercept = coef(fit)[1],
    Slope = coef(fit)[2],
    R_squared = summary(fit)$r.squared
  )
}))

PO4_labels <- std_curve_PO4 %>%
  mutate(
    R2_display = signif(R_squared, 12),
    label = paste0("R² = ", formatC(R2_display, digits = 12, format = "f"))
  )

# SO4 — 2nd‑ORDER POLYNOMIAL PER RUN DATE----------------------------------------

SO4_runs <- split(stds_SO4, stds_SO4$Run_Date)

std_curve_SO4 <- bind_rows(lapply(SO4_runs, function(df) {

  fit <- lm(
    Conc ~ Absorbance + I(Absorbance^2),
    data = df
  )

  coefs <- coef(fit)

  data.frame(
    Test = "SO4",
    Run_Date = unique(df$Run_Date),
    Intercept = coefs[1],
    Linear = coefs[2],
    Quadratic = coefs[3],
    R_squared = summary(fit)$r.squared
  )
}))

SO4_labels <- std_curve_SO4 %>%
  mutate(
    R2_display = signif(R_squared, 12),
    label = paste0("R² = ", formatC(R2_display, digits = 12, format = "f"))
  )

# NO2 — LINEAR CURVE PER RUN DATE-----------------------------------------------

NO2_runs <- split(stds_NO2, stds_NO2$Run_Date)

std_curve_NO2 <- bind_rows(lapply(NO2_runs, function(df) {

  fit <- lm(Conc ~ Absorbance, data = df)

  data.frame(
    Test = "NO2",
    Run_Date = unique(df$Run_Date),
    Intercept = coef(fit)[1],
    Slope = coef(fit)[2],
    R_squared = summary(fit)$r.squared
  )
}))

NO2_labels <- std_curve_NO2 %>%
  mutate(
    R2_display = signif(R_squared, 12),
    label = paste0("R² = ", formatC(R2_display, digits = 12, format = "f"))
  )

# COMBINE OUTPUTS ---------------------------------------------------------------

stds_combo <- bind_rows(
  stds_NOx,
  stds_NH3,
  stds_PO4,
  stds_SO4,
  stds_NO2
)

std_curve_combo <- bind_rows(
  std_curve_NOx,
  std_curve_NH3,
  std_curve_PO4,
  std_curve_SO4,
  std_curve_NO2
)
```


## Plot standard curves 
```{r Plot Standard Curves, echo=FALSE}

#Plot standard Curve or Curves 
  #v-Nox
NOx_curve_plot <- ggplot(stds_NOx, aes(x = Conc, y = Absorbance)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "coral") +
  facet_wrap(~Run_Date) +
  geom_text(
    data = NOx_labels,
    aes(
      x = Inf,
      y = Inf,
      label = label
    ),
    hjust = 1.05,
    vjust = 1.5,
    inherit.aes = FALSE,
    size = 4
  ) +
  theme_bw() +
  labs(
    x = "Concentration",
    y = "Absorbance",
    title = "NOx Standard Curves"
  )
print(NOx_curve_plot)

  #NH3 
NH3_curve_plot <- ggplot(stds_NH3, aes(x = Conc, y = Absorbance)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "darkgoldenrod") +
  facet_wrap(~Run_Date) +
  geom_text(
    data = NH3_labels,
    aes(
      x = Inf,
      y = Inf,
      label = label
    ),
    hjust = 1.05,
    vjust = 1.5,
    inherit.aes = FALSE,
    size = 4
  ) +
  theme_bw() +
  labs(
    x = "Concentration",
    y = "Absorbance",
    title = "NOx Standard Curves"
  )
print(NH3_curve_plot)


  #PO4 
PO4_curve_plot <- ggplot(stds_PO4, aes(x = Conc, y = Absorbance)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "darkkhaki") +
  facet_wrap(~Run_Date) +
  geom_text(
    data = PO4_labels,
    aes(
      x = Inf,
      y = Inf,
      label = label
    ),
    hjust = 1.05,
    vjust = 1.5,
    inherit.aes = FALSE,
    size = 4
  ) +
  theme_bw() +
  labs(
    x = "Concentration",
    y = "Absorbance",
    title = "NOx Standard Curves"
  )
print(PO4_curve_plot)

  #SO4 
SO4_curve_plot <- ggplot(stds_SO4, aes(x = Conc, y = Absorbance)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2, raw = TRUE), se = FALSE, color = "darkgrey") +
  facet_wrap(~Run_Date) +
  geom_text(
    data = SO4_labels,
    aes(
      x = Inf,
      y = Inf,
      label = label
    ),
    hjust = 1.05,
    vjust = 1.5,
    inherit.aes = FALSE,
    size = 4
  ) +
  theme_bw() +
  labs(
    x = "Concentration",
    y = "Absorbance",
    title = "NOx Standard Curves"
  )
print(SO4_curve_plot)


  #NO2 
NO2_curve_plot <- ggplot(stds_NO2, aes(x = Conc, y = Absorbance)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "darkgreen") +
  facet_wrap(~Run_Date) +
  geom_text(
    data = NO2_labels,
    aes(
      x = Inf,
      y = Inf,
      label = label
    ),
    hjust = 1.05,
    vjust = 1.5,
    inherit.aes = FALSE,
    size = 4
  ) +
  theme_bw() +
  labs(
    x = "Concentration",
    y = "Absorbance",
    title = "NOx Standard Curves"
  )
print(NO2_curve_plot)


############## Report on Cutoffs 

#Report out a flag if the run has an R2 lower than appropriate 
#Write out to the user whether or not the r2 is above the cutoff of 0.98
#For R2
options(digits = 10)

# NOx
NOx_r2_qc <- std_curve_NOx %>%
  mutate(
    decision = ifelse(
      R_squared < r2_cutoff,
      "NOx Curve r2 below cutoff : REASSESS",
      "NOx Curve r2 acceptable : PROCEED"
    )
  )

# NH3
NH3_r2_qc <- std_curve_NH3 %>%
  mutate(
    decision = ifelse(
      R_squared < r2_cutoff,
      "NH3 Curve r2 below cutoff : REASSESS",
      "NH3 Curve r2 acceptable : PROCEED"
    )
  )

# PO4
PO4_r2_qc <- std_curve_PO4 %>%
  mutate(
    decision = ifelse(
      R_squared < r2_cutoff,
      "PO4 Curve r2 below cutoff : REASSESS",
      "PO4 Curve r2 acceptable : PROCEED"
    )
  )

# SO4
SO4_r2_qc <- std_curve_SO4 %>%
  mutate(
    decision = ifelse(
      R_squared < r2_cutoff,
      "SO4 Curve r2 below cutoff : REASSESS",
      "SO4 Curve r2 acceptable : PROCEED"
    )
  )

# NO2
NO2_r2_qc <- std_curve_NO2 %>%
  mutate(
    decision = ifelse(
      R_squared < r2_cutoff,
      "NO2 Curve r2 below cutoff : REASSESS",
      "NO2 Curve r2 acceptable : PROCEED"
    )
  )  
  
# combine
r2_qc_all <- bind_rows(
  NOx_r2_qc,
  NH3_r2_qc,
  PO4_r2_qc,
  SO4_r2_qc,
  NO2_r2_qc
)%>%
  dplyr::select(
    Test, Run_Date,
    Intercept, Slope, Linear, Quadratic,
    R_squared, decision
  )%>%
  mutate(
    Curve_Type = ifelse(Test == "SO4", "Quadratic", "Linear")
  )

# include BOX ID
box_id <- stringr::str_extract(basename(file_path), "Box\\d+")
r2_qc_all <- r2_qc_all %>%
  mutate(Box = box_id)


# add metadata to current run ----------------------------------
stdf_cols <- c(
  "Box",
  "Test",
  "Run_Date",
  "Curve_Type",
  "Intercept",
  "Slope",
  "Linear",
  "Quadratic",
  "R_squared",
  "decision",
  "User",
  "Date_QAQC"        # day of running code (today!)
)


stdf_new <- r2_qc_all %>%
  mutate(
    User = user,
    Date_QAQC = Sys.Date()
  ) %>%
  dplyr::select(all_of(stdf_cols))


# Read existing log OR create it -----------------------------
if (file.exists(file_path)) {

  stdf_log <- read.csv(
    file_path,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  missing_cols <- setdiff(stdf_cols, names(stdf_log))
  stdf_log[missing_cols] <- NA

  stdf_log <- stdf_log %>%
    dplyr::select(all_of(stdf_cols))

} else {

  # correct empty log with proper column types
  stdf_log <- stdf_new[0, ]
}

# missing columns -------------------------------------------
missing_cols <- setdiff(stdf_cols, names(stdf_log))
stdf_log[missing_cols] <- NA

stdf_log <- stdf_log %>%
  dplyr::select(all_of(stdf_cols))


stdf_log <- as.data.frame(
  setNames(
    replicate(length(stdf_cols), logical(0), simplify = FALSE),
    stdf_cols
  )
)

# merging data from std curves to log ----------------------
stdf_log <- stdf_new[0, ]

stdf_log <- stdf_log %>%
  dplyr::anti_join(
    stdf_new,
    by = c("Box", "Test", "Run_Date")
  ) %>%
  dplyr::bind_rows(stdf_new)


# Filter to only rows in Slopes that are NOT already in log (by run_date + analyte) but need to add if none
if (nrow(stdf_log) == 0) {

  # First run: everything is new
  stdf_log <- stdf_new

} else {

  # Future runs: append only genuinely new records
  stdf_log <- stdf_log %>%
    dplyr::anti_join(
      stdf_new,
      by = c("Box", "Test", "Run_Date")
    ) %>%
    dplyr::bind_rows(stdf_new)

}

# Write add_log to CSV
write.csv(stdf_log, file_path, row.names = FALSE)


log <- stdf_log %>%
  mutate(Test = factor(Test, levels = c("NOx", "NH3", "PO4", "SO4", "NO2")))


log_all <- dplyr::filter(log, Test != "SO4")
log_SO4 <- dplyr::filter(log, Test == "SO4")

#Plot the slopes through time 
Slopes_chk1 <- ggplot(log_all, aes(Run_Date, Slope, col=Test)) + 
  geom_point(size=4) + 
  geom_line(aes(group = Test)) + 
  facet_wrap(~Test, scale="free")+
  theme_bw() + 
  labs(title="Slope Drift Assessment", x="Run Date", y="Slope") +
  scale_color_manual(values=c("coral", "darkgoldenrod", "darkkhaki", "darkgreen"))

Slopes_chk2 <- ggplot(log_SO4, aes(Run_Date, Quadratic, col=Test)) + 
  geom_point(size=4) + 
  geom_line(aes(group = Test)) + 
  facet_wrap(~Test, scale="free")+
  theme_bw() + 
  labs(title="Slope Drift Assessment", x="Run Date", y="quadratic: x2") +
  scale_color_manual(values=c("darkgrey"))

Slopes_chk3 <- ggplot(log_SO4, aes(Run_Date, Linear, col=Test)) + 
  geom_point(size=4) + 
  geom_line(aes(group = Test)) + 
  facet_wrap(~Test, scale="free")+
  theme_bw() + 
  labs(title="", x="Run Date", y="linear: x") +
  scale_color_manual(values=c("grey40"))

Slopes_chk4<-Slopes_chk1 + (Slopes_chk2 / Slopes_chk3) +
  plot_layout(widths = c(2.5, 1))


# make printable:
knitr::kable(
  log,
  digits = 12,
  caption = "Standard Curve R² Acceptance by Run Date"
)


#write a flag to the sample dataframe if the r2 is below the cutoff of 0.98

labels_all <- bind_rows(
  NOx_labels %>% select(Test, Run_Date, R_squared),
  NH3_labels %>% select(Test, Run_Date, R_squared),
  PO4_labels %>% select(Test, Run_Date, R_squared),
  SO4_labels %>% select(Test, Run_Date, R_squared),
  NO2_labels %>% select(Test, Run_Date, R_squared)
)

log2 <- log %>%
  select(Test, Run_Date, R_squared)


df_all1 <- df_all %>%
  left_join(log2, 
            by = c("Test2" = "Test", "Run_Date" = "Run_Date"))

#For R2
options(digits = 10)
options(pillar.sigfig = 10)

df_all2 <- df_all1 %>%
  group_by(Run_Date, Test2) %>%
  mutate(flag_value = ifelse(any(R_squared < r2_cutoff),
                             paste0(unique(Test2), " r2 low"),
                             "")) %>%
  ungroup()


df_all <- df_all2 %>%
  mutate(
    NOx_flag = ifelse(Test2 == "NOx", flag_value, ""),
    NH3_flag = ifelse(Test2 == "NH3", flag_value, ""),
    PO4_flag = ifelse(Test2 == "PO4", flag_value, ""),
    SO4_flag = ifelse(Test2 == "SO4", flag_value, ""),
    NO2_flag = ifelse(Test2 == "NO2", flag_value, "")
  )






```
 
 \newpage


##Dilution Corrections - ensure the latest dilution is kept
```{r Dilution Corrections, echo=FALSE}

####It looks like all duplicates were already removed?####

#Calculate the concentration when accounting for the dilution factor if applicable

#checking to see if duplicate samples (mislabeled/wrongly inputted/reran)
df_duplicates <- df_all %>%
  filter(!if_all(everything(), is.na)) %>%  # Remove rows that are entirely NA

  # Exclude common QC sample names and standard patterns
  filter(
    !(Sample_Name %in% c(
      "Duplicate", "CCV", "CCB", "1mg/L ammonia", "DI",
      "Blank", "Auto Spike", "0.15 mg/L"
    )) &
    !startsWith(Sample_Name, "Standard ") &
    !startsWith(Sample_Name, "Nitrite ") &
    !startsWith(Sample_Name, "Nitrate ") &
    !startsWith(Sample_Name, "Sulfate ") &
    !startsWith(Sample_Name, "Abs_Chk_") &
    !startsWith(Sample_Name, "peChk_") &
    !grepl("^\\d{1,2}/\\d{1,2}/\\d{4}\\s+\\d{1,2}:\\d{2}$", Sample_Name) &  # Remove timestamp-like names
    !grepl("^\\d+$", Sample_Name) &  # Remove sample names that are just numbers
    !is.na(Sample_Name) & Sample_Name != ""  # 🔍 Remove blank or missing names
  ) %>%
  group_by(Test, Sample_Name) %>%  # Group by test and sample
  filter(n() > 1) %>%              # Keep only those that appear more than once
  ungroup()

#pulling them out into a string and printing them out
dil_dups_string <- df_duplicates %>%
  distinct(Sample_Name) %>%
  pull(Sample_Name) %>%
  paste(collapse = ", ")

# Make sure Dilution is numeric
df_duplicates$Dilution <- as.numeric(df_duplicates$Dilution)

# Report duplicated samples
if (dil_dups_string != "") {
  message("Duplicated samples: ", dil_dups_string)

  bad_dups <- df_duplicates %>%
    group_by(Test, Sample_Name) %>%
    summarise(
      n_dups = n(),
      any_dilution = any(Dilution > 1, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    filter(!any_dilution)

  if (nrow(bad_dups) > 0) {
    cat("\nWARNING: Duplicated samples with NO dilution (possible input error):\n")
    print(bad_dups)
  } else {
    cat("\n All duplicated samples have valid dilutions. No naming issues detected.\n")
  }

} else {
  message("No duplicated samples.")
}


df_all_cor <- df_all %>%
  arrange(row_number()) %>%
  mutate(
    Dilution = case_when(
      Dilution == "0" ~ 1.0,
      TRUE ~ as.numeric(Dilution)
    ),
    Pair_ID = ifelse(
      grepl("Duplicate", Sample_Name, ignore.case = TRUE),
      paste0(lag(Sample_Name), "_", lag(Run_Number)),  # pair only if same run
      paste0(Sample_Name, "_", Run_Number)
    ),
    Original_Name = Sample_Name
  ) %>%
  group_by(Pair_ID, Test, Run_Number) %>%  # group by Pair_ID, Test, and Run_Number
  filter(
    Dilution == ifelse(any(Dilution > 1, na.rm = TRUE), max(Dilution, na.rm = TRUE), 1)
  ) %>%
  ungroup()


```

 
## Check NOx Reduction Efficiency 
```{r Assess reduction Efficiency, echo=FALSE}

#check on the reduction efficiency for the NOx test and recovery for NO2 
red_eff_NOx <- stds %>% filter(Test == "Nitrate (NOx)", Target > 0.1) #10% of max value 1
rec_eff_NO2 <- stds %>% filter(Test == "Nitrite", Target > 20)        #10% of max value 200

# calculate 
red_eff_NOx<-red_eff_NOx%>%
  group_by(Run_Date)%>%
  mutate(
  Eff_Percent= ifelse(Target > 0, Conc / Target * 100, NA_real_),
  Bias = (Conc - Target) / Target * 100,
  Eff_flag = ifelse(Eff_Percent >= 90 & Eff_Percent <= 105, 'YES', 'NO, rerun'),
  row_num = row_number()
  )

rec_eff_NO2<-rec_eff_NO2%>%
  group_by(Run_Date)%>%
  mutate(
  Eff_Percent= ifelse(Target > 0, Conc / Target * 100, NA_real_),
  Bias = (Conc - Target) / Target * 100,
  Eff_flag = ifelse(Eff_Percent >= 90 & Eff_Percent <= 105, 'YES', 'NO, rerun'),
  row_num = row_number()
  )

#Plot these efficencies 
#NO3
red_eff_NOx_plot <-  ggplot(data = red_eff_NOx, aes(x = ID, y = Eff_Percent, fill=Eff_flag)) +
       geom_bar(stat = 'identity') + 
       facet_wrap(~Run_Date) +
       scale_fill_manual(values = c("YES" = "midnightblue", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="NO3 Reduction Efficiency (%))", title=" ") + 
       theme(legend.position="bottom") +  
       guides(fill=guide_legend(title="Reduction Efficiency >90%"))+
      scale_x_discrete(drop = TRUE)
print(red_eff_NOx_plot)

#NO2
rec_eff_NO2_plot <-  ggplot(data = rec_eff_NO2, aes(x = ID, y = Eff_Percent, fill=Eff_flag)) +
       geom_bar(stat = 'identity') + 
       facet_wrap(~Run_Date) +
       scale_fill_manual(values = c("YES" = "midnightblue", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="NO2 Recovery Efficiency (%))", title=" ") + 
       theme(legend.position="bottom") +  
       guides(fill=guide_legend(title="Reduction Efficiency >90%"))+
      scale_x_discrete(drop = TRUE)
print(rec_eff_NO2_plot)

#Plot these efficencies per concentration
#NO3
red_eff_NOx_plot <- ggplot(data = red_eff_NOx,  aes(x = Target, y = Eff_Percent)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 1) +
  facet_wrap(~Test, scale="free") +
  theme_classic() +
  labs(x = "Target", y = "Reduction Efficiency (%)", title = "Stds. vs Target") +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = "Reduction Efficiency >90%"))
print(red_eff_NOx_plot)

#NO2
rec_eff_NO2_plot <- ggplot(data = rec_eff_NO2,  aes(x = Target, y = Eff_Percent)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 1) +
  facet_wrap(~Test, scale="free") +
  theme_classic() +
  labs(x = "Target", y = "Recovery Efficiency (%)", title = "Stds. vs Target") +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = "Reduction Efficiency >90%"))
print(rec_eff_NO2_plot)

#create a flag 
NOx_mean_red_eff_NOx <- mean(red_eff_NOx$Eff_Percent, na.rm = TRUE)
NO2_mean_rec_eff_NO2 <- mean(rec_eff_NO2$Eff_Percent, na.rm = TRUE)

#report out if flags indicate need for rerun and what the value is
ifelse(NOx_mean_red_eff_NOx >= 95,
       "Mean NOx Reduction Efficiency >95% - PROCEED",
       "Mean NOx Reduction Efficiency <95% - REASSESS")
print(NOx_mean_red_eff_NOx)

ifelse(NO2_mean_rec_eff_NO2 >= 95,
       "Mean NOx Recovery Efficiency >95% - PROCEED",
       "Mean NOx Recovery Efficiency <95% - REASSESS")
print(NO2_mean_rec_eff_NO2)

#write out a flag to the sample dataframe if eff is low
#NO3 or NOx
if (!is.na(NOx_mean_red_eff_NOx) && NOx_mean_red_eff_NOx < 95) {
  df_all$NOx_flag <- "NOx reduction efficiency low"
}

#NO2
if (!is.na(NO2_mean_rec_eff_NO2) && NO2_mean_rec_eff_NO2 < 95) {
  df_all$NO2_flag <- "NO2 recovery efficiency low"
}





```

\newpage

## Analyze the Check Standards
```{r Check Standards, echo=FALSE}

#Pull out check standards from raw file --------------------------------------------
CCV_chks <- df_all %>%
  filter(str_detect(Sample_Name, "CCV"))

#NOx CCVs
NOx_chks <- CCV_chks %>%
  filter(str_detect(Test, "Nitrate")) %>%
  group_by(Run_Date) %>%
  mutate(rep = row_number()) %>%
  ungroup()

# NH3 CCVs
NH3_chks <- CCV_chks %>%
  filter(str_detect(Test, "Ammonia")) %>%
  group_by(Run_Date) %>%
  mutate(rep = row_number()) %>%
  ungroup()

# PO4 CCVs
PO4_chks <- CCV_chks %>%
  filter(str_detect(Test, "Phosphate")) %>%
  group_by(Run_Date) %>%
  mutate(rep = row_number()) %>%
  ungroup()

# SO4 CCVs
SO4_chks <- CCV_chks %>%
  filter(str_detect(Test, "Sulfate")) %>%
  group_by(Run_Date) %>%
  mutate(rep = row_number()) %>%
  ungroup()

# NO2 CCVs
NO2_chks <- CCV_chks %>%
  filter(str_detect(Test, "Nitrite")) %>%
  group_by(Run_Date) %>%
  mutate(rep = row_number()) %>%
  ungroup()


#RSV of standards ------------------------------------------------
#NOx
chks_NOx_rsd <- NOx_chks %>%
  group_by(Run_Date) %>%
  summarise(
    RSD = sd(Conc) / mean(Conc) * 100,
    .groups = "drop"
  )

#NH3
chks_NH3_rsd <- NH3_chks %>%
  group_by(Run_Date) %>%
  summarise(
    RSD = sd(Conc) / mean(Conc) * 100,
    .groups = "drop"
  )

#PO4
chks_PO4_rsd <- PO4_chks %>%
  group_by(Run_Date) %>%
  summarise(
    RSD = sd(Conc) / mean(Conc) * 100,
    .groups = "drop"
  )

#SO4
chks_SO4_rsd <- SO4_chks %>%
  group_by(Run_Date) %>%
  summarise(
    RSD = sd(Conc) / mean(Conc) * 100,
    .groups = "drop"
  )

#NO2
chks_NO2_rsd <- NO2_chks %>%
  group_by(Run_Date) %>%
  summarise(
    RSD = sd(Conc) / mean(Conc) * 100,
    .groups = "drop"
  )

#write out to user about whether or not to continue -------------------------

chk_flag_NOx <- 5    # %
chk_flag_NH3 <- 10
chk_flag_PO4 <- 10
chk_flag_SO4 <- 10  # some labs allow 15 with justification
chk_flag_NO2 <- 10


# NO3
chks_NOx_flag <- chks_NOx_rsd %>%
  mutate(
    chk_flag = ifelse(
      RSD >= chk_flag_NOx,
      "NOx CHECK STANDARD RSD TOO HIGH: REASSESS",
      "NOx Check Standard RSD within Range: PROCEED"
    )
  )

# NH3

chks_NH3_flag <- chks_NH3_rsd %>%
  mutate(
    chk_flag = ifelse(
      RSD >= chk_flag_NH3,
      "NH3 CHECK STANDARD RSD TOO HIGH: REASSESS",
      "NH3 Check Standard RSD within Range: PROCEED"
    )
  )


chks_PO4_flag <- chks_PO4_rsd %>%
  mutate(
    chk_flag = ifelse(
      RSD >= chk_flag_PO4,
      "PO4 CHECK STANDARD RSD TOO HIGH: REASSESS",
      "PO4 Check Standard RSD within Range: PROCEED"
    )
  )



chks_SO4_flag <- chks_SO4_rsd %>%
  mutate(
    chk_flag = ifelse(
      RSD >= chk_flag_SO4,
      "SO4 CHECK STANDARD RSD TOO HIGH: REASSESS",
      "SO4 Check Standard RSD within Range: PROCEED"
    )
  )



chks_NO2_flag <- chks_NO2_rsd %>%
  mutate(
    chk_flag = ifelse(
      RSD >= chk_flag_NO2,
      "NO2 CHECK STANDARD RSD TOO HIGH: REASSESS",
      "NO2 Check Standard RSD within Range: PROCEED"
    )
  )


#calculate percent difference between check standards & expected concentration 
#flag if the percent difference is over X% (defined in setup)
#calculate percent difference of check standards 

#NOx
NOx_chks <- NOx_chks %>%
  group_by(Run_Date) %>%
  mutate(
    NOx_diff = abs(Conc - NOx_CCV) / ((Conc + NOx_CCV) / 2) * 100,
    NOx_diff_flag = ifelse(
      NOx_diff <= chk_conc_flag,
      "YES",
      "NO, rerun"
    )
  ) %>%
  ungroup()


#NH3
NH3_chks <- NH3_chks %>%
  group_by(Run_Date) %>%
  mutate(
    NH3_diff = abs(Conc - NH3_CCV) / ((Conc + NH3_CCV) / 2) * 100,
    NH3_diff_flag = ifelse(
      NH3_diff <= chk_conc_flag,
      "YES",
      "NO, rerun"
    )
  ) %>%
  ungroup()


#PO4
PO4_chks <- PO4_chks %>%
  group_by(Run_Date) %>%
  mutate(
    PO4_diff = abs(Conc - PO4_CCV) / ((Conc + PO4_CCV) / 2) * 100,
    PO4_diff_flag = ifelse(
      PO4_diff <= chk_conc_flag,
      "YES",
      "NO, rerun"
    )
  ) %>%
  ungroup()


#SO4
SO4_chks <- SO4_chks %>%
  group_by(Run_Date) %>%
  mutate(
    SO4_diff = abs(Conc - SO4_CCV) / ((Conc + SO4_CCV) / 2) * 100,
    SO4_diff_flag = ifelse(
      SO4_diff <= chk_conc_flag,
      "YES",
      "NO, rerun"
    )
  ) %>%
  ungroup()


#NO2
NO2_chks <- NO2_chks %>%
  group_by(Run_Date) %>%
  mutate(
    NO2_diff = abs(Conc - NO2_CCV) / ((Conc + NO2_CCV) / 2) * 100,
    NO2_diff_flag = ifelse(
      NO2_diff <= chk_conc_flag,
      "YES",
      "NO, rerun"
    )
  ) %>%
  ungroup()


#plot the check standards
NOx_chks_plot <-  ggplot(data = NOx_chks, aes(x = rep, y = Conc, fill=NOx_diff_flag)) +
       geom_bar(stat = 'identity') + 
        scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred")) +
        labs(x= " ", y="NOx (mg/L)", title="Check Stds: NOx") + 
        theme(legend.position="bottom") +  
        geom_hline(yintercept=NOx_CCV,
              linetype="dashed", color = "black", linewidth=1) + 
        guides(fill=guide_legend(title="% Difference <10%")) +
        theme_classic()+
        facet_wrap(~Run_Date)

NH3_chks_plot <-  ggplot(data = NH3_chks, aes(x = rep, y = Conc, fill=NH3_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred")) +
       theme_classic() + 
       labs(x= " ", y="NH3  (mg/L)", title="Check Stds: NH3") + 
       theme(legend.position="bottom") +  
       geom_hline(yintercept=NH3_CCV,
              linetype="dashed",  color = "black", linewidth=1) + 
       guides(fill=guide_legend(title="% Difference <10%"))+
       facet_wrap(~Run_Date)

PO4_chks_plot <-  ggplot(data = PO4_chks, aes(x = rep, y = Conc, fill=PO4_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred")) +
        theme_classic() + labs(x= " ", y="PO4  (mg/L)", title="Check Stds: PO4") + 
        theme(legend.position="bottom") +  
        geom_hline(yintercept=PO4_CCV,
              linetype="dashed",  color = "black", linewidth=1) + 
        guides(fill=guide_legend(title="% Difference <10%"))+
        facet_wrap(~Run_Date)

SO4_chks_plot <-  ggplot(data = SO4_chks, aes(x = rep, y = Conc, fill=SO4_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred")) +
        theme_classic() + labs(x= " ", y="SO4  (mg/L)", title="Check Stds: SO4") + 
        theme(legend.position="bottom") +  
        geom_hline(yintercept=SO4_CCV,
              linetype="dashed",  color = "black", linewidth=1) + 
        guides(fill=guide_legend(title="% Difference <10%"))+
        facet_wrap(~Run_Date)

NO2_chks_plot <-  ggplot(data = NO2_chks, aes(x = rep, y = Conc, fill=NO2_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred")) +
        theme_classic() + labs(x= " ", y="NO2  (mg/L)", title="Check Stds: NO2") + 
        theme(legend.position="bottom") +  
        geom_hline(yintercept=NO2_CCV,
              linetype="dashed",  color = "black", linewidth=1) + 
        guides(fill=guide_legend(title="% Difference <10%"))+
        facet_wrap(~Run_Date)

#Combined Plot
Chk_Std_Plots <- ggarrange(
  NOx_chks_plot, 
  NH3_chks_plot, 
  PO4_chks_plot, 
  SO4_chks_plot,   
  NO2_chks_plot,   
  nrow = 2, 
  ncol = 3,
  common.legend = TRUE,
  legend = "bottom")
print(Chk_Std_Plots)

#calculate the percent of check standards that are within the range based on the flag -------

#NOx
NOx_chks_percent <- NOx_chks %>%
  group_by(Run_Date) %>%
  summarise(
    percent_within = mean(NOx_diff_flag == "YES") * 100,
    decision = ifelse(
      percent_within >= chks_flag,
      ">60% of NOx Check Standards are within range of expected concentration: PROCEED",
      "<60% of NOx Check Standards are within range of expected concentration: REASSESS"
    ),
    .groups = "drop"
  )

#NH3
NH3_chks_percent <- NH3_chks %>%
  group_by(Run_Date) %>%
  summarise(
    percent_within = mean(NH3_diff_flag == "YES") * 100,
    decision = ifelse(
      percent_within >= chks_flag,
      ">60% of NH3 Check Standards are within range of expected concentration: PROCEED",
      "<60% of NH3 Check Standards are within range of expected concentration: REASSESS"
    ),
    .groups = "drop"
  )

#PO4
PO4_chks_percent <- PO4_chks %>%
  group_by(Run_Date) %>%
  summarise(
    percent_within = mean(PO4_diff_flag == "YES") * 100,
    decision = ifelse(
      percent_within >= chks_flag,
      ">60% of PO4 Check Standards are within range of expected concentration: PROCEED",
      "<60% of PO4 Check Standards are within range of expected concentration: REASSESS"
    ),
    .groups = "drop"
  )

#SO4
SO4_chks_percent <- SO4_chks %>%
  group_by(Run_Date) %>%
  summarise(
    percent_within = mean(SO4_diff_flag == "YES") * 100,
    decision = ifelse(
      percent_within >= chks_flag,
      ">60% of SO4 Check Standards are within range of expected concentration: PROCEED",
      "<60% of SO4 Check Standards are within range of expected concentration: REASSESS"
    ),
    .groups = "drop"
  )

#NO2
NO2_chks_percent <- NO2_chks %>%
  group_by(Run_Date) %>%
  summarise(
    percent_within = mean(NO2_diff_flag == "YES") * 100,
    decision = ifelse(
      percent_within >= chks_flag,
      ">60% of NO2 Check Standards are within range of expected concentration: PROCEED",
      "<60% of NO2 Check Standards are within range of expected concentration: REASSESS"
    ),
    .groups = "drop"
  )


#write out a flag to the sample dataframe if less than 60% of the checks are within the expected CV
# NO3
df_all_cor <- df_all_cor %>%
  left_join(
    NOx_chks_percent %>% select(Run_Date, decision),
    by = "Run_Date"
  ) %>%
  mutate(
    NOx_flag = ifelse(
      Test == "Nitrate (NOx)" & decision == "<60% of NOx Check Standards are within range of expected concentration: REASSESS",
      "NOx checks out of range",
      NOx_flag
    )
  ) %>%
  select(-decision)

#NH4
df_all_cor <- df_all_cor %>%
  left_join(
    NH3_chks_percent %>% select(Run_Date, decision),
    by = "Run_Date"
  ) %>%
  mutate(
    NH3_flag = ifelse(
      Test == "Ammonia (PPS)" & decision == "<60% of NH3 Check Standards are within range of expected concentration: REASSESS",
      "NH3 checks out of range",
      NH3_flag
    )
  ) %>%
  select(-decision)

#PO4
df_all_cor <- df_all_cor %>%
  left_join(
    PO4_chks_percent %>% select(Run_Date, decision),
    by = "Run_Date"
  ) %>%
  mutate(
    PO4_flag = ifelse(
      Test == "Phosphate" & decision == "<60% of PO4 Check Standards are within range of expected concentration: REASSESS",
      "PO4 checks out of range",
      PO4_flag
    )
  ) %>%
  select(-decision)

#SO4
df_all_cor <- df_all_cor %>%
  left_join(
    SO4_chks_percent %>% select(Run_Date, decision),
    by = "Run_Date"
  ) %>%
  mutate(
    SO4_flag = ifelse(
      Test == "Sulfate" & decision == "<60% of SO4 Check Standards are within range of expected concentration: REASSESS",
      "SO4 checks out of range",
      SO4_flag
    )
  ) %>%
  select(-decision)

#NO2
df_all_cor <- df_all_cor %>%
  left_join(
    NO2_chks_percent %>% select(Run_Date, decision),
    by = "Run_Date"
  ) %>%
  mutate(
    NO2_flag = ifelse(
      Test == "Nitrite" & decision == "<60% of NO2 Check Standards are within range of expected concentration: REASSESS",
      "NO2 checks out of range",
      NO2_flag
    )
  ) %>%
  select(-decision)



```

\newpage

## Analyze Blanks 
```{r Analyze Blanks, echo=FALSE}

## CCB was used as the blank for all analytes
#Pull out check standards from raw file 
# NO3
NOx_blks <- df_all %>%
  filter(
    Test == "Nitrate (NOx)",
    str_detect(Sample_Name, "CCB")
  ) %>%
  arrange(Run_Date, datetime) %>%
  group_by(Run_Date) %>%
  mutate(rep = row_number()) %>%
  ungroup()

# NH3
NH3_blks <- df_all %>%
  filter(
    Test == "Ammonia (PPS)",
    str_detect(Sample_Name, "CCB")
  ) %>%
  group_by(Run_Date) %>%
  arrange(Run_Date, datetime) %>%
  mutate(rep = row_number()) %>%
  ungroup()

# PO4
PO4_blks <- df_all %>%
  filter(
    str_trim(Test) == "Phosphate",
    str_detect(Sample_Name, "CCB")
  ) %>%
  arrange(Run_Date, datetime) %>%
  group_by(Run_Date) %>%
  mutate(rep = row_number()) %>%
  ungroup()

# SO4
SO4_blks <- df_all %>%
  filter(
    Test == "Sulfate",
    str_detect(Sample_Name, "CCB")
  ) %>%
  arrange(Run_Date, datetime) %>%
  group_by(Run_Date) %>%
  mutate(rep = row_number()) %>%
  ungroup()

# NO2
NO2_blks <- df_all %>%
  filter(
    Test == "Nitrite",
    str_detect(Sample_Name, "CCB")
  ) %>%
  arrange(Run_Date, datetime) %>%
  group_by(Run_Date) %>%
  mutate(rep = row_number()) %>%
  ungroup()


#Pull out samples from df_all to calc quantile
samples <- df_all %>%
  filter(!is.na(Sample_Name)) %>%
  filter(grepl("^[0-9]+$", Sample_Name)) %>%     # only numbers in sample name
  mutate(Sample_Num = as.integer(Sample_Name))

#Calculating the lowest 25% of sample concentrations to compare to the blank concentrations (old)
# NOx_only <- samples %>%
#   filter(Test == "Nitrate (NOx)")
# blk_flag_NOx <- quantile(NOx_only$Conc, prob=c(.25))   #this gives you the lower 25% quartile of the data 
# blk_flag_NOx <- ifelse(blk_flag_NOx < NOx_dl, NOx_dl/2, blk_flag_NOx) #this will determine if the Q1 of samples is below the detection limit, if it is then it will replace the flag with half of the detection limit 
# NOx_blks$NOx_diff_flag <-  ifelse(NOx_blks$Conc <= blk_flag_NOx, 'YES', 'NO, rerun')


# NOx blanks vs run-specific sample Q1
NOx_only <- samples %>%
  filter(Test == "Nitrate (NOx)")

NOx_blk_flag <- NOx_only %>%
  group_by(Run_Date) %>%
  summarise(
    blk_flag = quantile(Conc, probs = 0.25, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    blk_flag = ifelse(blk_flag < NOx_dl, NOx_dl / 2, blk_flag)
  )

NOx_blks <- NOx_blks %>%
  left_join(NOx_blk_flag, by = "Run_Date") %>%
  mutate(
    NOx_diff_flag = ifelse(Conc <= blk_flag, "YES", "NO, rerun")
  )


# NH3 blanks vs run-specific sample Q1
NH3_only <- samples %>%
  filter(Test == "Ammonia (PPS)")

NH3_blk_flag <- NH3_only %>%
  group_by(Run_Date) %>%
  summarise(
    blk_flag = quantile(Conc, probs = 0.25, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    blk_flag = ifelse(blk_flag < NH3_dl, NH3_dl / 2, blk_flag)
  )

NH3_blks <- NH3_blks %>%
  left_join(NH3_blk_flag, by = "Run_Date") %>%
  mutate(
    NH3_diff_flag = ifelse(Conc <= blk_flag, "YES", "NO, rerun")
  )


# PO4 blanks vs run-specific sample Q1
PO4_only <- samples %>%
  filter(str_trim(Test) == "Phosphate")

PO4_blk_flag <- PO4_only %>%
  group_by(Run_Date) %>%
  summarise(
    blk_flag = quantile(Conc, probs = 0.25, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    blk_flag = ifelse(blk_flag < PO4_dl, PO4_dl / 2, blk_flag)
  )

PO4_blks <- PO4_blks %>%
  left_join(PO4_blk_flag, by = "Run_Date") %>%
  mutate(
    PO4_diff_flag = ifelse(Conc <= blk_flag, "YES", "NO, rerun")
  )


# SO4 blanks vs run-specific sample Q1
SO4_only <- samples %>%
  filter(Test == "Sulfate")

SO4_blk_flag <- SO4_only %>%
  group_by(Run_Date) %>%
  summarise(
    blk_flag = quantile(Conc, probs = 0.25, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    blk_flag = ifelse(blk_flag < SO4_dl, SO4_dl / 2, blk_flag)
  )

SO4_blks <- SO4_blks %>%
  left_join(SO4_blk_flag, by = "Run_Date") %>%
  mutate(
    SO4_diff_flag = ifelse(Conc <= blk_flag, "YES", "NO, rerun")
  )


# NO2 blanks vs run-specific sample Q1
NO2_only <- samples %>%
  filter(Test == "Nitrite")

NO2_blk_flag <- NO2_only %>%
  group_by(Run_Date) %>%
  summarise(
    blk_flag = quantile(Conc, probs = 0.25, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    blk_flag = ifelse(blk_flag < NO2_dl, NO2_dl / 2, blk_flag)
  )

NO2_blks <- NO2_blks %>%
  left_join(NO2_blk_flag, by = "Run_Date") %>%
  mutate(
    NO2_diff_flag = ifelse(Conc <= blk_flag, "YES", "NO, rerun")
  )

#OLD:
# #calculate the percent of check standards that are within the range based on the flag 
# NOx_blks_percent <- (sum(NOx_blks$NOx_diff_flag == "YES")/nrow(NOx_blks))*100
# NH3_blks_percent <- (sum(NH3_blks$NH3_diff_flag == "YES")/nrow(NH3_blks))*100
# PO4_blks_percent <- (sum(PO4_blks$PO4_diff_flag == "YES")/nrow(PO4_blks))*100
# SO4_blks_percent <- (sum(SO4_blks$SO4_diff_flag == "YES")/nrow(SO4_blks))*100
# NO2_blks_percent <- (sum(NO2_blks$NO2_diff_flag == "YES")/nrow(NO2_blks))*100
# 
# #report out if flags indicate need for rerun
# ifelse(NOx_blks_percent >= 60,
#        ">60% of NOx Blanks are below the lower 25% quartile of samples or 1/2 detection limit - PROCEED",
#        "<60% of NOx Blanks are lower 25% quartile of samples or 1/2 the detection limit - REASSESS")
# ifelse(NH3_blks_percent >= 60,
#        ">60% of NH3 Blanks are below the lower 25% quartile of samples - PROCEED",
#        "<60% of NH3 Blanks are lower 25% quartile of samples - REASSESS")
# ifelse(PO4_blks_percent >= 60,
#        ">60% of PO4 Blanks  are below the lower 25% quartile of samples- PROCEED",
#        "<60% of PO4 Blanks are lower 25% quartile of samples - REASSESS")
# ifelse(SO4_blks_percent >= 60,
#        ">60% of SO4 Blanks  are below the lower 25% quartile of samples- PROCEED",
#        "<60% of SO4 Blanks are lower 25% quartile of samples - REASSESS")
# ifelse(NO2_blks_percent >= 60,
#        ">60% of NO2 Blanks  are below the lower 25% quartile of samples- PROCEED",
#        "<60% of NO2 Blanks are lower 25% quartile of samples - REASSESS")


# NO3
NOx_blks_percent <- NOx_blks %>%
  group_by(Run_Date) %>%
  summarise(
    percent_within = mean(NOx_diff_flag == "YES") * 100,
    decision = ifelse(
      percent_within >= 60,
      ">60% of NOx Blanks are below the lower 25% quartile of samples or 1/2 detection limit: PROCEED",
      "<60% of NOx Blanks are below the lower 25% quartile of samples or 1/2 detection limit: REASSESS"
    ),
    .groups = "drop"
  )

# NH3
NH3_blks_percent <- NH3_blks %>%
  group_by(Run_Date) %>%
  summarise(
    percent_within = mean(NH3_diff_flag == "YES") * 100,
    decision = ifelse(
      percent_within >= 60,
      ">60% of NH3 Blanks are below the lower 25% quartile of samples: PROCEED",
      "<60% of NH3 Blanks are below the lower 25% quartile of samples: REASSESS"
    ),
    .groups = "drop"
  )

# PO4
PO4_blks_percent <- PO4_blks %>%
  group_by(Run_Date) %>%
  summarise(
    percent_within = mean(PO4_diff_flag == "YES") * 100,
    decision = ifelse(
      percent_within >= 60,
      ">60% of PO4 Blanks are below the lower 25% quartile of samples: PROCEED",
      "<60% of PO4 Blanks are below the lower 25% quartile of samples: REASSESS"
    ),
    .groups = "drop"
  )

# SO4
SO4_blks_percent <- SO4_blks %>%
  group_by(Run_Date) %>%
  summarise(
    percent_within = mean(SO4_diff_flag == "YES") * 100,
    decision = ifelse(
      percent_within >= 60,
      ">60% of SO4 Blanks are below the lower 25% quartile of samples: PROCEED",
      "<60% of SO4 Blanks are below the lower 25% quartile of samples: REASSESS"
    ),
    .groups = "drop"
  )

# NO2
NO2_blks_percent <- NO2_blks %>%
  group_by(Run_Date) %>%
  summarise(
    percent_within = mean(NO2_diff_flag == "YES") * 100,
    decision = ifelse(
      percent_within >= 60,
      ">60% of NO2 Blanks are below the lower 25% quartile of samples: PROCEED",
      "<60% of NO2 Blanks are below the lower 25% quartile of samples: REASSESS"
    ),
    .groups = "drop"
  )


#plotting the blanks compared to the lower 25% of conc (what the flag is)
NOx_blks_plot <-  ggplot(data = NOx_blks, aes(x = rep, y = Conc, fill=NOx_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="NOx (mg/L)", title="NOx Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(aes(yintercept = blk_flag),  linetype = "dashed",  color = "black",  linewidth = 1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))+
       facet_wrap(~Run_Date)

NH3_blks_plot <-  ggplot(data = NH3_blks, aes(x = rep, y = Conc, fill=NH3_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="NH3  (mg/L)", title="NH3 Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(aes(yintercept = blk_flag),  linetype = "dashed",  color = "black",  linewidth = 1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))+
       facet_wrap(~Run_Date)

PO4_blks_plot <-  ggplot(data = PO4_blks, aes(x = rep, y = Conc, fill=PO4_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="PO4 (mg/L)", title="PO4 Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(aes(yintercept = blk_flag),  linetype = "dashed",  color = "black",  linewidth = 1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))+
       facet_wrap(~Run_Date)

SO4_blks_plot <-  ggplot(data = SO4_blks, aes(x = rep, y = Conc, fill=SO4_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="SO4 (mg/L)", title="SO4 Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(aes(yintercept = blk_flag),  linetype = "dashed",  color = "black",  linewidth = 1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))+
       facet_wrap(~Run_Date)

NO2_blks_plot <-  ggplot(data = NO2_blks, aes(x = rep, y = Conc, fill=NO2_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="NO2 (mg/L)", title="NO2 Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(aes(yintercept = blk_flag),  linetype = "dashed",  color = "black",  linewidth = 1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))+
       facet_wrap(~Run_Date)


#Combined Plot
Blks_Plots <- ggarrange(
  NOx_blks_plot, 
  NH3_blks_plot, 
  PO4_blks_plot,
  SO4_blks_plot,
  NO2_blks_plot,
  nrow = 2, 
  ncol = 3, 
  common.legend = TRUE,
  legend = "bottom")
print(Blks_Plots)

# find average of run blanks for flagging samples later # Compute averages ------
# NOx
blk_avg_NOx <- NOx_blks %>%
  group_by(Run_Date) %>%
  summarise(
    blk_avg = mean(Conc, na.rm = TRUE),
    .groups = "drop"
  )%>%
  mutate(Test = "NOx")


# NH3
blk_avg_NH3 <- NH3_blks %>%
  group_by(Run_Date) %>%
  summarise(
    blk_avg = mean(Conc, na.rm = TRUE),
    .groups = "drop"
  )%>%
  mutate(Test = "NH3")


# PO4
blk_avg_PO4 <- PO4_blks %>%
  group_by(Run_Date) %>%
  summarise(
    blk_avg = mean(Conc, na.rm = TRUE),
    .groups = "drop"
  )%>%
  mutate(Test = "PO4")


# SO4
blk_avg_SO4 <- SO4_blks %>%
  group_by(Run_Date) %>%
  summarise(
    blk_avg = mean(Conc, na.rm = TRUE),
    .groups = "drop"
  )%>%
  mutate(Test = "SO4")


# NO2
blk_avg_NO2 <- NO2_blks %>%
  group_by(Run_Date) %>%
  summarise(
    blk_avg = mean(Conc, na.rm = TRUE),
    .groups = "drop"
  )%>%
  mutate(Test = "NO2")


# Create a data frame
blank_avgs <- dplyr::bind_rows(
  blk_avg_NOx,
  blk_avg_NH3,
  blk_avg_PO4,
  blk_avg_SO4,
  blk_avg_NO2
)

# Pretty print
knitr::kable(blank_avgs, caption = "Mean Concentration of Blanks", digits = 4)

# OLD:
#write out a flag to the sample dataframe if more than 60% of the blanks are above the lower 25% quantile of samples
# if (NOx_blks_percent <= chks_flag) {
#     df_all_cor$NOx_flag <- ifelse(
#     df_all_cor$NOx_flag != "",
#     paste0(df_all_cor$NOx_flag, "; NOx blanks out of range"),
#     "NOx blanks out of range"
#   )
# }
# 
# if (NH3_blks_percent <= chks_flag) {  # assuming you have tn_chks_percent similarly
#     df_all_cor$NH3_flag <- ifelse(
#     df_all_cor$NH3_flag != "",
#     paste0(df_all_cor$NH3_flag, "; NH3 blanks out of range"),
#     "NH3 blanks out of range"
#   )
# }
# 
# if (PO4_blks_percent <= chks_flag) {  # assuming you have tn_chks_percent similarly
#     df_all_cor$PO4_flag <- ifelse(
#     df_all_cor$PO4_flag != "",
#     paste0(df_all_cor$PO4_flag, "; PO4 blanks out of range"),
#     "PO4 blanks out of range"
#   )
# }
# 
# if (SO4_blks_percent <= chks_flag) {  # assuming you have tn_chks_percent similarly
#     df_all_cor$SO4_flag <- ifelse(
#     df_all_cor$SO4_flag != "",
#     paste0(df_all_cor$SO4_flag, "; PO4 blanks out of range"),
#     "PO4 blanks out of range"
#   )
# }
# 
# if (NO2_blks_percent <= chks_flag) {  # assuming you have tn_chks_percent similarly
#     df_all_cor$NO2_flag <- ifelse(
#     df_all_cor$NO2_flag != "",
#     paste0(df_all_cor$NO2_flag, "; PO4 blanks out of range"),
#     "PO4 blanks out of range"
#   )
# }

# write out a flag to the sample dataframe if more than 60% of the blanks are above the lower 25% quantile of samples

# NOx
df_all_cor <- df_all_cor %>%
  left_join(
    NOx_blks_percent %>% select(Run_Date, decision),
    by = "Run_Date"
  ) %>%
  mutate(
    NOx_flag = ifelse(
      Test == "Nitrate (NOx)" &
        decision == "<60% of NOx Blanks are below the lower 25% quartile of samples or 1/2 detection limit: REASSESS",
      ifelse(
        NOx_flag != "",
        paste0(NOx_flag, "; NOx blanks out of range"),
        "NOx blanks out of range"
      ),
      NOx_flag
    )
  ) %>%
  select(-decision)

# NH3
df_all_cor <- df_all_cor %>%
  left_join(
    NH3_blks_percent %>% select(Run_Date, decision),
    by = "Run_Date"
  ) %>%
  mutate(
    NH3_flag = ifelse(
      Test == "Ammonia (PPS)" &
        decision == "<60% of NH3 Blanks are below the lower 25% quartile of samples: REASSESS",
      ifelse(
        NH3_flag != "",
        paste0(NH3_flag, "; NH3 blanks out of range"),
        "NH3 blanks out of range"
      ),
      NH3_flag
    )
  ) %>%
  select(-decision)

# PO4
df_all_cor <- df_all_cor %>%
  left_join(
    PO4_blks_percent %>% select(Run_Date, decision),
    by = "Run_Date"
  ) %>%
  mutate(
    PO4_flag = ifelse(
      str_trim(Test) == "Phosphate" &
        decision == "<60% of PO4 Blanks are below the lower 25% quartile of samples: REASSESS",
      ifelse(
        PO4_flag != "",
        paste0(PO4_flag, "; PO4 blanks out of range"),
        "PO4 blanks out of range"
      ),
      PO4_flag
    )
  ) %>%
  select(-decision)

# SO4
df_all_cor <- df_all_cor %>%
  left_join(
    SO4_blks_percent %>% select(Run_Date, decision),
    by = "Run_Date"
  ) %>%
  mutate(
    SO4_flag = ifelse(
      Test == "Sulfate" &
        decision == "<60% of SO4 Blanks are below the lower 25% quartile of samples: REASSESS",
      ifelse(
        SO4_flag != "",
        paste0(SO4_flag, "; SO4 blanks out of range"),
        "SO4 blanks out of range"
      ),
      SO4_flag
    )
  ) %>%
  select(-decision)

# NO2
df_all_cor <- df_all_cor %>%
  left_join(
    NO2_blks_percent %>% select(Run_Date, decision),
    by = "Run_Date"
  ) %>%
  mutate(
    NO2_flag = ifelse(
      Test == "Nitrite" &
        decision == "<60% of NO2 Blanks are below the lower 25% quartile of samples: REASSESS",
      ifelse(
        NO2_flag != "",
        paste0(NO2_flag, "; NO2 blanks out of range"),
        "NO2 blanks out of range"
      ),
      NO2_flag
    )
  ) %>%
  select(-decision)






```

\newpage



```{r Unit conversion to uM, include=FALSE}

#Convert values based on the Test ID
####Need to check this####
##Everything is mg/L except Nitrite which is ug/L
# ((mg/L) /1000) = g/L
# (g/L) / (g/mol)) = mol/L
# (mol/L) * 1,000,000 = umol/L

##Nitrite
# ((ug/L) /100,000) = g/L
# (g/L) / (g/mol)) = mol/L
# (mol/L) * 100,000 = umol/L

samples <- samples %>%
  mutate(
    Conc_uM = case_when(
      Test == "Nitrate (NOx)" ~ (((as.numeric(samples$Conc))/Con1)/N_mw)*Con2,
      Test == "Ammonia (PPS)" ~ (((as.numeric(samples$Conc))/Con1)/N_mw)*Con2,
      Test == "Phosphate " ~ (((as.numeric(samples$Conc))/Con1)/P_mw)*Con2,
      Test == "Sulfate" ~ (((as.numeric(samples$Conc))/Con1)/S_mw)*Con2,
      Test == "Nitrite" ~ (((as.numeric(samples$Conc))/Con2)/N_mw)*Con2,

      TRUE ~ as.numeric(NA)  # fallback in case of unexpected value
    ),
    # Replace negatives with 0 in 'value' and 'value_converted'
    Conc = pmax(Conc, 0),
    Conc_uM = pmax(Conc_uM, 0)
  )

head(samples)
```


##Sample Flagging - Within range of standard curve
```{r Sample Flagging, echo=FALSE, include = FALSE}

####Need to know detection limit for all analytes####

cat("Sample Flagging")

#Flagging data if the concentration is outside the standards range 
samples_flagged <- samples %>%
  mutate(
    Conc_flag = case_when(
      Test == "Nitrate (NOx)" & Conc < NOx_dl ~ "bdl",
      Test == "Nitrate (NOx)" & Conc > NOx_top    ~ "adl",
      Test == "Nitrate (NOx)"               ~ "Within_Range",
      
      Test == "Ammonia (PPS)" & Conc < NH3_dl ~ "bdl",
      Test == "Ammonia (PPS)" & Conc > NH3_top    ~ "adl",
      Test == "Ammonia (PPS)"                ~ "Within_Range",
      
      Test == "Phosphate " & Conc < PO4_dl  ~ "bdl",
      Test == "Phosphate " & Conc > PO4_top   ~ "adl",
      Test == "Phosphate "                ~ "Within_Range",
      
      Test == "Sulfate" & Conc < SO4_dl  ~ "bdl",
      Test == "Sulfate" & Conc > SO4_top   ~ "adl",
      Test == "Sulfate"                ~ "Within_Range",
      
      Test == "Nitrite" & Conc < NO2_dl  ~ "bdl",
      Test == "Nitrite" & Conc > NO2_top   ~ "adl",
      Test == "Nitrite"                ~ "Within_Range",
      
      TRUE ~ NA_character_  # fallback for unexpected values
    )
  )

```

##Pull out sample id information 
```{r Sample Identification, echo=FALSE, include = FALSE}

cat("Sample Processing")

# 1_43_R76_5292024_ch6_trE_cm
head(nutr_metadata)
head(samples_flagged)

# (1) in case names are in seal_file and as consequence in sample_flagged use: ------------------------

# samples_flagged2 <- samples_flagged %>%
#   separate(
#     col = Sample_Name,
#     sep = "_",
#     into = c("Box ", "Vial_number", "Vial_ID", "Date", "Chamber_ID", "Treatment", "Depth_Units"),
#     remove = FALSE)

# samples_flagged$Date <- paste0("0", samples_flagged$Date)
# samples_flagged$Date_formatted <- as.Date(samples_flagged$Date, format = "%m%d%Y")
# samples_flagged$Date_formatted <- format(samples_flagged$Date_formatted, format = "%Y%m%d")
 
# samples_flagged_selected <- samples_flagged %>%
#   rename(
#     SEAL_Sample_Name = Sample_Name,
#     Old_Date = Date, 
#     Date = Date_formatted)  %>%
#   select(SEAL_Sample_Name, 
#          Vial_ID, 
#          Run_Number, 
#          Conc, 
#          Absorbance, 
#          Dilution, 
#          Unit, 
#          Test, 
#          Run_Time, 
#          NOx_flag, 
#          NH3_flag, 
#          PO4_flag, 
#          SO4_flag, 
#          NO2_flag, 
#          Conc_uM, 
#          Conc_flag)



# (2) in case names are NOT in seal_file and as consequence NOT in sample_flagged use: ------------------------

samples_flagged2 <- samples_flagged %>%
   mutate(Vial_number = Sample_Num)%>%
   dplyr::select(
     Vial_number, everything()
   )




```

## Check to see if samples run match metadata & merge info
```{r check sample ids with metadata, echo=FALSE}

cat("Check Sample IDs with Metadata")

#check to see if all samples are present in the metadata 
all_present <- all(samples_flagged2$Vial_number %in% nutr_metadata$Vial_number)

if (all_present) {
  message("All sample IDs are present in metadata.")
} else {
  message("Some sample IDs are missing from metadata.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(samples_flagged_selected$Vial_number, nutr_metadata$Vial_number)
  print(missing_ids)
}

#merge metadata with sample run data 
merged_data <- samples_flagged2 %>%
  left_join(nutr_metadata, by = "Vial_number")


```

\newpage

## Visualize data
```{r Visualize data, echo=FALSE, fig.height=10, fig.width=15}
cat("Visualize Data")

merged_data$Zone <- factor(
  merged_data$Zone,
  levels = c("UP", "TR", "W", "OW", "Stream"),
  ordered = TRUE
)

### Nitrite + Nitrate 

NOx_forplot <- merged_data %>%
  filter(Test == "Nitrate (NOx)")

#group the data for plotting
NOx_forplot <- NOx_forplot %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

viz_NOx_plot <-  ggplot(data = NOx_forplot, aes(x = Vial_ID, y = Conc, fill=Site)) +
       geom_bar(stat = "identity", position = position_dodge2(preserve = "single")) +
       facet_grid(~ Zone, scales="free") +
        theme_classic() + 
        labs(x= " ", y="NOx (mg/L)", title="Porewater NOx") + 
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
        theme_classic(base_size = 30) +
        theme(axis.text.x = element_blank())

viz_NOx_plot


### Ammonia 
NH3_forplot <- merged_data %>%
  filter(Test == "Ammonia (PPS)")

#group the data for plotting
NH3_forplot <- NH3_forplot %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

viz_NH3_plot <-  ggplot(data = NH3_forplot, aes(x = Vial_ID, y = Conc, fill=Site)) +
       geom_bar(stat = "identity", position = position_dodge2(preserve = "single")) +
       facet_grid(~ Zone, scales="free") +
        theme_classic() + 
        labs(x= " ", y="NH3 (mg/L)", title="Porewater NH3") + 
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
        theme_classic(base_size = 30) +
        theme(axis.text.x = element_blank())

viz_NH3_plot


### Phosphate 
PO4_forplot <- merged_data %>%
  filter(Test == "Phosphate ")

#group the data for plotting
PO4_forplot <- PO4_forplot %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

viz_PO4_plot <-  ggplot(data = PO4_forplot, aes(x = Vial_ID, y = Conc, fill=Site)) +
       geom_bar(stat = "identity", position = position_dodge2(preserve = "single")) +
       facet_grid(~ Zone, scales="free") +
        theme_classic() + 
        labs(x= " ", y="PO4 (mg/L)", title="Porewater PO4") + 
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
        theme_classic(base_size = 30) +
        theme(axis.text.x = element_blank())

viz_PO4_plot

### Sulfate 
SO4_forplot <- merged_data %>%
  filter(Test == "Sulfate")

#group the data for plotting
SO4_forplot <- SO4_forplot %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

viz_SO4_plot <-  ggplot(data = SO4_forplot, aes(x = Vial_ID, y = Conc, fill=Site)) +
       geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), 
                #color="black"
                ) +
       facet_grid(~ Zone, scales="free") +
        theme_classic() + 
        labs(x= " ", y="SO4 (mg/L)", title="Porewater SO4") + 
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
        theme_classic(base_size = 30) +
        theme(axis.text.x = element_blank())
viz_SO4_plot

### Nitrite 
NO2_forplot <- merged_data %>%
  filter(Test == "Nitrite")

#group the data for plotting
NO2_forplot <- NO2_forplot %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

viz_NO2_plot <-  ggplot(data = NO2_forplot, aes(x = Vial_ID, y = Conc, fill=Site)) +
       geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), 
                #color="black"
                ) +
       facet_grid(~ Zone, scales="free") +
        theme_classic() + 
        labs(x= " ", y="NO2 (mg/L)", title="Porewater NO2") + 
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
        theme_classic(base_size = 30) +
        theme(axis.text.x = element_blank())

viz_NO2_plot



```


##Export Processed Data
```{r Export Processed Data, include=FALSE}

cat("Export Processed Data")

# create unique labels discriminating test
merged_data2 <- merged_data %>%
  mutate(Vial_Test_ID = paste0(Vial_ID, "_", Test, "_", Dilution, "_", datetime))


#pivot the data set wider to make it wide format
final_data1 <- merged_data2 %>%
  pivot_wider(
    id_cols = c(Vial_Test_ID, Vial_ID),
    names_from = Test,
    values_from = Conc,
    names_glue = "{Test}_Conc"
  )

final_data2 <- merged_data2 %>%
  pivot_wider(
    id_cols = c(Vial_Test_ID, Vial_ID),
    names_from = Test,
    values_from = Conc_uM,
    names_glue = "{Test}_Conc_uM"
  )

final_data3 <- merged_data2 %>%
  pivot_wider(
    id_cols = c(Vial_Test_ID, Vial_ID),
    names_from = Test,
    values_from = Conc_flag,
    names_glue = "{Test}_Conc_flag"
  )

head(merged_data2)

# take out only the columns we want to merge
df_all_clean_cols <- merged_data2 %>%
  select(Vial_Test_ID, Vial_ID, Box, box_trail, Vial_number, Site, Zone, Water, Repl, Depth, Year, Month, Day,
         Analysis, Test, Run_Date, Dilution,
         Conc, Absorbance, Conc_uM,
         Conc_flag,
         NOx_flag, NH3_flag, PO4_flag, SO4_flag, NO2_flag)

# df_all_clean_cols_one_row <- df_all_clean_cols %>%
#   group_by(Vial_ID) %>%
#   slice(1) %>%      # or use summarise() if you want to aggregate
#   ungroup()


#merge these together
data_list <- list(df_all_clean_cols, final_data1, final_data2, final_data3)

final_data4 <- reduce(data_list, full_join, by = c("Vial_Test_ID"))



final_data5 <- final_data4 %>%
  group_by(Vial_ID.x) %>%
  summarise(
    # Keep ALL non-numeric, non-logical metadata (character, factor, ordered factor, Date, POSIXct, etc.)
    across(
      !where(is.numeric) & !where(is.logical),
      ~ .x[which(!is.na(.x))[1]]
    ),

    # Combine numeric columns (max non-NA value)
    across(
      where(is.numeric),
      ~ if (all(is.na(.x))) NA_real_ else max(.x, na.rm = TRUE)
    ),

    # Combine logical columns
    across(
      where(is.logical),
      ~ if (all(is.na(.x))) NA else any(.x, na.rm = TRUE)
    ),

    .groups = "drop"
  )



final_data6<-final_data5%>%
  dplyr::select(Vial_ID.x, 
                Box, box_trail, Vial_number,
                Year, Month, Day,
                Site, Zone, Water, Repl, Depth,
                Analysis, Dilution, Run_Date, 
                Absorbance, 
                Conc,
                Conc_uM,
                Sulfate_Conc, 
                "Nitrate (NOx)_Conc", 
                Nitrite_Conc, 
                "Ammonia (PPS)_Conc", 
                "Phosphate _Conc", 
                Sulfate_Conc_uM,
                "Nitrate (NOx)_Conc_uM", 
                Nitrite_Conc_uM, 
                "Ammonia (PPS)_Conc_uM", 
                "Phosphate _Conc_uM",    
                Conc_flag,
                NOx_flag,
                NH3_flag,
                PO4_flag, 
                SO4_flag, 
                NO2_flag,
                Sulfate_Conc_flag,
                "Nitrate (NOx)_Conc_flag",
                Nitrite_Conc_flag,
                "Ammonia (PPS)_Conc_flag",
                "Phosphate _Conc_flag",
                )%>%
  dplyr::rename(
    Vial_ID = Vial_ID.x
  )

#Add project information 
final_data_labeled <- final_data6 %>% 
  mutate(
    Project = "COMPASS_LE_Synoptic_Sites_Waters"
  ) %>%
  dplyr::arrange(Vial_number)

#Prepare data to be exported 
final_data <- final_data_labeled %>%
    rename(
    
    NOx_Conc_mgL  = `Nitrate (NOx)_Conc`,
    NOx_Conc_uM   = `Nitrate (NOx)_Conc_uM`,
    NOx_Conc_Flag = `Nitrate (NOx)_Conc_flag`,
    NOx_QAQC_Flag = NOx_flag,
    
    NH3_Conc_mgL  = `Ammonia (PPS)_Conc`,
    NH3_Conc_uM   = `Ammonia (PPS)_Conc_uM`,
    NH3_Conc_Flag =`Ammonia (PPS)_Conc_flag`,
    NH3_QAQC_Flag = NH3_flag,
    
    PO4_Conc_mgL  = `Phosphate _Conc`,
    PO4_Conc_uM   = `Phosphate _Conc_uM`,
    PO4_Conc_Flag =`Phosphate _Conc_flag`,
    PO4_QAQC_Flag = PO4_flag,
    
    SO4_Conc_mgL  = `Sulfate_Conc`,
    SO4_Conc_uM   = `Sulfate_Conc_uM`,
    SO4_Conc_Flag =`Sulfate_Conc_flag`,
    SO4_QAQC_Flag = SO4_flag,
    
    NO2_Conc_mgL  = `Nitrite_Conc`,
    NO2_Conc_uM   = `Nitrite_Conc_uM`,
    NO2_Conc_Flag =`Nitrite_Conc_flag`,
    NO2_QAQC_Flag = NO2_flag,
    # add more rename pairs as needed
  ) %>%
  select(Project, Vial_ID, Box, box_trail, Vial_number, Year, Month, Day, Site, Zone, Water, Repl, Depth,  
         NOx_Conc_mgL, NOx_Conc_uM, NOx_Conc_Flag, NOx_QAQC_Flag,
         NH3_Conc_mgL, NH3_Conc_uM, NH3_Conc_Flag, NH3_QAQC_Flag,
         PO4_Conc_mgL, PO4_Conc_uM, PO4_Conc_Flag, PO4_QAQC_Flag,
         SO4_Conc_mgL, SO4_Conc_uM, SO4_Conc_Flag, SO4_QAQC_Flag,
         NO2_Conc_mgL, NO2_Conc_uM, NO2_Conc_Flag, NO2_QAQC_Flag,
         Run_Date)


#Write out data frame 
  write.csv(final_data, final_path)
  

```


#end



