---
title: "COMPASS_SynopticCB_Tree_GHG_Flux_Calcs""
author: "Roberta Peioxto"
date: "2026-04-07"
output: 
  html_document:
    self_contained: true 
---

## Add Required Packages

```{r packages, include = FALSE}
lapply(c( "fluxfinder", "readxl",
  "dplyr", "ggplot2", "ggpubr", "stringr",
  "purrr", "tidyverse", "here", "broom", "tibble",
  "googledrive", "googlesheets4", "data.table", 
  "matrixStats", "gridExtra", "grid", "tidyverse", "knitr"), 
  library, character.only = TRUE)
```

## Pull in Raw Datafiles

```{r}

#list files in raw data folder
L <- list.files(path = "Raw Data/", pattern = "\\.data$", #"txt$", 
                full.names = TRUE)
head(L)

#Read in the data 
alldat1 <- lapply(L, ffi_read_LI7810)

alldat <- data.table::rbindlist(alldat1, fill=TRUE)

#fixing timestamp errors
#unique(alldat$TZ) [1] "US/Eastern" "America/New_York" "EST5EDT"
alldat <- alldat %>% 
  mutate(TIMESTAMP = with_tz(TIMESTAMP, tzone = "US/Eastern"))
```

## Read in metadata file and match to raw data

```{r}
md <- "Raw Data/COMPASS_SynopticLE_TreeFlux_Metadata_2022b.csv"
metadat <- read.csv(md)

#can remove any row that have NA's in the date column 
metadat <- metadat %>%
  filter(!is.na(Date))

head(metadat) 

#need to make a unique plot id that includes the date or an obs number
metadat$Plot_ID <- paste0(metadat$Plot_ID, "-", seq_len(nrow(metadat)))
#metadat$Obs_length <- 90

#creating the obs_length column based on the stop time
metadat<-metadat%>% 
  filter(SN=="TG10-01448"| SN=="TG10-01426")%>%
  mutate(#Obs_length = Msmt_stop
        start = as.POSIXct(Time_start, format = "%Y-%m-%d %H:%M:%S"),
    end   = as.POSIXct(Stop.time,  format = "%Y-%m-%d %H:%M:%S"),
    Obs_length = as.numeric(difftime(end, start, units = "secs"))
  )

head(metadat)
```

## Check start times overlaps

```{r}

# calculating end time and overlaps
metadat <- metadat %>%
  mutate(
    start = parse_date_time(TIMESTAMP, orders = c("m/d/Y H:M:S", "m/d/Y H:M")),
    end   = start + seconds(Msmt_stop)
  ) %>%
  arrange(SN, start) %>%  
  group_by(SN) %>%
  mutate(
    next_start = lead(start),
    next_plot  = lead(Plot),
    overlap    = end > next_start
  ) %>%
  ungroup()

# example
metadat %>%
  select(SN, Plot, start, end, next_plot, next_start, overlap)

#filtering the overlaps until I come back to fix it =)

metadat <-metadat %>% filter(overlap == FALSE)
```

## Match metadata to raw data

```{r}

#match the metadata to the concentration data 
# matching SN

# alldat$metadat_row <- ffi_metadata_match(
#   data_timestamps = alldat$TIMESTAMP,
#   start_dates = metadat$Date,
#   start_times = metadat$Time_start,
#   obs_lengths = metadat$Obs_length) 
# head(alldat)

# testing only the SN I checked the start time
alldat<-alldat%>%filter(SN=="TG10-01448"| SN=="TG10-01426")

alldat  <- alldat %>%
  group_by(SN) %>%
  mutate(
    metadat_row = ffi_metadata_match(
      data_timestamps = TIMESTAMP,
      start_dates     = metadat$Date[metadat$SN == first(SN)],
      start_times     = metadat$Time_start[metadat$SN == first(SN)],
      obs_lengths     = metadat$Obs_length[metadat$SN == first(SN)]
    )
  ) %>%
  ungroup()


head(alldat)

#Set plot, temp, volume, etc. here 
alldat$Plot <- metadat$Plot[alldat$metadat_row]
alldat$Volume <- metadat$Volume[alldat$metadat_row]
alldat$Temp <- metadat$Temp[alldat$metadat_row]
alldat$Site <- metadat$Site[alldat$metadat_row]
metadat$metadat_row <- seq_len(nrow(metadat))
alldat$Dead_band <- metadat$Dead_band[alldat$metadat_row]

#can remove any row that have NA's in the CH4 column 
#alldat <- alldat %>%
 # filter(!is.na(CH4))

```

## Calculate fluxes

```{r}

# NEED TO PULL TEMP from the weather station
#change the units
alldat$CH4_nmol <- ffi_ppb_to_nmol(alldat$CH4, 
                                volume = alldat$Volume, # m3
                                temp = alldat$Temp)    # degrees C
#> Assuming atm = 101325 Pa
#> Using R = 8.31446261815324 m3 Pa K-1 mol-1

alldat$CH4_nmol_m2 <- alldat$CH4_nmol / 0.16

fluxes <- ffi_compute_fluxes(alldat,
                             group_column = "Plot", 
                             time_column = "TIMESTAMP", 
                             gas_column = "CH4_nmol_m2", 
                             dead_band = "Dead_band")
head(fluxes)

#take a look at the fluxes we calculated with R2 
ggplot(fluxes, aes(Plot, lin_flux.estimate, color = lin_r.squared)) +
  geom_point(size=4) + theme_classic() + 
  geom_linerange(aes(ymin = lin_flux.estimate- lin_flux.std.error,
                   ymax = lin_flux.estimate+ lin_flux.std.error)) +
  scale_color_gradientn(colours = c("darkred",  "red", "white","lightblue", "darkblue"),
                        values = c(0, 0.5, 0.6, 0.7, 1)) +
  ylab("CH4 flux (nmol/m2/s)")

#sort fluxes by R2 
fluxes_good <- as.data.frame(subset(fluxes, lin_r.squared >= .90, select = Plot:TIMESTAMP_max))
fluxes_lowr2 <- as.data.frame(subset(fluxes, lin_r.squared < .90 & lin_flux.estimate < 1, select = Plot:TIMESTAMP_max))
fluxes_weird <- as.data.frame(subset(fluxes, lin_r.squared < .90, select = Plot:TIMESTAMP_max))

#find fluxes that represent the first and fourth quartile of the data 
low <- quantile(fluxes$lin_flux.estimate, 0.05) 
high <- quantile(fluxes$lin_flux.estimate, 0.95) 

fluxes_low <- as.data.frame(subset(fluxes, lin_flux.estimate <= low, select = Plot:TIMESTAMP_max))
fluxes_high <- as.data.frame(subset(fluxes, lin_flux.estimate >= high, select = Plot:TIMESTAMP_max))

tails <- rbind(fluxes_low, fluxes_high)
```

## check the fluxes for which I changed the start time due to lags or high values (recorded in notes of metadata)

```{r}
#### If something is in a V shape should we take the second part to make a positive linear curve?

# Subset data and create relative time (seconds since flux start)
r_all <- alldat %>%
  filter(Plot %in% deadband_overrides$Plot) %>%
  group_by(Plot) %>%
  mutate(
    time_sec = as.numeric(
      difftime(TIMESTAMP, min(TIMESTAMP), units = "secs")
    )
  ) %>%
  ungroup()


# Random sample of plots for visualization
plots_sample <- sample(unique(r_all$Plot), 6,replace = FALSE)

r_plot <- r_all %>%
  filter(Plot %in% plots_sample)

# Plot
ggplot(
  r_plot,
  aes(
    x = TIMESTAMP,
    y = co2,
    color = time_sec >= DeadBand_used
  )
) +
  geom_point(size = 1.8) +
  facet_wrap(~ Plot_ID, scales = "free", ncol = 2) +
  scale_color_manual(
    values = c("grey60", "red"),
    labels = c("Discarded (deadband)", "Used after deadband")
  ) +
  theme_classic() +
  ylab(expression(paste(CO[2], " (ppm)"))) +
  xlab("Time")
```

## Spot check a few fluxes with R2 \> 0.9

```{r, echo=FALSE}

### pull a random sample of good fluxes to check and plot 
good_plots <- unique(fluxes_good$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
good_sample <- sample(good_plots, 9, replace=FALSE)

#Loop to run five random plots subset above 
for(P in good_sample){ 
  r1 <- subset(alldat, Plot == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(as.POSIXct(TIMESTAMP), CH4)) + geom_point() + 
  scale_x_datetime(breaks = "15 sec") + 
  theme_classic() +
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

####################

```

## Take a look at the fluxes \<0.9 and a flux estimate below 1

```{r, echo = FALSE}

####Do we just leave these alone because the fluxes are so low?####

### pull out fluxes with R2 < 0.9 and put them in a list 
low_plots <- unique(fluxes_lowr2$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
low_sample <- sample(low_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in low_sample){ 
  r1 <- subset(alldat, Plot == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() +
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

####################
```

## Take a look at the bad fluxes \<0.9 and have a high flux estimate

## See if we want to remove these fluxes or if we want to try and recalculate these ones?

```{r}
### pull out fluxes with R2 < 0.9 and put them in a list 
bad_plots <- unique(fluxes_weird$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
bad_sample <- sample(bad_plots, 2, replace=FALSE)

#Loop to run five random plots subset above 
for(P in bad_sample){ 
  r1 <- subset(alldat, Plot == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth()
  print(p1)
  }

####################
```

## Look at the really high or really low ones to see if there is a timing issue or something

```{r}
### pull out fluxes in the 5% and 95% tails to look at really high and really low 
tail_plots <- unique(tails$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
tail_sample <- sample(tail_plots, 2, replace=FALSE)

#Loop to run five random plots subset above 
for(P in tail_sample){ 
  r1 <- subset(alldat, Plot == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CH4 [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

#would be nice to put on the plot the flux estimate and if it is a high or low or something 
####################
```

## Check all fluxes 

```{r}
##Put all fluxes together
fluxes_all <- as.data.frame(subset(fluxes, select = Plot:TIMESTAMP_max))

all_plots <- unique(fluxes_all$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
all_sample <- sample(all_plots, 5, replace=FALSE)

#Loop to run five random plots subset above
for(P in all_sample){
  r1 <- subset(alldat, Plot== P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() +
  scale_x_datetime(breaks = "15 sec") +
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth(color="red")
  print(p1)
  }

####################
```

## Remove any fluxes that do not fit criteria 

```{r}

# Create a new dataframe by removing rows that were present in fluxes_weird (11 obs)
#allflux_filtered <- allflux[!(allflux$Plot_ID %in% fluxes_weird$Plot_ID), ] # not all of them are bad, so I am removing below the ones we visually confirmed that were bad

#Any others that are not in the weird category that need to be removed: 
  #also removing ones that are crazy off but were not in the fluxes_weird dataframe 

allflux_filtered <- allflux %>%
  filter(!Plot_ID %in% c("2022_09_23_PTR_TR_9_1", #leak 
                         "2022_07_07_PTR_TR_9_1", #leak
                         "2022_07_08_CRC_TR_36_1" #plateau + sharp increase - soil disturbed during measurement?
                      ))


# View the result
head(allflux_filtered)
```

## Do some visualization of QAQC plots to assess fluxes 

```{r}

p1 <- ggplot(fluxes, aes(lin_flux.estimate, rob_flux.estimate, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  labs(title = "A") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Robust Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) +
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p1

p2 <- ggplot(fluxes, aes(lin_r.squared, poly_r.squared, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  xlim(0,1) + ylim(0,1) +
  labs(title = "B") + 
  ylab(expression(paste("R"^2* " Second-Order Polynomial"))) +
  xlab(expression(paste("R"^2* " Linear Model"))) +
  guides(color = guide_legend(title = "Plot_ID")) + 
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p2

p3 <- ggplot(fluxes, aes(lin_flux.estimate, HM81_flux.estimate, color = lin_r.squared )) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  #xlim(0,50) + #ylim(0,50) +
  labs(title = "C") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") HM1981 Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot_ID")) + 
  theme(legend.position="right") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)

p3

#ggarrange(p1, p2, p3, ncol=3, nrow=1, common.legend = TRUE, legend="none")
```

## Read out final dataframe and export for use elsewhere

```{r}
# allflux_filtered_renamed <- allflux_filtered %>%
#     separate(
#     Plot_ID,
#     into = c("Year", "Month_num", "Day", "Site", "Zone", "Rest"),
#     sep = "_",
#     fill = "right",
#     extra = "merge",
#     remove = FALSE
#   ) %>%
#   
#   mutate(
#     Month = month.abb[as.integer(Month_num)],
#         # RepNumber is always the final number
#     RepNumber = as.integer(str_extract(Plot_ID, "[0-9]+$")),
#         # Replicate = numeric collar ID only (removes letters like a, b)
#     Replicate = as.integer(str_extract(Rest, "^[0-9]+"))
#   ) %>%
#       
#   rename(Plot = Plot_ID) %>%
#   
#   select(
#     RepNumber, Month, Site, Zone, Replicate, Plot,
#     everything(),
#     -Year, -Day, -Month_num, -Rest, -base_id
#   )
# 
# 
# #export as a csv to the folder that you want 
# write.csv(allflux_filtered_renamed, file="Processed Data/COMPASS_Synoptic_LE_TreeStem_Fluxes_Final_2022_CH4.csv", row.names = FALSE) 
# #Change file name
```

##  
