---
title: "Synoptic LE: Manual Redox Probes"
author: "Roberta Peixoto & Fausto Machado-Silva"
date: today
format:
  pdf:
    toc: true
    keep-tex: true
    output-dir: "LE-Synoptic-discrete-data/Manual_Redox_Probe_Data"
---
  
# Measurements taken with SWAP Redox Probes & Calomel Ref probe
# Protocol: https://docs.google.com/document/d/
# 1dL2ZPhLFBAZJmVBCfKxup_KOYY96qA5IWfJmQen50pY/edit
# Units = mV 

##Add Required Packages
```{r packages, include = FALSE}

#Packages that are required 
lapply(c(
  "dplyr", "ggplot2", "ggpubr", "stringr",
  "purrr", "tidyverse", "here", "broom", "tibble",
  "googledrive", "googlesheets4", "data.table", 
  "matrixStats", "gridExtra", "grid", "tidyverse", "knitr"), 
  library, character.only = TRUE)

```

##Setup
```{r setup}

###### File Names - PLEASE CHANGE 
#file path and name for raw summary data file 
dat <- read.csv("Raw_Data/COMPASS_Synoptic_LE_Redox_2022_2025_Raw.csv")
#soildat <- read.csv("Raw Data/LE_Soil_pH.csv") 

##Fix Sample Zone naming
dat <- dat %>%
  mutate(Zone = ifelse(Zone == "WE", "WC", Zone))%>%
  rename(Depth=Depth_cm)

##Std error function
std.error <- function (x, na.rm=FALSE){
  sqrt(var(x, na.rm = na.rm) / sum(!is.na(x)))}

```

## Adjust Raw Readings by pH
## Need to adjust for pH: 
# redox = redoxi + [(ph - 5)59 + 224] This is the equation from Megonigal, 
# Patrick, & Faulkner 1993
# They used a calomel electrode and used 224 to adjust to H2, 
# we use a KCl ref probe, so we need to add 202 instead 
```{r adjust for pH, message = FALSE, warning = FALSE}

##Remove unnecessary column in redox data
prelong <- dat %>% 
  select (c(1:11), -c("Campaign","Time", "Day"))%>% 
  mutate(Notes = NA, 
         Time_24hr = NA, 
         Time_Zone = NA, 
         Topography = NA)

#switch redox data to long format 
datlong <-  as.data.frame(melt(setDT(prelong), id.vars = c("Date", "Month", "Year", "Site", "Zone", "Depth", "Time_24hr", "Time_Zone", "Topography", "Notes"), 
                               variable.name = "RP_mV"))

# ##Add in soil pH data
# all_corr <- merge(datlong, soildat, by = c("Site", "Zone"), all = TRUE)
# 
# ##Adjust redox for pH
# all_corr <- all_corr %>% 
#   mutate(value_corr = value + ((pH - 5)*59 + 202))

##Adjust redox for KCl ref probe
datlong <- datlong %>%
  mutate(value = as.numeric(value),
         value_corr = value + 202)

##Average across all 5-6 measurements for each sampling Site, Zone, Depth, Month, Year
red1 <- datlong %>%
  group_by(Date, Month, Year, Site, Zone, Depth, Time_24hr, Time_Zone, Topography, Notes) %>%
  summarise(Date = first(Date), 
            Avg_mV = mean(value_corr, na.rm=TRUE), 
            Std_dv = sd(value_corr, na.rm=TRUE), 
            Std_err = std.error(value_corr, na.rm=TRUE), 
            Notes = first(Notes),
            .groups = "drop")

# head(red1)

```

##Write Out Adjusted & Avg'd Data
```{r write out processed data, include = FALSE}

years <- unique(red1$Year)

walk(years, function(yr) {
  out_dir <- paste0("Processed_Data_", yr)
  dir.create(out_dir, showWarnings = FALSE)
  
  processed_file_name <- file.path(
    out_dir,
    paste0("COMPASS_Synoptic_LE_Redox_", yr, "_processed.csv")
  )
  
  dat_yr <- red1 %>% filter(Year == yr)
  write.csv(dat_yr, processed_file_name, row.names = FALSE)
})

##Clean field notes and creating columns for ESSDive standard 
unique(red1$Notes, na.rm=TRUE)

red2 <- red1 %>%
  mutate(Day = as.integer(sapply(strsplit(Date, "/"), `[`, 2)),
         Region = "LE",
         Project= "COMPASS: Synoptic",
        Notes = case_when(
    str_detect(Notes, "NA dry soil - very hard") ~ "dry soil - very hard", 
    str_detect(Notes, "average temp") ~ NA, 
    str_detect(Notes, "S58-GHG soilcollar") ~ NA, 
    str_detect(Notes, "time on sheet says 11:05") ~ NA, 
    Notes == "" ~ NA, 
    TRUE ~ Notes
  )) %>% 
  mutate(Notes = str_to_sentence(Notes))%>%
  rename(Plot=Zone)%>%
  dplyr::select(-Date)

unique(red2$Notes)

#write out a csv of all the data to the main folder: 
write.csv(red2, "COMPASS_SynopticLE_Redox_AllData.csv", row.names = FALSE)

```

## Vizualize Data
```{r visualize data, echo = FALSE, warning = FALSE, message = FALSE}

#Plot samples to get a first look at redox (sanity check)
data_plotting <- red1

#Order by Zone from Upland to Surface Water
data_plotting$Zone = factor(data_plotting$Zone, levels = c("UP",  "TR", "WC"))
data_plotting <- data_plotting[order(data_plotting$Zone), ]

#group the data for plotting
data_plotting_year <- data_plotting %>%
  group_by(Site, Year) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

#Plot all data and change colors based on zone:
viz_redox_plot <- ggplot(data_plotting_year, aes(x = row_num, y = Avg_mV, fill = Zone)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "single")) +
  facet_grid(Year ~ Site, scales="free_x") +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "WC" = "#419973"
  )) +
  theme_classic() +
  theme(axis.text.x = element_blank(), 
        legend.position = "bottom") +
  labs(x = " ", y = "Redox Potential (mV)", title = "All Redox Data") +
  scale_x_discrete(drop = TRUE)

viz_redox_plot

#Plot averaged data by depth and change colors based on zone:
red <- datlong %>%
  group_by(Site, Zone, Depth, Year) %>%
  summarise(Date    = mean(Date),
            Avg_mV  = mean(value_corr, na.rm = TRUE),
            Std_dv  = sd(value_corr, na.rm = TRUE),
            Std_err = std.error(value_corr, na.rm = TRUE),
            .groups = "drop")

build_redox_depth_plot <- function(yr) {
  d <- red %>% filter(Year == yr)
  
  ggplot(d, aes(x = Depth, y = Avg_mV, col = Zone, fill = Zone)) +
    geom_point(size = 3) +
    geom_line(linewidth = 1.5) +
    geom_errorbar(aes(ymin = Avg_mV - Std_err, ymax = Avg_mV + Std_err),
                  width = 0.3, position = position_dodge(.1)) +
    scale_x_reverse() +
    coord_flip() +
    theme_classic() +
    facet_grid(~ factor(Site, levels = c("CRC", "PTR", "OWC"))) +
    labs(x = "Depth (cm)", y = "Redox Potential (mV)",
         title = paste("Averaged Redox Data —", yr)) +
    scale_colour_manual(values = c(
      "UP" = "#20063B",
      "TR" = "#FFBC42",
      "WC" = "#419973"
    )) +
    theme(panel.spacing     = unit(.5, "lines"),
          panel.border      = element_rect(color = "black", fill = NA, linewidth = 1),
          strip.background  = element_rect(color = "black", linewidth = 1),
          plot.margin       = margin(t = 0.1, r = 0.1, l = 0.1, b = 0.1, "cm"),
          plot.title        = element_text(hjust = 0.0),
          axis.text         = element_text(size = 12),
          axis.title        = element_text(size = 12),
          legend.text       = element_text(size = 12),
          legend.title      = element_blank(),
          legend.position   = "bottom")
}

years <- sort(unique(red$Year))
redox_depth_plots <- purrr::map(years, build_redox_depth_plot)

# print each one so they each render as a separate figure in the Quarto output
walk(redox_depth_plots, print)

```


## Summarized data for Site and Zone 
```{r Summarize and Visualize Data, echo=FALSE, warning=FALSE, message = FALSE, fig.width=22, fig.height=8, dpi = 500}

##Make Date Column with just Year and Month for Sample ID
all_dat_dates <- data_plotting %>%
  mutate(
    YearMonth = if_else(
      Month < 10,
      # Modify the string:
      paste(Year, Month, sep = "-0"), 
      paste(Year, Month, sep = "-"), 
    )
  )

##Summarize SO4 and Cl by Date, Site, and Zone
Redox_avg <- all_dat_dates %>%
  group_by(Month, Year, Site, Zone) %>%
   dplyr::summarize(mean_mV = mean(Avg_mV, na.rm=TRUE), 
            YearMonth = first(YearMonth), 
            Date = first(Date))

# Concatenate with a day (e.g., "01" for the first day)
Redox_avg$Date <- paste(Redox_avg$YearMonth, Redox_avg$Date, sep = "-")

# Convert to Date object
Redox_avg$Date <- as.Date(Redox_avg$Date)

# Define the specific date breaks you want to display
specific_breaks <- as.Date(c("2022-01-01", "2023-01-01", "2024-01-01", "2025-01-01"))

viz_redox_plot_avgs <- ggplot(Redox_avg, aes(x = Date, y = mean_mV, color = Zone, group = Zone)) +
  geom_point(size = 3) + 
  # geom_errorbar(aes(ymin = mean_Cl - sd_Cl, ymax = mean_Cl + sd_Cl), width = 0.2) + 
  geom_line() + 
  facet_grid( ~ Site, scales="free_y") +
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "WC" = "#419973")) +
  theme_classic() +
  labs(x = "Date", y = "Redox Potential (mV)", title = "Samples: Redox\n (No std dev because values were averaged previously,\n so new std dev would be incorrect)") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  # theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
  theme(text = element_text(size = 30))+
  scale_x_date(breaks = specific_breaks, date_labels = "%Y", expand = expansion(mult = c(0.18, 0.05)))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))

viz_redox_plot_avgs

```


## Summarized Data by Depth
```{r Summarize Data by Depth, echo=FALSE, warning=FALSE, message = FALSE, fig.width=22, fig.height=8, dpi = 500}

##Remove SW and calculate average and sd for each depth
Redox_avg <- all_dat_dates %>%
  group_by(Site, Zone, Depth) %>%
   dplyr::summarize(mean_mV = mean(Avg_mV, na.rm=TRUE))

viz_redox_plot_depth <- ggplot(Redox_avg, aes(x = as.numeric(Depth), y = mean_mV, color = Zone, group = Zone)) +
  geom_point(size = 3) + 
  # Zone_geom_errorbar(aes(ymin = mean_Cl - sd_Cl, ymax = mean_Cl + sd_Cl), width = 0.5, size = 1) + 
  geom_line(size = 1.5) + 
  facet_wrap( ~ Site, scales="free", nrow = 1) + 
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "WC" = "#419973")) +
  theme_classic() +
  #labs(x = "Depth (cm)", y = "Redox Potential (mV)", title = "Samples: Redox profiles across transects") +
  labs(x = "Depth (cm)", y = "Redox Potential (mV)", title = "Samples: Redox\n (No std dev because values were averaged previously,\n so new std dev would be incorrect)") +
  theme(text = element_text(size = 20))+
  coord_flip() + 
  scale_x_continuous(trans = "reverse") + 
  labs(color = "Zone") 

viz_redox_plot_depth

```

## Boxplot by depth
```{r Boxplot by depth, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 500}

##Remove SW
data_plotting_depth <- data_plotting %>%
  mutate(Depth_cm = factor(Depth, levels = c("5", "10", "20")))


#Plot data and change colors based on Zone:
viz_redox_depth <- ggplot(data_plotting_depth, aes(x = Depth_cm, y = Avg_mV, fill = Zone)) +
  geom_boxplot() +
  facet_wrap(~ Site, scales = "free", nrow = 1) +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "WC" = "#419973")) +
  theme_classic() +
  labs(x = "Depth (cm)", y = "Redox Potential (mV)", title = "Samples: Redox") +
  theme(text = element_text(size = 30)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  coord_flip()

viz_redox_depth

```

## Boxplot summary
```{r Overall Boxplot, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}


##Summarize SO4 and Cl by Date, Site, and Zone
viz_redox_boxplot <- ggplot(all_dat_dates, aes(x = Zone, y = Avg_mV, fill = Zone, group = Zone)) +
  geom_boxplot() + 
  facet_grid( ~ Site, scales = "free_x") +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "WC" = "#419973")) +
  theme_classic() +
  labs(x = " ", y = "Redox Potential (mV)", title = "Synoptic LE Soil: Redox Potential") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

viz_redox_boxplot

# Save as a PNG
png("COMPASS_SynopticLE_Redox_alldata.png", width = 2000, height = 1200, res = 100) 
print(viz_redox_boxplot)  # For ggplot objects, use print()
dev.off() 

```

##Calculate summary metrics
```{r Additional summary metrics, include=FALSE, dpi = 300}
# It would also be great to add a summary chunk that will calculate value ranges, means, and standard deviations (and errors) for each site, zone, depth for each year; also for each site zone, depth overall; and for each site and zone overall so that we have some easy access to metrics. Was just thinking this might be helpful today when re-collating the DIC data.

colnames(red1)

##Metrics for each site, zone, depth for each year
summary_by_year <- red1 %>%
  group_by(Year, Site, Zone, Depth) %>%
   reframe(
     min_mV = min(Avg_mV, na.rm=TRUE),
     max_mV = max(Avg_mV, na.rm=TRUE), 
     mean_mV = mean(Avg_mV, na.rm=TRUE)) %>%
  mutate(across(where(is.numeric), round, digits = 4)) 

##Metrics for each site, zone, depth for all years
summary_all_years <- red1 %>%
  group_by(Site, Zone, Depth) %>%
   reframe(
     min_mV = min(Avg_mV, na.rm=TRUE),
     max_mV = max(Avg_mV, na.rm=TRUE), 
     mean_mV = mean(Avg_mV, na.rm=TRUE)) %>%
  mutate(across(where(is.numeric), round, digits = 4)) 

##Metrics for each site and zone for all depths and years
summary_all_depths_years <- red1 %>%
  group_by(Site, Zone) %>%
   reframe(
     min_mV = min(Avg_mV, na.rm=TRUE),
     max_mV = max(Avg_mV, na.rm=TRUE), 
     mean_mV = mean(Avg_mV, na.rm=TRUE)) %>%
  mutate(across(where(is.numeric), round, digits = 2)) %>%
  rename(Plot = Zone)

```

##Calculate number of samples 
```{r Number of samples, include=FALSE, dpi = 300}

##Metrics for each year
sample_numbers_years <- red1 %>%
  group_by(Year) %>%
   reframe(
     Number_of_samples = n())

##Metrics for each site, zone, depth for all years
sample_numbers <- red1 %>%
  group_by(Site, Zone, Depth) %>%
   reframe(
     Number_of_samples = n())
     

```

## Summary Tables
```{r Summary table metrics, echo = FALSE, message = FALSE, results = "asis", out.width="100%"}

library(kableExtra)

summary_table <- kable(
  summary_all_depths_years[, c("Site", "Plot", "min_mV", "max_mV", "mean_mV")],
  format = "latex",
  booktabs = TRUE,
  escape = TRUE,
  digits = 3,
  caption = "Summary Statistics",
  col.names = c("Site", "Plot", "Min (mV)", "Max (mV)", "Mean (mV)")
) %>%
  kable_styling(latex_options = c("hold_position", "striped"),
                font_size = 12) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, bold = TRUE, width = "0.8in") %>%
  column_spec(2, width = "1.6in") %>%
  column_spec(3:5, width = "0.9in")

summary_table

```

## Number of Samples
```{r number of samples tables, echo = FALSE, message = FALSE, dpi = 100}

kable(
  sample_numbers_years[, c("Year", "Number_of_samples")],  
  format = "latex",
  booktabs = TRUE,
  caption = paste0("Total Samples: ", nrow(red1))
) %>%
  kable_styling(
    latex_options = c("hold_position", "striped", "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%   # header
  column_spec(1, bold = TRUE)

```

```{r Number of Samples Barplots, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

##Number of Samples Plots
sample_num_barplot <- ggplot(sample_numbers, aes(x = as.character(Depth), y = Number_of_samples, fill = Zone)) +
  geom_col(position = "dodge", color = "black", width = 0.75) + # 'color' adds a border to the bars
  facet_grid(~Site , scales = "free_x") +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "WC" = "#419973")) +
  theme_classic() +
  labs(x = "Depth (cm)", y = "Number of Samples", title = "Synoptic LE Manual Redox Probe Sample Numbers") +
  theme(
    panel.spacing = unit(0.5, "lines"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))
  
sample_num_barplot
```



