---
title: "COMPASS_Synoptic_LE_SoilFluxes_2024_CH4"
author: "Roberta Peixoto "
date: today
format:
  pdf:
    toc: true
    number-sections: true
output-dir: "2024/PDF"

---
## Goals of this code: 
1.	Reads in raw smart chamber
2.	If no metadatafile, write one out & fields will be: 
Label 
the IDs pulled from the label, 
Plot *fixed so that if there are any issues in the label they are good here 
Drop column – false as long as label looks right 
adjust time – would be initially populated by the start time in the OR blank 
adjust observation length – initially populated by the obs length  OR blank 
notes / comment 
3.	Code reads in the meta data and applies it to the data 
a.	i.e. drops fluxes that have the drop column TRUE or FLAG this 
b.	adjusting times as needed 
4.	Now you compute fluxes 
5.	Visualize and look for issues 
6.	If found, report to meta data file then go back to 3 and reanalyze 
7.	Apply algorithmic qaqc and generate flags 


## Set Up & Load Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("remotes") # if necessary
#remotes::install_github("COMPASS-DOE/fluxfinder")

library(fluxfinder)
library(ggplot2)
library(ggpubr)
library(lubridate)
library(dplyr)
library(tidyr)
library(tidyverse)
library(stringr)
library(kableExtra)
library(data.table)
```

## Pull in Smart Chamber Files from the Raw Data folder
```{r, message=FALSE, warning = FALSE, results ='hide', echo=FALSE}
# Caminho para o ZIP
zip_file <- "D:/GitHub/LE-Synoptic-discrete-data/Soil_Fluxes_Synoptic_LE/2024/Raw Data_LE_2024.zip"

# List files inside the zip
zip_contents <- unzip(zip_file, list = TRUE)

# Select SmartChamber files (json, txt, data)
raw_files <- zip_contents$Name[ grepl("\\.(json|txt|data)$", zip_contents$Name, ignore.case = TRUE) &
  !grepl("manifest|test|generic", zip_contents$Name, ignore.case = TRUE)]

# Extract selected files
unzip(zip_file,
  files = raw_files,
  exdir = tmp_dir)

# List extracted files (REAL paths on disk)
file.list <- list.files(
  path = tmp_dir,
  pattern = "\\.(json|txt|data)$",
  full.names = TRUE,
  recursive = TRUE)

# Read ALL SmartChamber files (json, txt, data)
alldat1 <- lapply(file.list, function(f) {
  tryCatch(
    ffi_read_LIsmartchamber(f),
    error = function(e) {
      message("Skipped file: ", basename(f))
      NULL} )})

# Keep only successful reads with data
alldat1 <- alldat1[
  sapply(alldat1, function(x) !is.null(x) && nrow(x) > 0)]

# Combine into a single table
alldat <- data.table::rbindlist(alldat1, fill = TRUE)

alldat_filtered<-alldat%>%filter(!is.na(TIMESTAMP))

# there is a group of CRC UP and TR data with  TIMESTAMP = NA - need to track them ??
# this code is reading only json files but we have txt too - To Do here
```

## Label format
```{r, echo = FALSE}
# Table describing the days we have visited each site - site is already included manually in the label during measurement
# site_lookup <- tibble::tribble(
#   ~Date,         ~Site,
#   "2024-02-02",  "CRC",
#   "2024-04-26",  "CRC",
#   "2024-07-30",  "CRC",
#   "2024-10-29",  "CRC",
#   "2024-02-23",  "OWC",
#   "2024-04-19",  "OWC",
#   "2024-07-31",  "OWC",
#   "2024-10-18",  "OWC",
#   "2024-02-13",  "PTR",
#   "2024-10-22",  "PTR",
#   "2024-08-07",  "PTR",
#   ) %>%
#   mutate(Date = as.Date(Date))

#fwrite(alldat, "Soil_Fluxes_LE_2024_all_sources_raw.csv")


```

## Unfortunatly - we had many problems with labels - we need to include the sites in the Plot_ID
```{r}
site_lookup <- tibble::tribble(
  ~Date,         ~Site,
  "2024-02-02",  "CRC",
  "2024-04-26",  "CRC",
  "2024-07-30",  "CRC",
  "2024-10-29",  "CRC",
  "2024-02-23",  "OWC",
  "2024-04-19",  "OWC",
  "2024-07-31",  "OWC",
  "2024-10-18",  "OWC",
  "2024-02-13",  "PTR",
  "2024-08-07",  "PTR",
  "2024-10-22",  "PTR"
) %>%
  mutate(Date = as.Date(Date))
# alldat  - date formatting
alldat <- alldat %>%
  mutate(Date = as.Date(TIMESTAMP))%>%filter(lubridate::year(Date) == 2024)


# Fixing label code problems 
alldat <- alldat %>%
  mutate(
    label = label %>%
      str_replace_all("-", "_") %>%
      str_replace_all("\\s+", "") %>%
      toupper(),

    label = label %>%
      str_replace_all("_WE_", "_W_") %>%
      str_replace_all("_WET_", "_W_"),

    label = str_replace(label, "^LE_(CRC|OWC|PTR)_", "\\1_"),

    Site = str_extract(label, "^(CRC|OWC|PTR)"),

    Zone = str_extract(label, "(?<=_)(UP|TR|W)(?=_)"),

    Replicate = as.integer(
      str_extract(label, "(?<=_(UP|TR|W)_)[0-9]+")),

    label = str_replace(label, "([0-9])([A-Z])$", "\\1_\\2"))
 

#correcting labels that were recorded without zone
alldat <- alldat %>%
    mutate(Replicate = coalesce(
            Replicate,
            as.integer(str_extract(label, "(?<=_)[0-9]+")))) %>%
    mutate(Zone = if_else(
            is.na(Zone) &
                Site == "CRC" &
                Replicate %in% c(33, 34, 36, 37, 38, 39, 40),"TR",Zone)) %>%
    mutate(label = if_else(
            Site == "CRC" &
                !grepl("_(UP|TR|W)_", label) &
                Replicate %in% c(33, 34, 36, 37, 38, 39, 40),
            str_replace(label, "^CRC_", "CRC_TR_"),
            label )  )

#I have to change labels that were recorded wrong - tracked in the field notes

#correcting the zone that was saved wrongly

# alldat <- alldat %>%
#   mutate( label = case_when(
#       # CRC 02-21-2023 — TR was saved wrongly as W
#       Site == "CRC" & Date == as.Date("2023-02-21") & PlotNum %in% c(36, 37) ~
#         stringr::str_replace(label, "_W_", "_TR_"),
# 
#       Site == "CRC" & Date == as.Date("2023-02-21") & PlotNum %in% c(38, 40) ~
#         stringr::str_replace(label, "_UP_", "_TR_"),
# 
#       # CRC 03-24-2023 — W was saved wrongly as TR
#       Site == "CRC" & Date == as.Date("2023-03-24") & PlotNum == 48 ~
#         stringr::str_replace(label, "_TR_", "_W_"),
# 
#       # CRC 10-06-2023 — TR was saved wrongly as UP
#       Site == "CRC" & Date == as.Date("2023-10-06") & PlotNum %in% c(37, 39, 40) ~
#         stringr::str_replace(label, "_UP_", "_TR_"),
#       
#       # OWC 04-21-2023 — UP was saved wrongly as TR
#       Site == "OWC" & Date == as.Date("2023-04-21") & PlotNum %in% c(58, 60, 61, 63, 64) ~
#         stringr::str_replace(label, "_UP_", "_TR_"),
# 
#       TRUE ~ label
#     )
#   )
#checking missing field days relabel

# alldat %>%
#      filter(is.na(Site)) %>%
#      distinct(Date) %>%
#      arrange(Date)
```

## Create a unique Plot_ID, check if there is a meta-data file, if not, create a meta-data file
```{r, echo=FALSE}

############### only run this first part if you don't already have a meta-data file 
        #because if you do then all the data for each one will have a plot and you won't be able to change start times

#create unique Plot ID for each measurement: 
#detach("package:data.table", unload=TRUE)

# I am changing this from a month_ID to a date_ID as I have more than one day per month
alldat <- alldat %>%
  mutate(
    Date_str = format(as.Date(TIMESTAMP), "%Y_%m_%d"),
    Plot_ID = str_c(Date_str, label, RepNum, sep = "_")  )

#this needs to be PLot ID so that later we ID plot from metadata and not the metadata row 

# Path inside the zip
zip_file <- "Raw Data_LE_2024.zip"
csv_inside_zip <- "Raw Data/COMPASS_SmartChamber_Metadata_2024.csv"

# Make sure the destination folder exists
dir.create("Raw Data", showWarnings = FALSE)

# Extract only the metadata CSV if it doesn't already exist
if (!file.exists(csv_inside_zip)) {
  unzip(
    zip_file,
    files = csv_inside_zip,
    exdir = "."
  )
}

# Define the file path where you want to save the CSV file
file_path <- "Raw Data/COMPASS_SmartChamber_Metadata_2024.csv"

# Check if the CSV file exists
if (file.exists(file_path)) {
  # If it exists, read it back into R
  metadata <- read.csv(file_path)
  print("CSV file exists and has been read into the code.")
} else {
  # If it does not exist, create the CSV file
    #create a metadata file from the Plot IDs 
      # pull timestart timestamps to put in the metadata and obs lengths 
  metadata <- alldat %>%
  group_by(InstrumentSerialNumber, Plot_ID) %>%
  summarise(SN = first(InstrumentSerialNumber), #Roberta => I included this to process more than one licor per time
    Date = format(min(TIMESTAMP), "%Y-%m-%d"),
    Time_start = format(min(TIMESTAMP), "%H:%M:%S"),
    First_Timestamp = min(TIMESTAMP),  # Minimum (earliest) timestamp for each Plot_ID
    Last_Timestamp = max(TIMESTAMP),   # Maximum (latest) timestamp for each Plot_ID
    Total_Time = as.numeric(difftime(max(TIMESTAMP), min(TIMESTAMP), units = "secs")),  # Time difference in seconds
    ) %>%
  ungroup() %>%
  mutate(Total_Time = pmin(Total_Time, 60),
          drop = FALSE,# Add a 'drop' column with 'FALSE' for every row 
          Notes = "") 
  #mutate(notes = " ")  # Add a 'drop' column with 'FALSE' for every row
  #mutate(Deadband = 5)
  #mutate(Plot = Plot_ID) #think need this because we need to make a plot column
  
  # Write df_unique to CSV
  write.csv(metadata, file = file_path, row.names = FALSE)
  
  print("CSV file does not exist. It has been created.")
}
```

## Mark fluxes that were retaken - this is the first issue! Need to get all but last flux to be TRUE for drop - fieldnotes with explanation for flux retaken - usually weird zeros, Steven probe not inserted in the soil, or label mistake  
```{r, echo=FALSE}

# fluxes that were retaken have similar names with a letter after it 
metadata_filtered <- metadata %>%
  mutate(
    # Base plot ID (everything before the last _number or _number_letter)
    base_id = str_replace(Plot_ID, "_[0-9]+(_[a-z])?$", ""),
    
    # Replicate number (e.g., 32 from 1_Jul_CRC_UP_32 or 1_Jul_CRC_UP_32_b)
    replicate = as.numeric(
      str_extract(Plot_ID, "(?<=_)[0-9]+(?=(_[a-z])?$)")
    ),
    
    # Suffix letter (if present); otherwise empty string
    letter = str_extract(base_id, "(?<=_)[A-Za-z]$"),
    letter = if_else(is.na(letter), "", letter),
    
    # Letter order to define order: "" < a < b < c < d ...
    letter_order = if_else(
      letter == "",
      0L,
      match(letter, letters)  ) ) 


#creating a core ID to compare the retaken obs
metadata_filtered <- metadata_filtered %>%
  mutate(
    # ID for plot not considering retake (remove _a, _b, _c, etc.)
    core_id = str_remove(base_id, "_[A-Za-z]$")
  ) %>%
  group_by(core_id) %>%
  mutate(
    drop = letter_order < max(letter_order, na.rm = TRUE)
  ) %>%
  ungroup()
 
```

##Dont need to pull L1 temp data or any other temps like we do for macrophyte because the smartchamber records the temp

## Match up the metadata to the rest of the data   
```{r, echo = FALSE}
#remove any metadata rows that have NAs for the PlotID
metadata_filtered  <- metadata_filtered %>%
  filter(!is.na(Plot_ID)) %>%
  filter(str_detect(Plot_ID, paste(c("OWC", "CRC", "PTR"), collapse = "|"))) 

alldat<-alldat%>% filter(!is.na(label))

#match the altered metadata to the concentration data with fluxfinder ffi_metadata_match 
# this next part was adapted to include SN and process all Licors at the same time, handling overlapping timestamp among >1 licor instruments. 

# matching SN
alldat <- alldat %>%
  rename(SN=InstrumentSerialNumber)%>%
  group_by(SN) %>%
  mutate(
    metadat_row = ffi_metadata_match(
      data_timestamps = TIMESTAMP,
      start_dates     = metadata_filtered$Date[metadata_filtered$SN == first(SN)],
      start_times     = metadata_filtered$Time_start[metadata_filtered$SN == first(SN)],
      obs_lengths     = metadata_filtered$Total_Time[metadata_filtered$SN == first(SN)]
    )
  ) %>%
  ungroup()

alldat <- alldat %>%
  left_join(
    metadata_filtered %>% mutate(metadat_row = row_number()),
    by = c("SN", "metadat_row", "Plot_ID")
  )

#problems with labels - there was a chamber Offset of 75 instead of 5 and chamber volume of 0 from the smart chamber

alldat <- alldat %>%
  mutate(
    # Fix the Offset bug Plot_ID OWC_02-10-2023 W 67: 75 -> 5 and recalculating volume based on different Offsets
    Offset_cm = if_else(Offset == 75, 5, Offset),

    # Fix chamber bug: volume should never be zero
    ChamVolume_fixed_cm3 = if_else(
      ChamVolume == 0,
      4244.1,
      ChamVolume
    ),

    # Recalculate collar volume (cm3)
    CollarVolume_cm3 = Area * Offset_cm,

    # Recalculate total volume (cm3)
    TotalVolume_fixed_cm3 =
      ChamVolume_fixed_cm3 +
      IrgaVolume +
      CollarVolume_cm3,
  )

```

## fixing some Deadbands  Revisit here, it is not working to fix the dead band
```{r, echo=FALSE}
# Table of plots with custom deadband adjustments

time_overrides <- tibble::tibble(
  Plot_ID = c("2024_07_30_CRC_UP_26_b_2",
              "2024_07_31_OWC_UP_52_2",
              "2024_07_31_OWC_TR_63_2",
              "2024_07_31_OWC_W_72_1",
              "2024_07_31_OWC_W_68_2",
              "2024_02_23_OWC_W_66_b_1",
              "2024_07_30_CRC_UP_32_a_1",
              "2024_07_30_CRC_UP_27_1",
              "2024_02_23_OWC_UP_51_B_1",
              "2024_08_07_PTR_W_18_2",
              "2024_08_07_PTR_W_22_1",
              "2024_07_30_CRC_L_CRC_36_2",  # TO DO: CORRECT LABEL
              "2024_07_30_CRC_UP_25_2",
              "2024_07_31_OWC_UP_53_1",
              "2024_02_23_OWC_W_65_1",
              "2024_07_31_OWC_W_72_2",
              "2024_07_31_OWC_UP_51_2",
              "2024_02_02_CRC_TR_39_1",
              "2024_02_02_CRC_TR_39_2",
              "2024_07_30_CRC_L_CRC_38_2",   # TO DO: CORRECT LABEL
              "2024_07_31_OWC_TR_61_2",
              "2024_02_23_OWC_W_72_1",
              "2024_02_23_OWC_TR_58_2",
              "2024_02_23_OWC_TR_60_1",
              "2024_02_23_OWC_TR_58_1",
              "2024_02_02_CRC_W_43_1",
              "2024_02_23_OWC_W_67_2",
              "2024_07_31_OWC_W_68_1"),
  DeadBand_override = c( 15,
                         7,
                         12,
                         15,
                         15,
                         5,
                         15,
                         15,
                         10,
                         5,
                         6,
                         3,
                         15,
                         3,
                         5,
                         15,
                         12,
                         6,
                         4,
                         15,
                         5,
                         6,
                         10,
                         10,
                         8,
                         15,
                         13,
                         15),
  stop_offset_secs = NA_real_ )

#apply new incubation period
alldat_trimmed <- alldat %>%
  left_join(time_overrides, by = "Plot_ID") %>%
  group_by(Plot_ID) %>%
  mutate(
    t_start_original = min(TIMESTAMP),
    t_end_original   = max(TIMESTAMP),
    
    # deadband manual or metadat
    t_start_adjusted = if_else(
      !is.na(DeadBand_override),
      t_start_original + DeadBand_override,
      t_start_original + as.numeric(coalesce(first(DeadBand), 0))
    ),
    
    # new offset 
    t_end_adjusted = if_else(
      !is.na(stop_offset_secs),
      t_end_original + stop_offset_secs,
      t_end_original
    )
  ) %>%
  filter(TIMESTAMP >= t_start_adjusted,   
         TIMESTAMP <= t_end_adjusted) %>% 
  ungroup() %>%
  select(-t_start_original, -t_end_original, 
         -t_start_adjusted, -t_end_adjusted,
         -DeadBand_override,-stop_offset_secs)

```

##Extracting initial CH4 concentration
```{r, echo=FALSE}

## Extract initial concentration (t0) for each Plot for QAQC later
#saving df to use after flux calculation
df_CH4_in <- alldat %>%
  arrange(Plot_ID, TIMESTAMP) %>%
  group_by(Plot_ID) %>%
  summarise(CH4_initial = first(ch4), .groups = "drop")  

```

## Change the units from ppb to nmol  
```{r}
#change the units
alldat_trimmed$CH4_nmol <- ffi_ppb_to_nmol(alldat_trimmed$ch4, 
                                volume = (alldat_trimmed$TotalVolume_fixed_cm3 / 1000000), # have to divide to get to m3
                                temp = alldat_trimmed$chamber_t)    # degrees C
#> Assuming atm = 101325 Pa
#> Using R = 8.31446261815324 m3 Pa K-1 mol-1

alldat_trimmed$CH4_nmol_m2 <- alldat_trimmed$CH4_nmol / 0.0318 #adjust for the area 


```

## Calculate fluxes
```{r, message=FALSE, warning = FALSE, results ='hide', echo=FALSE}

#selecting the base id to include in the flux df later
plot_base_lookup <- alldat_trimmed%>%
  select(Plot_ID, base_id)%>% 
  filter(!is.na(base_id)) %>%   
  distinct(Plot_ID, .keep_all = TRUE)


#calculate fluxes for all other sites 
fluxes <- ffi_compute_fluxes(alldat_trimmed,
                             group_column = "Plot_ID", 
                             time_column = "TIMESTAMP", 
                             gas_column = "CH4_nmol_m2", 
                             dead_band = "DeadBand")
head(fluxes)
#including the unique ID 
#fluxes <- fluxes %>%
#  left_join(plot_base_lookup, by = "Plot_ID")

##Create a dataframe with averaged soil probe data per Plot and add to flux data
soildat <- alldat_trimmed %>% 
    group_by(Plot_ID) %>%
  dplyr::summarize(
     soilp_c = mean(soilp_c, na.rm=TRUE),
     soilp_m = mean(soilp_m, na.rm=TRUE),
     soilp_t = mean(soilp_t, na.rm=TRUE))

allflux <- left_join(fluxes, soildat, by = "Plot_ID")

```


## Visualize fluxes and sort by r2 
```{r, message=FALSE, warning = FALSE, results ='hide', echo = FALSE}

#take a look at the fluxes we calculated with R2 
ggplot(allflux, aes(Plot_ID, lin_flux.estimate, color = lin_r.squared)) +
  geom_point(size=4) + theme_classic() + #ylim(-50,2000) + 
  geom_linerange(aes(ymin = lin_flux.estimate- lin_flux.std.error,
                   ymax = lin_flux.estimate+ lin_flux.std.error)) +
  scale_color_gradientn(colours = c("darkred",  "red", "white","lightblue", "darkblue"),
                        values = c(0, 0.5, 0.6, 0.7, 1)) +
  ylab("CH4 flux (nmol/m2/s)")

#sort fluxes by R2 
fluxes_good <- as.data.frame(subset(allflux, lin_r.squared >= .90, select = Plot_ID:TIMESTAMP_max))
fluxes_lowr2 <- as.data.frame(subset(allflux, lin_r.squared < .90 & lin_flux.estimate < 1, select = Plot_ID:TIMESTAMP_max))
fluxes_weird <- as.data.frame(subset(allflux, lin_r.squared < .90 & lin_flux.estimate > 1, select = Plot_ID:TIMESTAMP_max))

#find fluxes that represent the first and fourth quartile of the data 
low <- quantile(allflux$lin_flux.estimate, 0.05, na.rm=TRUE) 
high <- quantile(allflux$lin_flux.estimate, 0.95, na.rm=TRUE) 

fluxes_low <- as.data.frame(subset(allflux, lin_flux.estimate <= low, select = Plot_ID:TIMESTAMP_max))
fluxes_high <- as.data.frame(subset(allflux, lin_flux.estimate >= high, select = Plot_ID:TIMESTAMP_max))

tails <- rbind(fluxes_low, fluxes_high)

```

## check the fluxes for which I changed the start time due to lags or ebullition (recorded in notes of metadata)
```{r}


# grab trim limits for each overridden Plot
trim_limits <- alldat_trimmed %>%
  filter(Plot_ID %in% time_overrides$Plot_ID) %>%
  group_by(Plot_ID) %>%
  summarise(
    t_start_trim = min(TIMESTAMP),
    t_end_trim   = max(TIMESTAMP),
    .groups = "drop"
  )

# use full alldat + trim limits to classify each point
r_all <- alldat %>%
  filter(Plot_ID %in% time_overrides$Plot_ID) %>%
  left_join(trim_limits, by = "Plot_ID") %>%
  group_by(Plot_ID) %>%
  mutate(
    time_sec = as.numeric(difftime(TIMESTAMP, min(TIMESTAMP), units = "secs")),
    point_status = case_when(
      TIMESTAMP < t_start_trim ~ "Discarded (DeadBand)",
      TIMESTAMP > t_end_trim   ~ "Discarded (stop trim)",
      TRUE                     ~ "Used"
    )
  ) %>%
  ungroup()

# ---- PLOTTING: one plot per Plot_ID (PDF friendly) ----

plot_ids <- unique(r_all$Plot_ID)

for (P in plot_ids) {

  p <- ggplot(
    r_all %>% filter(Plot_ID == P),
    aes(x = TIMESTAMP, y = ch4, color = point_status)
  ) +
    geom_point(size = 1.8) +
    scale_color_manual(
      values = c(
        "Used"                  = "#E41A1C",
        "Discarded (DeadBand)"  = "grey60",
        "Discarded (stop trim)" = "grey20"
      )
    ) +
    theme_classic(base_size = 12) +
    ylab(expression(paste(CH[4], " (ppb)"))) +
    xlab("Time") +
    labs(
      title = paste("Incubation Period Adjustments:", P),
      color = "Point Status"
    ) +
    theme(
      legend.position = "bottom",
      plot.title = element_text(face = "bold")
    )

  print(p)   # IMPORTANT: each print() = new page in PDF
}

```


## Spot check a few fluxes with R2 > 0.9  
```{r, echo=FALSE}

### pull a random sample of good fluxes to check and plot 
good_plots <- unique(fluxes_good$Plot_ID)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
good_sample <- sample(good_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in good_sample){ 
  r1 <- subset(alldat_trimmed, Plot_ID == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(as.POSIXct(TIMESTAMP), ch4)) + geom_point() + 
  scale_x_datetime(breaks = "15 sec") + 
  theme_classic() +
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

####################
```

## Take a look at the  fluxes <0.9 and a flux estimate below 1
```{r, echo=FALSE}

####Do we just leave these alone because the fluxes are so low?####

### pull out fluxes with R2 < 0.9 and put them in a list 
low_plots <- unique(fluxes_lowr2$Plot_ID)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
low_sample <- sample(low_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in low_sample){ 
  r1 <- subset(alldat_trimmed, Plot_ID == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, ch4)) + geom_point() + theme_classic() +
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

####################

```

## Take a look at the bad fluxes <0.9 and have a high flux estimate 
# See if we want to remove these fluxes or if we want to try and recalculate these ones? 
```{r, echo=FALSE}

### pull out fluxes with R2 < 0.9 and put them in a list 
bad_plots <- unique(fluxes_weird$Plot_ID)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
bad_sample <- sample(bad_plots, 3, replace=FALSE)

#Loop to run five random plots subset above 
for(P in bad_sample){ 
  r1 <- subset(alldat_trimmed, Plot_ID == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, ch4)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth()
  print(p1)
  }

####################


```

## Look at the really high or really low ones to see if there is a timing issue or something
```{r, echo=FALSE}

### pull out fluxes in the 5% and 95% tails to look at really high and really low 
tail_plots <- unique(tails$Plot_ID)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
tail_sample <- sample(tail_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in tail_sample){ 
  r1 <- subset(alldat_trimmed, Plot_ID == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, ch4)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

#would be nice to put on the plot the flux estimate and if it is a high or low or something 
####################


```

## Check all fluxes
```{r, echo=FALSE}

##Put all fluxes together
fluxes_all <- as.data.frame(subset(allflux, select = Plot_ID:TIMESTAMP_max))

all_plots <- unique(fluxes_all$Plot_ID)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
all_sample <- sample(all_plots, 5, replace=FALSE)

#Loop to run five random plots subset above
for(P in all_sample){
  r1 <- subset(alldat_trimmed, Plot_ID == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, ch4)) + geom_point() + theme_classic() +
  scale_x_datetime(breaks = "15 sec") +
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth(color="red")
  print(p1)
  }

####################
```

## QAQC - Flags

```{r, echo=FALSE}

# Adding CH4 intital per Plot 
 allflux<- allflux %>%
  left_join(df_CH4_in, by = "Plot_ID") 

# QAQC flags 
fluxes_flagged <- allflux %>%
  mutate(
    # model fit flags
    CH4_lin_pass  = (lin_r.squared  >= 0.85 | lin_RMSE  < 5) & !is.na(lin_r.squared),
    CH4_HM81_pass = (HM81_r.squared >= 0.85 | HM81_RMSE < 5) & !is.na(HM81_r.squared),
    CH4_poly_pass = (poly_r.squared >= 0.85 | poly_RMSE < 5) & !is.na(poly_r.squared),
    # initial conc flag (CH4 initial conc)
    CH4_init_conc        = CH4_initial > 2600 | is.na(lin_flux.estimate),
    # rejection flags
    CH4_rejected_model  = !(CH4_lin_pass | CH4_HM81_pass | CH4_poly_pass)
  )

# percentage of problems
r2_summary <- fluxes_flagged %>%
  summarise(
    CH4_total_obs          = n(),
    CH4_n_rejected_model   = sum(CH4_rejected_model,    na.rm = TRUE),
    CH4_pct_rejected_model = 100 * CH4_n_rejected_model / CH4_total_obs,
    CH4_n_rejected_init_conc =  sum(CH4_init_conc,    na.rm = TRUE),       
    CH4_perct_init_conc    = 100 * CH4_n_rejected_init_conc / CH4_total_obs,
    CH4_n_lin_pass         = sum(CH4_lin_pass,  na.rm = TRUE),
    CH4_n_HM81_pass        = sum(CH4_HM81_pass, na.rm = TRUE),
    CH4_n_poly_pass        = sum(CH4_poly_pass, na.rm = TRUE)
  )

# formatted table for PDF
r2_summary %>%
  tidyr::pivot_longer(everything(), names_to = "Metric", values_to = "Value") %>%
  mutate(
    Metric = recode(Metric,
      CH4_total_obs          = "Total observations",
      CH4_n_rejected_init_conc = "Rejected - [CH4] (n)" ,
      CH4_perct_init_conc    = "Rejected — [CH4] (%)",
      CH4_n_rejected_model   = "Rejected — poor model fit (n)",
      CH4_pct_rejected_model = "Rejected — model (%)",
      CH4_n_lin_pass         = "Passed — Linear model (n)",
      CH4_n_HM81_pass        = "Passed — HM81 model (n)",
      CH4_n_poly_pass        = "Passed — Polynomial model (n)"
    ),
    Value = ifelse(Metric == "Rejected — model (%)",
                   round(Value, 1),
                   as.integer(Value))
  ) %>%
  kable(
    caption  = "QA/QC CH4 Flux Summary",
    align    = c("l", "r"),
    booktabs = TRUE,
    format   = "latex"
  ) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"),
    full_width    = FALSE
  ) %>%
  pack_rows("Rejections",        2, 5) %>%
  pack_rows("Model performance", 6, 8)

# R² vs RMSE diagnostic plot
# pivot_longer keeps R²/RMSE columns and uses CH4_*_pass as the pass indicator
set.seed(42)
plot_data <- fluxes_flagged %>%
  slice_sample(n = min(15000, nrow(.))) %>%
  select(
    lin_r.squared,  lin_RMSE,  CH4_lin_pass,
    HM81_r.squared, HM81_RMSE, CH4_HM81_pass,
    poly_r.squared, poly_RMSE, CH4_poly_pass
  ) %>%
  rename(
    lin_pass  = CH4_lin_pass,
    HM81_pass = CH4_HM81_pass,
    poly_pass = CH4_poly_pass
  ) %>%
  pivot_longer(cols = everything(),
    names_to      = c("model", ".value"),
    names_pattern = "(lin|HM81|poly)_(.*)"
  ) %>%
  rename(r2 = r.squared) %>%
  mutate(
    model  = recode(model, lin = "Linear", HM81 = "HM81", poly = "Polynomial"),
    status = case_when(
      pass                 ~ "Accepted",
      TRUE                 ~ "Rejected"
    )
  ) %>%
  filter(!is.na(r2), !is.na(RMSE))

#graphing
QAQC_plot <- ggplot(plot_data, aes(x = r2, y = RMSE, color = model, shape = status)) +
  geom_point(size = 2) +
  geom_vline(xintercept = 0.85, linetype = "dashed", color = "black", linewidth = 0.8) +
  geom_hline(yintercept = 5,    linetype = "dashed", color = "black", linewidth = 0.8) +
  annotate("text", x = 0.7, y = max(plot_data$RMSE, na.rm = TRUE) * 0.95,
           label = "R² = 0.85", hjust = -0.1, size = 3.5, fontface = "bold") +
  annotate("text", x = 0.1, y = 5,
           label = "RMSE = 5", vjust = -0.5, size = 3.5, fontface = "bold") +
  scale_color_manual(values = c(
    "Linear"     = "#E41A1C",
    "HM81"       = "#377EB8",
    "Polynomial" = "#4DAF4A"
  )) +
  scale_shape_manual(values = c(
    "Accepted"     = 16,
    "Rejected"     = 4
  )) +
  labs(
    title = "R² vs RMSE — CH4 Flux QA/QC",
    x     = "R²",
    y     = "RMSE",
    color = "Model",
    shape = "Status"
  ) +
  theme_bw(base_size = 13) +
  theme(legend.position = "right")

print(QAQC_plot)
```

## Remove any fluxes that do not fit criteria 
```{r, echo=FALSE}

#Any others that are not in the weird category that need to be removed: 
  #also removing ones that are crazy off but were not in the fluxes_weird dataframe 

manual_reject  <- c(
                      )

fluxes_flagged <- fluxes_flagged  %>%
  mutate(
    CH4_rejected_manual  = Plot_ID %in% manual_reject,
    CH4_flux_rejected   = CH4_rejected_model |CH4_rejected_manual)

# View the result
head(fluxes_flagged)

```


## Do some visualization of QAQC plots to assess fluxes  
```{r, echo=FALSE}

p1 <- ggplot(fluxes_flagged , aes(lin_flux.estimate, rob_flux.estimate, color = Plot_ID)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  labs(title = "A") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Robust Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot_ID")) +
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p1


p2 <- ggplot(fluxes_flagged , aes(lin_r.squared, poly_r.squared, color = Plot_ID)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  xlim(0,1) + ylim(0,1) +
  labs(title = "B") + 
  ylab(expression(paste("R"^2* " Second-Order Polynomial"))) +
  xlab(expression(paste("R"^2* " Linear Model"))) +
  guides(color = guide_legend(title = "Plot_ID")) + 
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p2

p3 <- ggplot(fluxes_flagged , aes(lin_flux.estimate, HM81_flux.estimate, color = lin_r.squared )) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  #xlim(0,50) + #ylim(0,50) +
  labs(title = "C") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") HM1981 Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot_ID")) + 
  theme(legend.position="right") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p3

#ggarrange(p1, p2, p3, ncol=3, nrow=1) #, common.legend = TRUE, legend="none")
```

## Read out final dataframe and export for use elsewhere 
```{r, echo=FALSE}

allflux_renamed <- fluxes_flagged %>%
  separate(
    Plot_ID,
    into = c("Year", "Month_num", "Day", "Site", "Zone_raw", "Rest"),
    sep = "_",
    fill = "right",
    extra = "merge",
    remove = FALSE
  ) %>%
  mutate(
    # (TR / UP / W)
    Zone = str_extract(Zone_raw, "UP|TR|W"),

    # collar ID - number after zone
    Replicate = as.integer(
      str_extract(Plot_ID, "(?<=_(UP|TR|W)_)[0-9]+")
    ),

    # meansurement replicate _1, _2)
    RepNumber = as.integer(
      str_extract(Plot_ID, "[0-9]+$")
    ),

    Month = month.abb[as.integer(Month_num)]
  ) %>%
  rename(Plot = Plot_ID) %>%
  select(
    RepNumber, Month, Site, Zone, Replicate, Plot,
    everything(),
    -Year, -Day, -Month_num, -Zone_raw, -Rest
  )

#checkf for labels 
unique(allflux_renamed$Zone)
unique(allflux_renamed$Replicate)
unique(allflux_renamed$RepNumber)


#export as a csv to the folder that you want 
write.csv(
  allflux_renamed,
  file = "Processed Data/COMPASS_Synoptic_LE_SmartChamber_Fluxes_Final_2024_CO2.csv",
  row.names = FALSE
)

```



