---
title: "COMPASS_Synoptic_LE_Tree_CO2_Flux_Calcs_2023b"
author: "Roberta Peixoto"
date: today
format:
  pdf:
    toc: true
    number-sections: true
output-dir: "2023/PDF"
---

## Load Required Packages

```{r packages, include = FALSE}
# File reading
library(here)
library(readr)
library(fluxfinder)
#library(readxl)
# Data processing
library(dplyr)
library(stringr)
library(lubridate)
# Statistics
library(broom)
# Plotting
library(ggplot2)
#library(ggpubr)

# invisible(lapply(c( "fluxfinder", "readxl",
#                     "dplyr", "ggplot2", "ggpubr", "stringr",
#                     "here", "broom", "tibble",
#                     "matrixStats", "gridExtra", "grid", "knitr", "kableExtra", "arrow"), 
#                  library, character.only = TRUE))
```

## Pull in raw datafiles

```{r read-data}
# Defining where the raw data is

DATA_DIR_ROOT <- here::here(
    "Tree_Fluxes_Synoptic_LE",
    "2023",
    "Raw Data"
)

now <- function() format(Sys.time(), "%Y-%m-%d %H:%M:%S")
message(now(), " | Script started")

# Find the zip files
message(now(), " | Looking for ZIP files in: ", DATA_DIR_ROOT)
zip_files <- list.files(
    DATA_DIR_ROOT,
    pattern = "\\.zip$",
    full.names = TRUE)

message(now(), " | ZIP files found: ", length(zip_files))
print(basename(zip_files))
if (length(zip_files) == 0) {
    stop("No ZIP files found. Check DATA_DIR_ROOT.")
}

cache <- list()

# Reader function
read_li7810_zip <- function(zip_file, tz = "US/Eastern") {
    message(now(), " | Processing ZIP: ", basename(zip_file))
    if (zip_file %in% names(cache)) {
        message("\t-> Using cached version")
        return(cache[[zip_file]])  }
    
    tmp_dir <- tempdir()
    # Technically this next step isn't necessary, because tempdir()
    # gives us a directory in the system "temporary" space, which
    # will automatically get cleaned up. But still not a bad idea!
    on.exit(unlink(tmp_dir, recursive = TRUE), add = TRUE)
    
    # Extract and list .data files
    unzip(zip_file, exdir = tmp_dir)
    data_files <- list.files(tmp_dir,
                             pattern = "\\.data$",
                             full.names = TRUE,
                             recursive = TRUE)
    
    message("\t-> .data files found: ", length(data_files))
    if (length(data_files) == 0) {
        stop("No .data files inside ZIP: ", zip_file) }
    
    dat_list <- lapply(data_files, function(f) {
        message("\tReading: ", basename(f))
        ffi_read_LI7810(f) %>% 
            as_tibble() %>% 
            # these are big files, so immediately drop unneeded columns;
            # this saves memory and speeds operations
            select(-DATAH, -SECONDS, -NANOSECONDS, -NDX, -REMARK, 
                   -INPUT_VOLTAGE, -RING_DOWN_TIME, -PHASE_ERROR, 
                   -LASER_T_SHIFT, -LASER_PHASE_P, -LASER_T,
                   -RESIDUAL, -THERMAL_ENCLOSURE_T)
    })
    
    dat <- bind_rows(dat_list) %>%
        # Some machines can be in the wrong time zone; 
        # ensure everything is EST
        mutate(TIMESTAMP = with_tz(TIMESTAMP, tz))
    cache[[zip_file]] <<- dat
    message("\t-> ZIP finished: ", basename(zip_file))
    return(dat)
}

# ACTUAL EXECUTION 
message(now(), " | Starting data read")
alldat <- lapply(zip_files, read_li7810_zip) %>% bind_rows()
message(now(), " | DONE")

# Show result
print(head(alldat))

# REMOVING WEIRD VERY LOW VALUES ON "2022-12-20 11:41:00" co2=-4000 licor bug???
na_co2 <- is.na(alldat$CO2)
message("Removing ", sum(na_co2), " NA CO2 values")
alldat <- alldat[!na_co2,]
weird_co2 <- alldat$CO2 <= -1000 | alldat$CO2 >= 10000
message("Removing ", sum(weird_co2), " weird CO2 values")
alldat <- alldat[!weird_co2,]

message("\tFinal number of rows: ", nrow(alldat))
```

## Read in metadata file

```{r read-metadaa}
md <- "Raw Data/COMPASS_SynopticLE_TreeFlux_Metadata_2023.csv"
dat <- read_csv(md)

metadat <- dat %>%
    # Remove rows without date
    filter(!is.na(Date)) %>%
    # Keep only desired instruments
    filter(SN %in% c("TG10-01448", "TG10-01426", "TG10-01028")) %>%
    # Normalize plot labels
    mutate(Plot_raw = toupper(Plot),
           # Extract plot components 
           #Site = str_extract(Plot_raw, "^[A-Z]{3}"),
           #Zone = str_extract(Plot_raw, "(TR|UP)"), # somehow all of the UP PTR measurements turned into TR
           Site = toupper(Site),
           Zone = toupper(Zone), 
           Replicate = str_extract(Plot_raw, "(?<=TR|UP)[0-9]+"),
           # Letter ONLY if it comes AFTER a number
           Letter = str_extract(Plot_raw, "(?<=[0-9])[A-Z]$"),
           # Observation length
           Obs_length = as.numeric(`Stop time` - Time_start, units = "secs"),
           # Construct timestamp
           Date = mdy(Date, tz = "EST"),
           TIMESTAMP = Date + Time_start,
        end_ts = TIMESTAMP + Obs_length,
           # Sequential ID
           Seq = row_number(),
           # INAL standardized Plot label
           Plot = if_else(
               is.na(Letter),
               paste(Date, Site, Zone, Replicate, Seq, sep = "_"),
               paste(Date, Site, Zone, Replicate, Letter, Seq, sep = "_")) )

# Quick check
head(metadat$Plot)
```

## Check for start time overlaps or missing values

```{r overlap-check}

# Check if there is any start and stop time overlap from the field notes
overlaps_detected <- metadat %>%
    arrange(SN, TIMESTAMP) %>%
    group_by(SN) %>%
    mutate(next_plot = lead(Plot),
           next_start_ts = lead(TIMESTAMP)) %>%
    # is the next timestamp is less than the current end timestamp?
    filter(!is.na(next_start_ts) & next_start_ts <= end_ts) %>%
    select(SN, Plot,
           TIMESTAMP, end_ts,
           next_plot, next_start_ts,
           Obs_length)

print(overlaps_detected)
# TODO -- should the code stop() if overlaps are detected?

# Check problems with obs_length - this is a frequent bug for manual field notes 
obs_length_problems <- metadat %>%
    filter(is.na(Obs_length) | Obs_length <= 0) %>%
    select(SN, Date, Plot, TIMESTAMP, end_ts, Obs_length)
if(nrow(obs_length_problems) > 0) {
    print(obs_length_problems)
    stop("Observations length <=0 detected")
}

# Check for duplicated plot entries
if(any(duplicated(metadat$Plot))) {
    stop("There are duplicated Plot entries!")
}

# checking unique measurment start time and stop time
alldat %>%
    filter(SN == "TG10-01448",
           TIMESTAMP > ymd_hms("2023-01-05 13:28:00"),
           TIMESTAMP < ymd_hms("2023-01-05 13:33:00")) %>%
    ggplot(aes(x = TIMESTAMP, y = CO2)) +
    geom_point()#+ylim(300,500)

```

## Match metadata to raw data

```{r data-metadata-match}
PROCESSED_FILE_BASE <- "alldat_matched_2023"
matched_list <- list()
for(sn in unique(alldat$SN)) {
    message("\tMatching ", sn)
    
    # Pre-processed data available?
    pf <- paste0(PROCESSED_FILE_BASE, "_", sn, ".rds")
    if(file.exists(pf)) {
        message("\tReading existing processed file ", pf)
        matched_list[[sn]] <- readRDS(pf)
        next
    }
    
    # No processed file available for this NS
    alldat %>% 
        filter(SN == sn) ->
        dat_sn
    message("\tData rows = ", nrow(dat_sn))
    metadat %>% 
        filter(SN == sn) %>% 
        arrange(Date, Time_start) %>%
        mutate(metadat_row = row_number()) ->
        metadat_sn
    message("\tMetadata rows = ", nrow(metadat_sn))
    
    if(nrow(metadat_sn) == 0) {
        message("\tNo metadata rows! Moving on")
        next
    }
    # The actual match step
    dat_sn$metadat_row <- ffi_metadata_match(
        data_timestamps = dat_sn$TIMESTAMP,
        start_dates     = metadat_sn$Date,
        start_times     = as.character(metadat_sn$Time_start),
        obs_lengths     = metadat_sn$Obs_length)
    # Drop unmatched rows
    n_before <- nrow(dat_sn)
    dat_i <- dat_sn %>% filter(!is.na(metadat_row))
    message("\tMatched rows: ", nrow(dat_sn), " / ", n_before)
    
    # ...and then copy needed metadata into data frame
    dat_sn %>%
        mutate(Plot      = metadat_sn$Plot[metadat_row],
               Site      = metadat_sn$Site[metadat_row],
               Zone      = metadat_sn$Zone[metadat_row],
               Volume    = metadat_sn$Volume[metadat_row],
               # TODO: this column is not in the metadata
               #Temp      = metadat_sn$Temp[metadat_row],
               Area      = metadat_sn$Area[metadat_row],
               Dead_band = metadat_sn$Dead_band[metadat_row]) ->
        matched_list[[sn]]
    
    message("\tSaving ", pf)
    saveRDS(matched_list[[sn]], pf)
}

# All done! Pull data together and do final sanity checks
message("Matching finished successfully")
alldat_matched <- bind_rows(matched_list)
stopifnot(
    !anyNA(alldat_matched$Plot),
    nrow(alldat_matched) > 0
)

message("Matched dataset rows: ", nrow(alldat_matched))
head(alldat_matched)

```

##Extracting initial CO2 concentration

```{r}

## Extract initial CO2 concentration (t0) for each Plot for QAQC later
#saving df to use after flux calculation
df_CO2_in <- alldat_matched %>%
    arrange(Plot, TIMESTAMP) %>%
    group_by(Plot) %>%
    summarise(CO2_initial = first(CO2), .groups = "drop")  

```

## Creating temperature csv from the L2-weather station in the wetland

```{#r}

# Getting Air temp from the weather station 
## Read L2 temperature data from ZIP stored outside GitHub,
## extract desired variable, save lightweight CSV,
## and remove temporary unzipped files afterwards

## Read extracted L2 Parquet files and save temperature time series
BASE_DIR <- "D:/COMPASS_L2-v2-1/v2-1"   
SITE     <- "(OWC|CRC|PTR)"             
VARIABLE <- "wx-tempavg15"                  

OUTFILE  <- "Raw Data/L2_temp_timeseries.csv"

# Build regex that matches REAL L2 filenames
regex <- paste0(SITE, ".*", VARIABLE,".*_L2_v2-1.*\\.parquet$")

# Locate Parquet files
files <- list.files(BASE_DIR,pattern = regex,recursive = TRUE,full.names = TRUE)

message("Files found: ", length(files))
print(basename(files))

# Fail fast if nothing is found
stopifnot(length(files) > 0)

# Read and combine Parquet files
dat <- lapply(files, function(f) {
message("Reading ", basename(f))
read_parquet(f)
})

dat <- bind_rows(dat)

# Optional QA filtering (only if column exists)
if ("QA_flag" %in% names(dat)) {
dat <- dat %>% filter(QA_flag == 0)
}

# Keep only necessary columns
temps <- dat %>%
select(TIMESTAMP, Plot, Site, Value) %>%
rename(Temp = Value)

head(temps)

# Save lightweight output only
write.csv(
temps,
file = OUTFILE,
row.names = FALSE
)
```


## Reading Temp df to merge to metadat
```{r}
# reading temp 
temps <- read.csv("Raw Data/L2_temp_timeseries.csv")

# make sure timestamps are in the correct format
# convert to tibble early to avoid data.table list-column issues
alldat_matched <- as_tibble(alldat_matched) %>%
    mutate(across(where(is.list), ~ as.numeric(unlist(.))))

alldat_matched$TIMESTAMP <- as.POSIXct(alldat_matched$TIMESTAMP, format = "%Y-%m-%d %H:%M:%OS")
alldat_matched$Dead_band <- as.numeric(alldat_matched$Dead_band)

# fix midnight timestamps in L2 data
temps$TIMESTAMP[!grepl(":", temps$TIMESTAMP)] <- 
    paste0(temps$TIMESTAMP[!grepl(":", temps$TIMESTAMP)], " 00:00:00")
temps$TIMESTAMP <- as.POSIXct(temps$TIMESTAMP, format = "%Y-%m-%d %H:%M:%OS") 

# time rounding
alldat_matched$TIMESTAMP_round <- round_date(alldat_matched$TIMESTAMP, "60 mins")
temps$TIMESTAMP_round           <- round_date(temps$TIMESTAMP, "60 mins")

# one observation per hour
temps_clean <- temps %>% 
    group_by(TIMESTAMP_round, Site) %>%
    summarize(Temp = mean(Temp), .groups = "drop") 

# join weather station temp
alldat_matched <- alldat_matched %>% 
    left_join(temps_clean, by = c("Site", "TIMESTAMP_round"))

# ── DIAGNOSTIC: which plots are missing Temp after join ──────────────────────
temp_missing_check <- alldat_matched %>%
    group_by(Plot, Site) %>%
    summarise(
        n_obs       = n(),
        n_temp_na   = sum(is.na(Temp)),
        pct_temp_na = round(100 * n_temp_na / n_obs, 1),
        month       = month(first(TIMESTAMP)),
        .groups     = "drop"
    ) %>%
    filter(n_temp_na > 0)

message("Plots with missing Temp: ", nrow(temp_missing_check))
print(temp_missing_check)

# ── IMPUTATION: monthly mean per site as fallback ────────────────────────────
monthly_temp <- alldat_matched %>%
    mutate(month = month(TIMESTAMP)) %>%
    filter(!is.na(Temp)) %>%
    group_by(Site, month) %>%
    summarise(Temp_monthly_mean = mean(Temp), .groups = "drop")

alldat_matched <- alldat_matched %>%
    mutate(
        month      = month(TIMESTAMP),
        Temp       = as.numeric(Temp)
    ) %>%
    left_join(monthly_temp, by = c("Site", "month")) %>%
    mutate(
        Temp_source = case_when(
            !is.na(Temp)              ~ "weather_station",
            !is.na(Temp_monthly_mean) ~ "monthly_mean_imputed",
            TRUE                      ~ "still_missing"
        ),
        Temp = coalesce(Temp, Temp_monthly_mean)
    ) %>%
    select(-Temp_monthly_mean, -month)

# ── FINAL CHECK ──────────────────────────────────────────────────────────────
message("Temp source summary:")
alldat_matched$Temp_source <- as.character(alldat_matched$Temp_source)
table(alldat_matched$Temp_source)

still_missing <- alldat_matched %>%
    filter(Temp_source == "still_missing") %>%
    distinct(Plot, Site)

if (nrow(still_missing) > 0) {
    message("WARNING: ", nrow(still_missing), " plots still have no Temp after imputation:")
    print(still_missing)
} else {
    message("All plots have Temp after imputation.")
}
```

## fixing some Deadbands and stop time

```{r}

# Table of plots with custom deadband adjustments
time_overrides <- tibble::tibble(
    Plot             = c(#"7/13/2023_CRC_TR_315_115",
        #"7/27/2023_PTR_UP_8_125",
        #"3/28/2023_OWC_TR_127_170",
        #"5/10/2023_OWC_UP_103_73",
        #"7/13/2023_CRC_TR_314_114",
        #"8/3/2023_CRC_TR_314_133",
        #"7/13/2023_CRC_TR_316_116",
        #"6/29/2023_CRC_TR_311_104" ,
        "2/24/2023_PTR_UP_4_100"),
    dead_band_manual = c(#8,9,22,16,8,16,3,34,
        22),
    stop_offset_secs = NA_real_ )

#apply new incubation period
alldat_trimmed <- alldat_matched %>%
    left_join(time_overrides, by = "Plot") %>%
    group_by(Plot) %>%
    mutate(
        t_start_original = min(TIMESTAMP),
        t_end_original   = max(TIMESTAMP),
        
        # deadband manual or metadat
        t_start_adjusted = if_else(
            !is.na(dead_band_manual),
            t_start_original + dead_band_manual,
            t_start_original + as.numeric(coalesce(first(Dead_band), 0))
        ),
        
        # new offset
        t_end_adjusted = if_else(
            !is.na(stop_offset_secs),
            t_end_original + stop_offset_secs,
            t_end_original
        )
    ) %>%
    filter(TIMESTAMP >= t_start_adjusted,
           TIMESTAMP <= t_end_adjusted) %>%
    ungroup() %>%
    select(-t_start_original, -t_end_original,
           -t_start_adjusted, -t_end_adjusted,
           -dead_band_manual, -stop_offset_secs)

# alldat_trimmed %>%
#     filter(Plot == "8/17/2022_OWC_UP_6_147") %>%
#     ggplot(aes(TIMESTAMP, CO2)) +
#     geom_point() + theme_classic() 

```

## Calculate fluxes

```{r}

#change the units
alldat_trimmed$CO2_umol <- ffi_ppm_to_umol(alldat_trimmed$CO2, 
                                           volume = alldat_trimmed$Volume, # m3
                                           temp = alldat_trimmed$Temp)    # degrees C from the weather station
#> Assuming atm = 101325 Pa
#> Using R = 8.31446261815324 m3 Pa K-1 mol-1

alldat_trimmed$CO2_umol_m2 <- alldat_trimmed$CO2_umol / 0.16

fluxes <- ffi_compute_fluxes(alldat_trimmed ,
                             group_column = "Plot", 
                             time_column = "TIMESTAMP", 
                             gas_column = "CO2_umol_m2", 
                             dead_band = 0)
head(fluxes)

#merging fluxes with initial CO2 concentration
fluxes <- fluxes %>%
    left_join(df_CO2_in,by = "Plot")

#take a look at the fluxes we calculated with R2 
ggplot(fluxes, aes(Plot, lin_flux.estimate, color = lin_r.squared)) +
    geom_point(size=4) + theme_classic() + 
    geom_linerange(aes(ymin = lin_flux.estimate- lin_flux.std.error,
                       ymax = lin_flux.estimate+ lin_flux.std.error)) +
    scale_color_gradientn(colours = c("darkred",  "red", "white","lightblue", "darkblue"),
                          values = c(0, 0.5, 0.6, 0.7, 1)) +
    ylab("CO2 flux (umol/m2/s)")

#sort fluxes by R2 
fluxes_good <- as.data.frame(subset(fluxes, lin_r.squared >= .90, select = Plot:TIMESTAMP_max))
fluxes_lowr2 <- as.data.frame(subset(fluxes, lin_r.squared < .90 & lin_flux.estimate < 1, select = Plot:TIMESTAMP_max))
fluxes_weird <- as.data.frame(subset(fluxes, lin_r.squared < .90, select = Plot:TIMESTAMP_max))

#find fluxes that represent the first and fourth quartile of the data 
low <- quantile(fluxes$lin_flux.estimate, 0.05) 
high <- quantile(fluxes$lin_flux.estimate, 0.95) 

fluxes_low <- as.data.frame(subset(fluxes, lin_flux.estimate <= low, select = Plot:TIMESTAMP_max))
fluxes_high <- as.data.frame(subset(fluxes, lin_flux.estimate >= high, select = Plot:TIMESTAMP_max))

tails <- rbind(fluxes_low, fluxes_high)
```

## check the fluxes for which I changed the start time due to lags or high values (recorded in notes of metadata)

```{r}

# grab trim limits for each overridden Plot
trim_limits <- alldat_trimmed %>%
    filter(Plot %in% time_overrides$Plot) %>%
    group_by(Plot) %>%
    summarise(
        t_start_trim = min(TIMESTAMP),
        t_end_trim   = max(TIMESTAMP) )

# use full alldat_matched + trim limits to classify each point
r_all <- alldat_matched %>%
    filter(Plot %in% time_overrides$Plot) %>%
    left_join(trim_limits, by = "Plot") %>%
    group_by(Plot) %>%
    mutate(
        time_sec = as.numeric(difftime(TIMESTAMP, min(TIMESTAMP), units = "secs")),
        point_status = case_when(
            TIMESTAMP < t_start_trim ~ "Discarded (deadband)",
            TIMESTAMP > t_end_trim   ~ "Discarded (stop trim)",
            TRUE                     ~ "Used"  )) %>%
    ungroup()

# plot all overridden plots
ggplot(r_all, aes(x = TIMESTAMP, y = CO2, color = point_status)) +
    geom_point(size = 1.8) +
    facet_wrap(~ Plot, scales = "free", ncol = 2) +
    scale_color_manual(
        values = c(
            "Used"                  = "#E41A1C",
            "Discarded (deadband)"  = "grey60",
            "Discarded (stop trim)" = "grey20"
        )
    ) +
    theme_classic() +
    ylab(expression(paste(CO[2], " (ppm)"))) +
    xlab("Time") +
    labs(
        title = "Incubation Period Adjustments",
        color = "Point Status") +
    theme(legend.position = "bottom")
```

## Spot check a few fluxes with R2 \> 0.9

```{r, echo=FALSE}

### pull a random sample of good fluxes to check and plot 
good_plots <- unique(fluxes_good$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
good_sample <- sample(good_plots, 9, replace=FALSE)

#Loop to run five random plots subset above 
for(P in good_sample){ 
    r1 <- subset(alldat_trimmed, Plot == P & !is.na(metadat_row))
    p1 <- ggplot(r1, aes(as.POSIXct(TIMESTAMP), CO2)) + geom_point() + 
        scale_x_datetime(breaks = "15 sec") + 
        theme_classic() +
        ylab(expression(paste( CO [2], " (ppm)"))) +
        xlab("Time") + labs(title = P) + 
        geom_smooth()
    print(p1)
}

####################

```

## Take a look at the fluxes \<0.9 and a flux estimate below 1

```{r, echo = FALSE}

####Do we just leave these alone because the fluxes are so low?####

### pull out fluxes with R2 < 0.9 and put them in a list 
low_plots <- unique(fluxes_lowr2$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
low_sample <- sample(low_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in low_sample){ 
    r1 <- subset(alldat_trimmed , Plot == P & !is.na(metadat_row))
    p1 <- ggplot(r1, aes(TIMESTAMP, CO2)) + geom_point() + theme_classic() +
        scale_x_datetime(breaks = "15 sec") + 
        ylab(expression(paste( CO [2], " (ppm)"))) +
        xlab("Time") + labs(title = P) + 
        geom_smooth()
    print(p1)
}

####################
```

## Take a look at the bad fluxes \<0.9 and have a high flux estimate

## See if we want to remove these fluxes or if we want to try and recalculate these ones?

```{r}
### pull out fluxes with R2 < 0.9 and put them in a list 
bad_plots <- unique(fluxes_weird$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
bad_sample <- sample(bad_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in bad_sample){ 
    r1 <- subset(alldat_trimmed, Plot == P & !is.na(metadat_row))
    p1 <- ggplot(r1, aes(TIMESTAMP, CO2)) + geom_point() + theme_classic() + 
        scale_x_datetime(breaks = "15 sec") + 
        ylab(expression(paste( CO[2], " (ppm)"))) +
        xlab("Time") + labs(title = P) +
        geom_smooth()
    print(p1)
}

####################
```

## Look at the really high or really low ones to see if there is a timing issue or something

```{r}
### pull out fluxes in the 5% and 95% tails to look at really high and really low 
tail_plots <- unique(tails$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
tail_sample <- sample(tail_plots, 2, replace=FALSE)

#Loop to run five random plots subset above 
for(P in tail_sample){ 
    r1 <- subset(alldat_trimmed, Plot == P & !is.na(metadat_row))
    p1 <- ggplot(r1, aes(TIMESTAMP, CO2)) + geom_point() + theme_classic() + 
        scale_x_datetime(breaks = "15 sec") + 
        ylab(expression(paste( CO [2], " (ppm)"))) +
        xlab("Time") + labs(title = P) + 
        geom_smooth()
    print(p1)
}

#would be nice to put on the plot the flux estimate and if it is a high or low or something 
####################
```

## Check all fluxes

```{r}
##Put all fluxes together
#fluxes_all <- as.data.frame(subset(fluxes, select = Plot:TIMESTAMP_max))

all_plots <- unique(fluxes$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
all_sample <- sample(all_plots, 5, replace=FALSE)

#Loop to run five random plots subset above
for(P in all_sample){
    r1 <- subset(alldat_trimmed, Plot== P & !is.na(metadat_row))
    p1 <- ggplot(r1, aes(TIMESTAMP, CO2)) + geom_point() + theme_classic() +
        scale_x_datetime(breaks = "15 sec") +
        ylab(expression(paste( CO [2], " (ppm)"))) +
        xlab("Time") + labs(title = P) +
        geom_smooth(color="red")
    print(p1)
}

####################
```


## QAQC - Flags

```{r}

# QAQC flags 
fluxes_flagged <- fluxes %>%
    mutate(
        # model fit flags
        CO2_lin_pass  = (lin_r.squared  >= 0.85 | lin_RMSE  < 5) & !is.na(lin_r.squared),
        CO2_HM81_pass = (HM81_r.squared >= 0.85 | HM81_RMSE < 5) & !is.na(HM81_r.squared),
        CO2_poly_pass = (poly_r.squared >= 0.85 | poly_RMSE < 5) & !is.na(poly_r.squared),
        # negative flux flag
        CO2_neg_flux       = lin_flux.estimate < 0 | is.na(lin_flux.estimate),
        # rejection flags
        CO2_rejected_model  = !(CO2_lin_pass | CO2_HM81_pass | CO2_poly_pass),
        CO2_rejected_negflux = CO2_neg_flux,
        CO2_flux_rejected   = CO2_rejected_model | CO2_rejected_negflux
    )

# percentage of problems
r2_summary <- fluxes_flagged %>%
    summarise(
        CO2_total_obs          = n(),
        CO2_n_rejected_model   = sum(CO2_rejected_model,   na.rm = TRUE),
        CO2_n_rejected_negflux = sum(CO2_rejected_negflux, na.rm = TRUE),
        CO2_n_rejected_both    = sum(CO2_rejected_model & CO2_rejected_negflux, na.rm = TRUE),
        CO2_n_rejected_total   = sum(CO2_flux_rejected,    na.rm = TRUE),
        CO2_pct_rejected_total = 100 * CO2_n_rejected_total / CO2_total_obs,
        CO2_n_lin_pass         = sum(CO2_lin_pass,  na.rm = TRUE),
        CO2_n_HM81_pass        = sum(CO2_HM81_pass, na.rm = TRUE),
        CO2_n_poly_pass        = sum(CO2_poly_pass, na.rm = TRUE)
    )

# formatted table for PDF
r2_summary %>%
    tidyr::pivot_longer(everything(), names_to = "Metric", values_to = "Value") %>%
    mutate(
        Metric = recode(Metric,
                        CO2_total_obs          = "Total observations",
                        CO2_n_rejected_model   = "Rejected — poor model fit (n)",
                        CO2_n_rejected_negflux = "Rejected — negative CO2 flux (n)",
                        CO2_n_rejected_both    = "Rejected — both reasons (n)",
                        CO2_n_rejected_total   = "Rejected — total (n)",
                        CO2_pct_rejected_total = "Rejected — total (%)",
                        CO2_n_lin_pass         = "Passed — Linear model (n)",
                        CO2_n_HM81_pass        = "Passed — HM81 model (n)",
                        CO2_n_poly_pass        = "Passed — Polynomial model (n)"
        ),
        Value = ifelse(Metric == "Rejected — total (%)",
                       round(Value, 1),
                       as.integer(Value)) ) %>%
    kable(
        caption  = "QA/QC CO2 Flux Summary",
        align    = c("l", "r"),
        booktabs = TRUE,
        format   = "latex") %>%
    kable_styling(
        latex_options = c("striped", "hold_position"),
        full_width    = FALSE ) %>%
    pack_rows("Rejections",        2, 6) %>%
    pack_rows("Model performance", 7, 9)

# R² vs RMSE diagnostic plot
# pivot_longer keeps R²/RMSE columns and uses CO2_*_pass as the pass indicator
set.seed(42)
plot_data <- fluxes_flagged %>%
    slice_sample(n = min(15000, nrow(.))) %>%
    select(
        lin_r.squared,  lin_RMSE,  CO2_lin_pass,
        HM81_r.squared, HM81_RMSE, CO2_HM81_pass,
        poly_r.squared, poly_RMSE, CO2_poly_pass,
        CO2_rejected_negflux
    ) %>%
    # rename pass columns to match model prefix before pivoting
    rename(
        lin_pass  = CO2_lin_pass,
        HM81_pass = CO2_HM81_pass,
        poly_pass = CO2_poly_pass
    ) %>%
    pivot_longer(
        cols = -CO2_rejected_negflux,
        names_to  = c("model", ".value"),
        names_pattern = "(lin|HM81|poly)_(.*)"
    ) %>%
    rename(r2 = r.squared) %>%
    mutate(
        model  = recode(model, lin = "Linear", HM81 = "HM81", poly = "Polynomial"),
        status = case_when(
            CO2_rejected_negflux ~ "Negative CO2",
            pass                 ~ "Accepted",
            TRUE                 ~ "Rejected"
        )
    ) %>%
    filter(!is.na(r2), !is.na(RMSE))

#graphing
QAQC_plot <- ggplot(plot_data, aes(x = r2, y = RMSE, color = model, shape = status)) +
    geom_point(size = 2) +
    geom_vline(xintercept = 0.85, linetype = "dashed", color = "black", linewidth = 0.8) +
    geom_hline(yintercept = 5,    linetype = "dashed", color = "black", linewidth = 0.8) +
    annotate("text", x = 0.7, y = max(plot_data$RMSE, na.rm = TRUE) * 0.95,
             label = "R² = 0.85", hjust = -0.1, size = 3.5, fontface = "bold") +
    annotate("text", x = 0.1, y = 5,
             label = "RMSE = 5", vjust = -0.5, size = 3.5, fontface = "bold") +
    scale_color_manual(values = c(
        "Linear"     = "#E41A1C",
        "HM81"       = "#377EB8",
        "Polynomial" = "#4DAF4A"
    )) +
    scale_shape_manual(values = c(
        "Accepted"     = 16,
        "Rejected"     = 4,
        "Negative CO2" = 17 )) +
    labs(
        title = "R² vs RMSE — CO2 Flux QA/QC",
        x     = "R²",
        y     = "RMSE",
        color = "Model",
        shape = "Status"  ) +
    theme_bw(base_size = 13) +
    theme(legend.position = "right")

print(QAQC_plot)
```

## Adding manual flag to reject any fluxes that do not fit criteria

```{r}

# adding manual flags to the ones that are crazy off based on visual checks
manual_reject <- c("2/24/2023_PTR_UP_6_101"  #leak
) 

fluxes_flagged <- fluxes_flagged %>%
    mutate(
        CO2_rejected_manual  = Plot %in% manual_reject,
        CO2_flux_rejected   = CO2_rejected_model | CO2_rejected_negflux |CO2_rejected_manual)

# fluxes_flagged %>% 
#   filter(CO2_rejected_manual) %>% 
#   select(Plot, lin_r.squared, lin_RMSE, CO2_rejected_model, CO2_rejected_negCO2, CO2_rejected_manual, CO2_flux_rejected)

```

## Do some visualization of QAQC plots to assess fluxes

```{r}

allflux_filtered <-fluxes_flagged%>%filter(!CO2_flux_rejected)


p1 <- ggplot(allflux_filtered, aes(lin_flux.estimate, rob_flux.estimate, color = Plot)) +
    geom_point(size=4) + geom_abline() + 
    theme_classic() + 
    labs(title = "A") + 
    ylab(expression(paste( CO [2], " Flux (Umol m"^-2* " s"^-1*") Robust Method"))) +
    xlab(expression(paste( CO [2], " Flux (Umol m"^-2* " s"^-1*") Linear Method"))) +
    guides(color = guide_legend(title = "Plot")) +
    theme(legend.position="none") +
    theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
          axis.title.y = element_text(size=12), legend.text=element_text(size=12),
          panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
          aspect.ratio = 1)
p1

p2 <- ggplot(allflux_filtered, aes(lin_r.squared, poly_r.squared, color = Plot)) +
    geom_point(size=4) + geom_abline() + 
    theme_classic() + 
    xlim(0,1) + ylim(0,1) +
    labs(title = "B") + 
    ylab(expression(paste("R"^2* " Second-Order Polynomial"))) +
    xlab(expression(paste("R"^2* " Linear Model"))) +
    guides(color = guide_legend(title = "Plot_ID")) + 
    theme(legend.position="none") +
    theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
          axis.title.y = element_text(size=12), legend.text=element_text(size=12),
          panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
          aspect.ratio = 1)
p2

p3 <- ggplot(allflux_filtered, aes(lin_flux.estimate, HM81_flux.estimate, color = lin_r.squared )) +
    geom_point(size=4) + geom_abline() + 
    theme_classic() + 
    #xlim(0,50) + #ylim(0,50) +
    labs(title = "C") + 
    ylab(expression(paste( CO [2], " Flux (Umol m"^-2* " s"^-1*") HM1981 Method"))) +
    xlab(expression(paste( CO [2], " Flux (Umol m"^-2* " s"^-1*") Linear Method"))) +
    guides(color = guide_legend(title = "Plot_ID")) + 
    theme(legend.position="right") +
    theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
          axis.title.y = element_text(size=12), legend.text=element_text(size=12),
          panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
          aspect.ratio = 1)

p3

#ggarrange(p1, p2, p3, ncol=3, nrow=1, common.legend = TRUE, legend="none")
```

## Read out final dataframe and export for use elsewhere

```{r}

allflux_filtered_renamed <- fluxes_flagged %>%
    mutate(
        Date = mdy(str_extract(Plot, "^[0-9]+/[0-9]+/[0-9]{4}")),
        Month = month.abb[month(Date)],
        Site = str_extract(Plot, "(PTR|OWC|CRC)"),
        Zone = str_extract(Plot, "_TR_|_UP_") %>%
            str_remove_all("_"),
        Replicate = as.integer(
            str_extract(Plot, "(?<=_(TR|UP)_)[0-9]+")
        ),
        Row_ID = as.integer(
            str_extract(Plot, "[0-9]+$")) ) %>%
    
    # Optional: rebuild a clean Plot label
    mutate(
        Plot_clean = paste(
            format(Date, "%Y-%m-%d"),
            Site,
            Zone,
            Replicate,
            Row_ID,
            sep = "_"
        )
    ) %>%
    
    select(
        Plot_clean,
        Date,
        Month,
        Site,
        Zone,
        Replicate,
        Row_ID,
        everything(),
        -Plot
    )




#export as a csv to the folder that you want
write.csv(allflux_filtered_renamed, file="Processed Data/COMPASS_Synoptic_LE_TreeStem_Fluxes_Final_2023_CO2.csv", row.names = FALSE)
# #Change file name
```

## 
