---
title: "COMPASS_Synoptic_LE_Tree_GHG_Flux_Calcs_2023"
author: "Roberta Peixoto"
date: "2026-04-17"
output: 
  html_document:
    self_contained: true 
---

## Add Required Packages

```{r packages, include = FALSE}
lapply(c( "fluxfinder", "readxl",
  "dplyr", "ggplot2", "ggpubr", "stringr",
  "purrr", "tidyverse", "here", "broom", "tibble",
  "googledrive", "googlesheets4", "data.table", 
  "matrixStats", "gridExtra", "grid", "tidyverse", "knitr"), 
  library, character.only = TRUE)
```

## Pull in Raw Datafiles

```{r}
#defining where the raw data is

DATA_DIR_ROOT <- here::here(
  "Tree_Fluxes_Synoptic_LE",
  "2022",
  "Raw Data"
)


#fucntion to read data
now <- function() format(Sys.time(), "%Y-%m-%d %H:%M:%S")
message(now(), " | Script started")
#finding the zip files
message(now(), " | Looking for ZIP files in: ", DATA_DIR_ROOT)

zip_files <- list.files(
  DATA_DIR_ROOT,
  pattern = "\\.zip$",
  full.names = TRUE)

message(now(), " | ZIP files found: ", length(zip_files))
print(zip_files)
  if (length(zip_files) == 0) {
  stop("No ZIP files found. Check DATA_DIR_ROOT.")}

cache <- list()

# Reader function
read_li7810_zip <- function(zip_file, tz = "US/Eastern") {
  message(now(), " | Processing ZIP: ", basename(zip_file))
  if (zip_file %in% names(cache)) {
    message("  -> Using cached version")
    return(cache[[zip_file]])  }

  tmp_dir <- tempfile("li7810_")
  dir.create(tmp_dir)
  on.exit(unlink(tmp_dir, recursive = TRUE), add = TRUE)
  unzip(zip_file, exdir = tmp_dir)
  data_files <- list.files(
    tmp_dir,
    pattern = "\\.data$",
    full.names = TRUE,
    recursive = TRUE  )

  message("  -> .data files found: ", length(data_files))
  if (length(data_files) == 0) {
    stop("No .data files inside ZIP: ", zip_file) }

  dat_list <- lapply(data_files, function(f) {
    message("     Reading: ", basename(f))
    ffi_read_LI7810(f)})

  dat <- data.table::rbindlist(dat_list, fill = TRUE) %>%
    mutate(TIMESTAMP = with_tz(TIMESTAMP, tz))
  cache[[zip_file]] <<- dat
  message("  -> ZIP finished: ", basename(zip_file))
  return(dat)}

# ACTUAL EXECUTION 
message(now(), " | Starting data read")
alldat_list <- lapply(zip_files, read_li7810_zip)
alldat <- data.table::rbindlist(alldat_list, fill = TRUE)
message(now(), " | DONE")
message("Final number of rows: ", nrow(alldat))

# Show result
print(head(alldat))


```

## Read in metadata file and match to raw data

```{r}
md <- "Raw Data/COMPASS_SynopticLE_TreeFlux_Metadata_2022b.csv"
dat <- read.csv(md)

metadat <- dat %>%
  # Remove rows without date
  filter(!is.na(Date)) %>%
  # Keep only desired instruments
  filter(SN %in% c("TG10-01448", "TG10-01426", "TG10-01028")) %>%
  # Normalize Plot labels
  mutate(
    Plot_raw = toupper(Plot),
    # Extract plot components 
    Site = str_extract(Plot_raw, "^[A-Z]{3}"),
    Zone = str_extract(Plot_raw, "(TR|UP)"),
    Replicate = str_extract(Plot_raw, "(?<=TR|UP)[0-9]+"),
    # Letter ONLY if it comes AFTER a number
    Letter = str_extract(Plot_raw, "(?<=[0-9])[A-Z]$"),
    # Time normalization
    Time_start = str_replace_all(as.character(Time_start), "\\.", ":") |> str_trim(),
    Stop.time  = str_replace_all(as.character(Stop.time),  "\\.", ":") |> str_trim(),
    # Safe time parsing (NO WARNINGS) 
    t_start = suppressWarnings(parse_date_time(Time_start, orders = c("H:M:S", "H:M"))),
    t_stop  = suppressWarnings(parse_date_time(Stop.time,  orders = c("H:M:S", "H:M"))),
    # Observation length
    Obs_length = as.numeric(t_stop - t_start, units = "secs"),
    # Timestamp
    TIMESTAMP = as.POSIXct(
      paste(Date, Time_start),
      format = "%m/%d/%Y %H:%M:%S",
      tz = "EST"
    ),
    # Sequential ID
    Seq = row_number(),
    # INAL standardized Plot label
    Plot = if_else(
      is.na(Letter),
      paste(Date, Site, Zone, Replicate, Seq, sep = "_"),
      paste(Date, Site, Zone, Replicate, Letter, Seq, sep = "_")) )

# Quick check
head(metadat$Plot)

```

## Check start times overlapsor missing values

```{r}

#checking if there is any start and stop time overlap from the field notes
overlaps_detected <- metadat %>%
  mutate(
    start_ts = as.POSIXct(
      paste(Date, Time_start),
      format = "%m/%d/%Y %H:%M:%S",
      tz = "EST"
    ),
    end_ts = start_ts + Obs_length
  ) %>%
  arrange(SN, start_ts) %>%
  group_by(SN) %>%
  mutate(
    next_plot     = lead(Plot),
    next_start_ts = lead(start_ts),
    overlap_ffi   = !is.na(next_start_ts) & next_start_ts <= end_ts
  ) %>%
  filter(overlap_ffi) %>%
  select(
    SN,
    Plot,
    start_ts,
    end_ts,
    next_plot,
    next_start_ts,
    Obs_length
  )

print(overlaps_detected)

#checking problems with obs_length - this is a frequent bug for manual field notes 
field_notes_overlap<-metadat %>%
  mutate(
    obs_problem =
      is.na(Obs_length) | Obs_length <= 0
  ) %>%
  filter(obs_problem) %>%
  select(
    SN,
    Date,
    Plot,
    Time_start,
    Stop.time,
    Obs_length
  )

print(field_notes_overlap)

# checking unique measurment start time and stop time
# alldat %>%
#     filter(
#         SN == "TG10-01448",
#         TIMESTAMP > as.POSIXct("2023-01-05 13:28:00"),
#         TIMESTAMP < as.POSIXct("2023-01-05 13:33:00")
#     ) %>%
#     ggplot(aes(x = TIMESTAMP, y = CO2)) +
#     geom_point()#+ylim(300,500)

```

## Match metadata to raw data

```{r}
#function to match data per equipment (considering different serial numbers) uhul!
match_flux_metadata <- function(alldat, metadat) {
  message("Starting metadata ↔ Licor matching")
  #Sanity checks (fail early)
  stopifnot(
    "TIMESTAMP" %in% names(alldat),
    "SN"        %in% names(alldat),
    "Date"      %in% names(metadat),
    "Time_start"%in% names(metadat),
    "Obs_length"%in% names(metadat),
    "Plot"      %in% names(metadat)
  )

  stopifnot(all(metadat$Obs_length > 0, na.rm = TRUE))
  stopifnot(anyDuplicated(metadat$Plot) == 0)
 # Instruments present in BOTH datasets
  instruments <- intersect(
    unique(alldat$SN),
    unique(metadat$SN)
  )
  message("Instruments to process: ", paste(instruments, collapse = ", "))
  results <- lapply(instruments, function(sn) {
    message("Processing SN: ", sn)
    dat_i <- alldat %>%
      filter(SN == sn)
    meta_i <- metadat %>%
      filter(SN == sn) %>%
      arrange(Date, Time_start) %>%
      mutate(metadat_row = row_number())
    stopifnot(nrow(meta_i) > 0)
  # Actual matching 
    dat_i$metadat_row <- ffi_metadata_match(
      data_timestamps = dat_i$TIMESTAMP,
      start_dates     = meta_i$Date,
      start_times     = meta_i$Time_start,
      obs_lengths     = meta_i$Obs_length  )
  # Attach metadata (vectorized, fast)
    dat_i <- dat_i %>%
      mutate(
        Plot      = meta_i$Plot[metadat_row],
        Volume    = meta_i$Volume[metadat_row],
        Temp      = meta_i$Temp[metadat_row],
        Area      = meta_i$Area[metadat_row],
        Dead_band = meta_i$Dead_band[metadat_row]      )
  # Drop unmatched rows
    n_before <- nrow(dat_i)
    dat_i <- dat_i %>% filter(!is.na(metadat_row))
    n_after <- nrow(dat_i)
    message("  Matched rows: ", n_after, " / ", n_before)
    dat_i })
  alldat_matched <- rbindlist(results, fill = TRUE)
  #Final sanity checks 
  stopifnot(
    !anyNA(alldat_matched$Plot),
    nrow(alldat_matched) > 0
  )
  message("Matching finished successfully")
  return(alldat_matched)}

#running the funciton
message("Running final metadata matching...")
alldat_matched <- match_flux_metadata(alldat, metadat)
message("Matched dataset rows: ", nrow(alldat_matched))
head(alldat_matched)

#saving matched data and metadata as raw data is too big (today, Licor files leave in the COMPASS drive)
saveRDS(alldat_matched, file = "alldat_matched_2023.rds")


#can remove any row that have NA's in the CH4 column 
#alldat <- alldat %>%
 # filter(!is.na(CH4))

```

## Temperature from weather station

```{r}
# I need to go back here to get temp from the weather station
```

## fixing some Deadbands

```{r}
# Table of plots with custom deadband adjustments
deadband_overrides <- tibble::tibble(
  Plot = c(
    "7/13/2023CRC_TR315-115",
    "7/27/2023PTR_UP8-125",
    "7/13/2023CRC_TR315-115",
    "3/28/2023OWC_TR127-170",
    "5/10/2023OWC_UP103-73",
    "7/13/2023CRC_TR314-114",
    "8/3/2023CRC_TR314-133",
    "7/13/2023CRC_TR316-116",
    "6/29/2023CRC_TR311-104"

  ),
  DeadBand_override = c(
    8,
    9,
    22,
    22,
    16,
    8,
    16,
    3,
    34
    
  )
)

# Default deadband
alldat_matched <- alldat_matched %>%
  mutate(
    DeadBand_used = as.numeric(Dead_band),
    DeadBand_used = if_else(
      is.na(DeadBand_used),
      0,     
      DeadBand_used    )
  )

# Apply per-plot overrides
alldat_matched <- alldat_matched %>%
  left_join(deadband_overrides, by = "Plot") %>%
  mutate(
    DeadBand_used = if_else(
      !is.na(DeadBand_override),
      DeadBand_override,
      DeadBand_used
    )
  ) %>%
  select(-DeadBand_override)


```

## Calculate fluxes

```{r}

# TO Do - NEED TO PULL TEMP from the weather station
#change the units
alldat_matched $CH4_nmol <- ffi_ppb_to_nmol(alldat_matched $CH4, 
                                volume = alldat_matched $Volume, # m3
                                temp = alldat_matched $Temp)    # degrees C
#> Assuming atm = 101325 Pa
#> Using R = 8.31446261815324 m3 Pa K-1 mol-1

alldat_matched $CH4_nmol_m2 <- alldat_matched $CH4_nmol / 0.16

fluxes <- ffi_compute_fluxes(alldat_matched ,
                             group_column = "Plot", 
                             time_column = "TIMESTAMP", 
                             gas_column = "CH4_nmol_m2", 
                             dead_band = "DeadBand_used")
head(fluxes)

#take a look at the fluxes we calculated with R2 
ggplot(fluxes, aes(Plot, lin_flux.estimate, color = lin_r.squared)) +
  geom_point(size=4) + theme_classic() + 
  geom_linerange(aes(ymin = lin_flux.estimate- lin_flux.std.error,
                   ymax = lin_flux.estimate+ lin_flux.std.error)) +
  scale_color_gradientn(colours = c("darkred",  "red", "white","lightblue", "darkblue"),
                        values = c(0, 0.5, 0.6, 0.7, 1)) +
  ylab("CH4 flux (nmol/m2/s)")

#sort fluxes by R2 
fluxes_good <- as.data.frame(subset(fluxes, lin_r.squared >= .90, select = Plot:TIMESTAMP_max))
fluxes_lowr2 <- as.data.frame(subset(fluxes, lin_r.squared < .90 & lin_flux.estimate < 1, select = Plot:TIMESTAMP_max))
fluxes_weird <- as.data.frame(subset(fluxes, lin_r.squared < .90, select = Plot:TIMESTAMP_max))

#find fluxes that represent the first and fourth quartile of the data 
low <- quantile(fluxes$lin_flux.estimate, 0.05) 
high <- quantile(fluxes$lin_flux.estimate, 0.95) 

fluxes_low <- as.data.frame(subset(fluxes, lin_flux.estimate <= low, select = Plot:TIMESTAMP_max))
fluxes_high <- as.data.frame(subset(fluxes, lin_flux.estimate >= high, select = Plot:TIMESTAMP_max))

tails <- rbind(fluxes_low, fluxes_high)
```

## check the fluxes for which I changed the start time due to lags or high values (recorded in notes of metadata)

```{r}
#### If something is in a V shape should we take the second part to make a positive linear curve?

# Subset data and create relative time (seconds since flux start)
# r_all <- alldat %>%
#   filter(Plot %in% deadband_overrides$Plot) %>%
#   group_by(Plot) %>%
#   mutate(
#     time_sec = as.numeric(
#       difftime(TIMESTAMP, min(TIMESTAMP), units = "secs")
#     )
#   ) %>%
#   ungroup()
# 
# 
# # Random sample of plots for visualization
# plots_sample <- sample(unique(r_all$Plot), 6,replace = FALSE)
# 
# r_plot <- r_all %>%
#   filter(Plot %in% plots_sample)
# 
# # Plot
# ggplot(
#   r_plot,
#   aes(
#     x = TIMESTAMP,
#     y = co2,
#     color = time_sec >= DeadBand_used
#   )
# ) +
#   geom_point(size = 1.8) +
#   facet_wrap(~ Plot_ID, scales = "free", ncol = 2) +
#   scale_color_manual(
#     values = c("grey60", "red"),
#     labels = c("Discarded (deadband)", "Used after deadband")
#   ) +
#   theme_classic() +
#   ylab(expression(paste(CO[2], " (ppm)"))) +
#   xlab("Time")
```

## Spot check a few fluxes with R2 \> 0.9

```{r, echo=FALSE}

### pull a random sample of good fluxes to check and plot 
good_plots <- unique(fluxes_good$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
good_sample <- sample(good_plots, 9, replace=FALSE)

#Loop to run five random plots subset above 
for(P in good_sample){ 
  r1 <- subset(alldat_matched, Plot == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(as.POSIXct(TIMESTAMP), CH4)) + geom_point() + 
  scale_x_datetime(breaks = "15 sec") + 
  theme_classic() +
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

####################

```

## Take a look at the fluxes \<0.9 and a flux estimate below 1

```{r, echo = FALSE}

####Do we just leave these alone because the fluxes are so low?####

### pull out fluxes with R2 < 0.9 and put them in a list 
low_plots <- unique(fluxes_lowr2$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
low_sample <- sample(low_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in low_sample){ 
  r1 <- subset(alldat_matched , Plot == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() +
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

####################
```

## Take a look at the bad fluxes \<0.9 and have a high flux estimate

## See if we want to remove these fluxes or if we want to try and recalculate these ones?

```{r}
### pull out fluxes with R2 < 0.9 and put them in a list 
bad_plots <- unique(fluxes_weird$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
bad_sample <- sample(bad_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in bad_sample){ 
  r1 <- subset(alldat_matched , Plot == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth()
  print(p1)
  }

####################
```

## Look at the really high or really low ones to see if there is a timing issue or something

```{r}
### pull out fluxes in the 5% and 95% tails to look at really high and really low 
tail_plots <- unique(tails$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
tail_sample <- sample(tail_plots, 2, replace=FALSE)

#Loop to run five random plots subset above 
for(P in tail_sample){ 
  r1 <- subset(alldat_matched , Plot == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CH4 [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

#would be nice to put on the plot the flux estimate and if it is a high or low or something 
####################
```

## Check all fluxes

```{r}
##Put all fluxes together
fluxes_all <- as.data.frame(subset(fluxes, select = Plot:TIMESTAMP_max))

all_plots <- unique(fluxes_all$Plot)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
all_sample <- sample(all_plots, 5, replace=FALSE)

#Loop to run five random plots subset above
for(P in all_sample){
  r1 <- subset(alldat_matched, Plot== P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, CH4)) + geom_point() + theme_classic() +
  scale_x_datetime(breaks = "15 sec") +
  ylab(expression(paste( CH [4], " (ppb)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth(color="red")
  print(p1)
  }

####################
```

## Remove any fluxes that do not fit criteria

```{r}

# Create a new dataframe by removing rows that were present in fluxes_weird (11 obs)
#allflux_filtered <- allflux[!(allflux$Plot %in% fluxes_weird$Plo), ] # not all of them are bad, so I am removing below the ones we visually confirmed that were bad

#Any others that are not in the weird category that need to be removed: 
  #also removing ones that are crazy off but were not in the fluxes_weird dataframe 

allflux_filtered <- fluxes_all #%>%
#   filter(!Plot %in% c("2022_09_23_PTR_TR_9_1", #leak 
#                          "2022_07_07_PTR_TR_9_1", #leak
#                          "2022_07_08_CRC_TR_36_1" #plateau + sharp increase - soil disturbed during measurement?
#                       ))
# 
# 
# # View the result
# head(allflux_filtered)
```

## Do some visualization of QAQC plots to assess fluxes

```{r}

p1 <- ggplot(fluxes, aes(lin_flux.estimate, rob_flux.estimate, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  labs(title = "A") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Robust Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot")) +
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p1

p2 <- ggplot(fluxes, aes(lin_r.squared, poly_r.squared, color = Plot)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  xlim(0,1) + ylim(0,1) +
  labs(title = "B") + 
  ylab(expression(paste("R"^2* " Second-Order Polynomial"))) +
  xlab(expression(paste("R"^2* " Linear Model"))) +
  guides(color = guide_legend(title = "Plot_ID")) + 
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p2

p3 <- ggplot(fluxes, aes(lin_flux.estimate, HM81_flux.estimate, color = lin_r.squared )) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  #xlim(0,50) + #ylim(0,50) +
  labs(title = "C") + 
  ylab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") HM1981 Method"))) +
  xlab(expression(paste( CH [4], " Flux (nmol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot_ID")) + 
  theme(legend.position="right") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)

p3

#ggarrange(p1, p2, p3, ncol=3, nrow=1, common.legend = TRUE, legend="none")
```

## Read out final dataframe and export for use elsewhere

```{r}

allflux_filtered_renamed <- fluxes %>%
  mutate(
    Date = mdy(str_extract(Plot, "^[0-9]+/[0-9]+/[0-9]{4}")),
    Month = month.abb[month(Date)],
    Site = str_extract(Plot, "(PTR|OWC|CRC)"),
    Zone = str_extract(Plot, "_TR_|_UP_") %>%
      str_remove_all("_"),
    Replicate = as.integer(
      str_extract(Plot, "(?<=_(TR|UP)_)[0-9]+")
    ),
   Plot_ID = as.integer(
      str_extract(Plot, "[0-9]+$")) ) %>%
  
  # Optional: rebuild a clean Plot label
  mutate(
    Plot_clean = paste(
      format(Date, "%Y-%m-%d"),
      Site,
      Zone,
      Replicate,
      Plot_ID,
      sep = "_"
    )
  ) %>%
  
  select(
    Plot_clean,
    Date,
    Month,
    Site,
    Zone,
    Replicate,
    Plot_ID,
    everything(),
    -Plot
  )



#export as a csv to the folder that you want
write.csv(allflux_filtered_renamed, file="Processed Data/COMPASS_Synoptic_LE_TreeStem_Fluxes_Final_2023_CH4.csv", row.names = FALSE)
# #Change file name
```

## 
