---
title: "COMPASS_Synoptic_LE_SoilFluxes_2023_CO2"
author: "Roberta Peixoto "
date: "2026-04-02"
output: 
  html_document:
    self_contained: true
---
## Goals of this code: 
1.	Reads in raw smart chamber
2.	If no metadatafile, write one out & fields will be: 
Label 
the IDs pulled from the label, 
Plot *fixed so that if there are any issues in the label they are good here 
Drop column – false as long as label looks right 
adjust time – would be initially populated by the start time in the OR blank 
adjust observation length – initially populated by the obs length  OR blank 
notes / comment 
3.	Code reads in the meta data and applies it to the data 
a.	i.e. drops fluxes that have the drop column TRUE or FLAG this 
b.	adjusting times as needed 
4.	Now you compute fluxes 
5.	Visualize and look for issues 
6.	If found, report to meta data file then go back to 3 and reanalyze 
7.	Apply algorithmic qaqc and generate flags 


## Set Up & Load Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("remotes") # if necessary
#remotes::install_github("COMPASS-DOE/fluxfinder")

library(fluxfinder)
library(ggplot2)
library(ggpubr)
library(lubridate)
library(dplyr)
library(tidyr)
library(tidyverse)
library(stringr)

```

## Pull in Smart Chamber Files from the Raw Data folder
```{r, message=FALSE, warning = FALSE, results ='hide'}

# Create a temporary directory
tmp_dir <- tempdir()

# Get list of files inside the zip
zip_contents <- unzip("Raw Data_LE_2023.zip", list = TRUE)

# Keep only JSON files
json_files <- zip_contents$Name[grepl("json$", zip_contents$Name, ignore.case = TRUE)]

# Extract only the JSON files
unzip(
  "Raw Data_LE_2023.zip",
  files = json_files,
  exdir = tmp_dir
)

# List extracted JSON files (full paths)
file.list <- list.files(
  path = tmp_dir,
  pattern = "json$",
  full.names = TRUE,
  recursive = TRUE
)

head(file.list)

# Read in the data
alldat1 <- lapply(file.list, ffi_read_LIsmartchamber)


# Combine into one data.table
alldat <- data.table::rbindlist(alldat1, fill = TRUE)


## Noticed that some site names were recorded in the file name and not in the flux label for 2023 
## dealing with it in the next chunk 



```

## Unfortunatly - we had many problems with labels - we need to include the sites in the Plot_ID
```{r}
# Table describing the days we have visited each site
site_lookup <- tibble::tribble(
  ~Date,         ~Site,
  "2023-01-20",  "CRC",
  "2023-02-21",  "CRC",
  "2023-03-24",  "CRC",
  "2023-04-14",  "CRC",
  "2023-05-30",  "CRC",
  "2023-06-29",  "CRC",
  "2023-07-13",  "CRC",
  "2023-08-03",  "CRC",
  "2023-09-01",  "CRC",
  "2023-10-06",  "CRC",
  "2023-01-05",  "OWC",
  "2023-02-10",  "OWC",
  "2023-03-28",  "OWC",
  "2023-04-21",  "OWC",
  "2023-05-10",  "OWC",
  "2023-06-22",  "OWC",
  "2023-07-25",  "OWC",
  "2023-08-10",  "OWC",
  "2023-09-08",  "OWC",
  "2023-10-20",  "OWC",
  "2023-01-12",  "PTR",
  "2023-02-24",  "PTR",
  "2023-07-26",  "PTR",
  "2023-08-17",  "PTR",
  "2023-10-11",  "PTR"
) %>%
  mutate(Date = as.Date(Date))

# alldat  - date formatting
alldat <- alldat %>%
  mutate(Date = as.Date(TIMESTAMP))%>%
  left_join(site_lookup, by = "Date")


# Fixing label code problems 
alldat <- alldat %>%
  
  ## Remove known invalid labels
  filter(!label %in% c("Test2", "Up30", "Up30c", "Up27")) %>%

  ## Normalize the raw label (do NOT trust embedded site)
  mutate(
    label = label %>%
      str_remove_all("[sSeE]") %>%          # remove stray s/e
      str_replace_all("-", "_") %>%         # standardize separators
      str_replace_all(c(
        "[Ww][tT]"     = "W",
        "[Tt][Rr][a]?" = "TR",
        "[Uu][Pp]"     = "UP",
        "[Ww][lL]"     = "W",
        "[w]"          = "W"
      )) %>%
      ## ensure ZONE_NUMBER format (W46 → W_46, TR34 → TR_34)
      str_replace("_(W|TR|UP)(\\d)", "_\\1_\\2") %>%
      ## separate suffix from number (48a → 48_a)
      str_replace("(\\d+)([A-Za-z])$", "\\1_\\2")
  ) %>%

  ## Remove ANY site embedded in the original label
  mutate(
    label_core = str_remove(label, "^(CRC|OWC|PTR)_?")
  ) %>%

  ## Rebuild label using ONLY the site from site_lookup  #OWC recorded wrongly on 2023-Jan-12 instead of PTR
  mutate(
    label = paste0(Site, "_", label_core)
  ) %>%

  ## Standardize suffixes (only when they exist)
  mutate(
    n_underscores = str_count(label, "_"),
    suffix = str_extract(label, "(?<=_)[^_]+$"),
    label = if_else(
      n_underscores >= 3 & !is.na(suffix),
      str_replace(
        label,
        paste0("_", suffix, "$"),
        paste0("_", tolower(suffix))
      ),
      label
    )
  ) %>%

  ## Final cleanup
  select(-label_core, -n_underscores, -suffix)



#I have to change labels that were recorded wrong - tracked in the field notes

# CRC 02-21-2023 TR 36 -> recorded as W but it is a TR sample
# CRC 02-21-2023 TR 37 -> recorded as W but it is a TR sample
# CRC 02-21-2023 TR 38 -> recorded as UP but it is a TR sample
# CRC 02-21-2023 TR 40 -> recorded as UP but it is a TR sample
# CRC 03-24-2023 W 48 -> recorded as TR but it is a W sample

# CRC 10-06-2023 TR 37 -> recorded as UP but it is a TR sample
# CRC 10-06-2023 TR 39 -> recorded as UP but it is a TR sample
# CRC 10-06-2023 TR 40 -> recorded as UP but it is a TR sample

# OWC 04-21-2023 TR 58 -> recorded as UP but it is a TR sample
# OWC 04-21-2023 TR 60 -> recorded as UP but it is a TR sample
# OWC 04-21-2023 TR 61 -> recorded as UP but it is a TR sample
# OWC 04-21-2023 TR 63 -> recorded as UP but it is a TR sample
# OWC 04-21-2023 TR 64 -> recorded as UP but it is a TR sample

#extracting the collar number from the label
alldat <- alldat %>%
  mutate(
    PlotNum = as.integer(stringr::str_extract(label, "(?<=_)\\d+(?=(_|$))"))
  )

#correcting the zone that was saved wrongly

alldat <- alldat %>%
  mutate( label = case_when(
      # CRC 02-21-2023 — TR was saved wrongly as W
      Site == "CRC" & Date == as.Date("2023-02-21") & PlotNum %in% c(36, 37) ~
        stringr::str_replace(label, "_W_", "_TR_"),

      Site == "CRC" & Date == as.Date("2023-02-21") & PlotNum %in% c(38, 40) ~
        stringr::str_replace(label, "_UP_", "_TR_"),

      # CRC 03-24-2023 — W was saved wrongly as TR
      Site == "CRC" & Date == as.Date("2023-03-24") & PlotNum == 48 ~
        stringr::str_replace(label, "_TR_", "_W_"),

      # CRC 10-06-2023 — TR was saved wrongly as UP
      Site == "CRC" & Date == as.Date("2023-10-06") & PlotNum %in% c(37, 39, 40) ~
        stringr::str_replace(label, "_UP_", "_TR_"),
      
      # OWC 04-21-2023 — UP was saved wrongly as TR
      Site == "OWC" & Date == as.Date("2023-04-21") & PlotNum %in% c(58, 60, 61, 63, 64) ~
        stringr::str_replace(label, "_UP_", "_TR_"),

      TRUE ~ label
    )
  )


```

## Create a unique Plot_ID, check if there is a meta-data file, if not, create a meta-data file
```{r}

############### only run this first part if you don't already have a meta-data file 
        #because if you do then all the data for each one will have a plot and you won't be able to change start times

#create unique Plot ID for each measurement: 
#detach("package:data.table", unload=TRUE)

# I am changing this from a month_ID to a date_ID as I have more than one day per month
alldat <- alldat %>%
  mutate(
    Date_str = format(as.Date(TIMESTAMP), "%Y_%m_%d"),
    Plot_ID = str_c(Date_str, label, RepNum, sep = "_")  )

#this needs to be PLot ID so that later we ID plot from metadata and not the metadata row 

# Path inside the zip
zip_file <- "Raw Data_LE_2023.zip"
csv_inside_zip <- "Raw Data/COMPASS_SmartChamber_Metadata_2023.csv"

# Make sure the destination folder exists
dir.create("Raw Data", showWarnings = FALSE)

# Extract only the metadata CSV if it doesn't already exist
if (!file.exists(csv_inside_zip)) {
  unzip(
    zip_file,
    files = csv_inside_zip,
    exdir = "."
  )
}

# Define the file path where you want to save the CSV file
file_path <- "Raw Data/COMPASS_SmartChamber_Metadata_2023.csv"

# Check if the CSV file exists
if (file.exists(file_path)) {
  # If it exists, read it back into R
  metadata <- read.csv(file_path)
  print("CSV file exists and has been read into the code.")
} else {
  # If it does not exist, create the CSV file
    #create a metadata file from the Plot IDs 
      # pull timestart timestamps to put in the metadata and obs lengths 
  metadata <- alldat %>%
  group_by(InstrumentSerialNumber, Plot_ID) %>%
  summarise(SN = first(InstrumentSerialNumber), #Roberta => I included this to process more than one licor per time
    Date = format(min(TIMESTAMP), "%Y-%m-%d"),
    Time_start = format(min(TIMESTAMP), "%H:%M:%S"),
    First_Timestamp = min(TIMESTAMP),  # Minimum (earliest) timestamp for each Plot_ID
    Last_Timestamp = max(TIMESTAMP),   # Maximum (latest) timestamp for each Plot_ID
    Total_Time = as.numeric(difftime(max(TIMESTAMP), min(TIMESTAMP), units = "secs")),  # Time difference in seconds
    ) %>%
  ungroup() %>%
  mutate(Total_Time = pmin(Total_Time, 60),
          drop = FALSE,# Add a 'drop' column with 'FALSE' for every row 
          Notes = "") 
  #mutate(notes = " ")  # Add a 'drop' column with 'FALSE' for every row
  #mutate(Deadband = 5)
  #mutate(Plot = Plot_ID) #think need this because we need to make a plot column
  
  # Write df_unique to CSV
  write.csv(metadata, file = file_path, row.names = FALSE)
  
  print("CSV file does not exist. It has been created.")
}

```

## Mark fluxes that were retaken - this is the first issue! Need to get all but last flux to be TRUE for drop - fieldnotes with explanation for flux retaken - usually weird zeros, Steven probe not inserted in the soil, or label mistake  
```{r}

# fluxes that were retaken have similar names with a letter after it 
metadata_filtered <- metadata %>%
  mutate(
    # Base plot ID (everything before the last _number or _number_letter)
    base_id = str_replace(Plot_ID, "_[0-9]+(_[a-z])?$", ""),
    
    # Replicate number (e.g., 32 from 1_Jul_CRC_UP_32 or 1_Jul_CRC_UP_32_b)
    replicate = as.numeric(
      str_extract(Plot_ID, "(?<=_)[0-9]+(?=(_[a-z])?$)")
    ),
    
    # Suffix letter (if present); otherwise empty string
    letter = str_extract(base_id, "(?<=_)[A-Za-z]$"),
    letter = if_else(is.na(letter), "", letter),
    
    # Letter order to define order: "" < a < b < c < d ...
    letter_order = if_else(
      letter == "",
      0L,
      match(letter, letters)  ) ) 


#creating a core ID to compare the retaken obs
metadata_filtered <- metadata_filtered %>%
  mutate(
    # ID for plot not considering retake (remove _a, _b, _c, etc.)
    core_id = str_remove(base_id, "_[A-Za-z]$")
  ) %>%
  group_by(core_id) %>%
  mutate(
    drop = letter_order < max(letter_order, na.rm = TRUE)
  ) %>%
  ungroup()
 

```

##Dont need to pull L1 temp data or any other temps like we do for macrophyte because the smartchamber records the temp

## Match up the metadata to the rest of the data   
```{r}
#remove any metadata rows that have NAs for the PlotID
metadata_filtered  <- metadata_filtered %>%
  filter(!is.na(Plot_ID)) %>%
  filter(str_detect(Plot_ID, paste(c("OWC", "CRC", "PTR"), collapse = "|"))) 

alldat<-alldat%>% filter(!is.na(label))

#match the altered metadata to the concentration data with fluxfinder ffi_metadata_match 
# this next part was adapted to include SN and process all Licors at the same time, handling overlapping timestamp among >1 licor instruments. 

# matching SN
alldat <- alldat %>%
  rename(SN=InstrumentSerialNumber)%>%
  group_by(SN) %>%
  mutate(
    metadat_row = ffi_metadata_match(
      data_timestamps = TIMESTAMP,
      start_dates     = metadata_filtered$Date[metadata_filtered$SN == first(SN)],
      start_times     = metadata_filtered$Time_start[metadata_filtered$SN == first(SN)],
      obs_lengths     = metadata_filtered$Total_Time[metadata_filtered$SN == first(SN)]
    )
  ) %>%
  ungroup()

alldat <- alldat %>%
  left_join(
    metadata_filtered %>% mutate(metadat_row = row_number()),
    by = c("SN", "metadat_row", "Plot_ID")
  )

#problems with labels - there was a chamber Offset of 75 instead of 5 and chamber volume of 0 from the smart chamber

alldat <- alldat %>%
  mutate(
    # Fix the Offset bug Plot_ID OWC_02-10-2023 W 67: 75 -> 5 and recalculating volume based on different Offsets
    Offset_cm = if_else(Offset == 75, 5, Offset),

    # Fix chamber bug: volume should never be zero
    ChamVolume_fixed_cm3 = if_else(
      ChamVolume == 0,
      4244.1,
      ChamVolume
    ),

    # Recalculate collar volume (cm3)
    CollarVolume_cm3 = Area * Offset_cm,

    # Recalculate total volume (cm3)
    TotalVolume_fixed_cm3 =
      ChamVolume_fixed_cm3 +
      IrgaVolume +
      CollarVolume_cm3,
  )


```

## fixing some Deadbands  
```{r}
# Table of plots with custom deadband adjustments
deadband_overrides <- tibble::tibble(
  Plot_ID = c(
    "2023_06_29_CRC_UP_29_1",
   "2023_05_10_OWC_TR_61_1",
   "2023_06_22_OWC_TR_60_2",
   "2023_06_22_OWC_TR_63_2",
   "2023_06_29_CRC_UP_26_2",
   "2023_06_22_OWC_TR_65_2",
   "2023_07_13_CRC_TR_39_2",
   "2023_01_05_OWC_TR_60_1",
   "2023_07_13_CRC_TR_33_2",
   "2023_06_22_OWC_TR_68_2",
   "2023_06_22_OWC_W_69_2",
   "2023_07_13_CRC_TR_37_2",
   "2023_06_22_OWC_UP_55_2",
   "2023_06_22_OWC_TR_64_2",
   "2023_04_21_OWC_UP_58_2",
   "2023_07_13_CRC_TR_34_3_2",
   "2023_06_22_OWC_UP_52_1",
   "2023_06_22_OWC_TR_60_2",
   "2023_05_10_OWC_TR_61_1",
   "2023_06_22_OWC_TR_58_1",
   "2023_06_22_OWC_W_67_2",
   "2023_06_29_CRC_TR_36_2",
   "2023_05_10_OWC_TR_63_2",
   "2023_01_05_OWC_TR_63_1"
  ),
  DeadBand_override = c(
    5,
    7,
    6,
    10,
    8,
    12,
    9,
    11,
    13,
    7,
    10,
    11,
    9,
    9,
    10,
    9,
    8,
    11,
    14,
    8,
    11,
    11,
    16,
    12
    
  )
)

# Default deadband
alldat <- alldat %>%
  mutate(DeadBand_used = DeadBand) 

# Apply per-plot overrides
alldat <- alldat %>%
  left_join(deadband_overrides, by = "Plot_ID") %>%
  mutate(
    DeadBand_used = if_else(
      !is.na(DeadBand_override),
      DeadBand_override,
      DeadBand_used
    )
  ) %>%
  select(-DeadBand_override)

```


## Change the units from ppb to nmol  
```{r}
#change the units
alldat$CO2_umol <- ffi_ppb_to_nmol(alldat$co2, 
                                volume = (alldat$TotalVolume_fixed_cm3 / 1000000), # have to divide to get to m3
                                temp = alldat$chamber_t)    # degrees C
#> Assuming atm = 101325 Pa
#> Using R = 8.31446261815324 m3 Pa K-1 mol-1

alldat$CO2_umol_m2 <- alldat$CO2_umol / 0.0318 #adjust for the area 


```

## Calculate fluxes
```{r, message=FALSE, warning = FALSE, results ='hide'}

#need to create a new line that would say only calculate the flux if the drop column is false
alldat_dropped<-alldat%>%filter(drop!=TRUE)

#selecting the base id to include in the flux df later
plot_base_lookup <- alldat_dropped%>%
  select(Plot_ID, base_id)%>% 
  filter(!is.na(base_id)) %>%   
  distinct(Plot_ID, .keep_all = TRUE)


#calculate fluxes for all other sites 
fluxes <- ffi_compute_fluxes(alldat_dropped,
                             group_column = "Plot_ID", 
                             time_column = "TIMESTAMP", 
                             gas_column = "CO2_umol_m2", 
                             dead_band = "DeadBand_used"
                             )
head(fluxes)
#including the unique ID 
fluxes <- fluxes %>%
  left_join(plot_base_lookup, by = "Plot_ID")

##Create a dataframe with averaged soil probe data per Plot and add to flux data
soildat <- alldat %>% 
    group_by(Plot_ID) %>%
  dplyr::summarize(
     soilp_c = mean(soilp_c, na.rm=TRUE),
     soilp_m = mean(soilp_m, na.rm=TRUE),
     soilp_t = mean(soilp_t, na.rm=TRUE))

alldf <- left_join(fluxes, soildat, by = "Plot_ID")

fluxes_neg <- as.data.frame(subset(alldf, lin_flux.estimate < 0, select = Plot_ID:TIMESTAMP_max))

(print(fluxes_neg))

allflux <-alldf%>%filter(lin_flux.estimate > 0)

# saving a negative CO2 flux table to remove int he CH4 flux script
drop_neg_CO2 <- alldf %>%
  filter(lin_flux.estimate < 0) %>%
  select(Plot_ID, TIMESTAMP_min, TIMESTAMP_max) %>%
  mutate(drop_CO2 = TRUE)

saveRDS(drop_neg_CO2, "drop_neg_CO2_2023.rds")

```


## Visualize fluxes and sort by r2 
```{r}

#take a look at the fluxes we calculated with R2 
ggplot(allflux, aes(Plot_ID, lin_flux.estimate, color = lin_r.squared)) +
  geom_point(size=4) + theme_classic() + #ylim(-10,50) + 
  geom_linerange(aes(ymin = lin_flux.estimate- lin_flux.std.error,
                   ymax = lin_flux.estimate+ lin_flux.std.error)) +
  scale_color_gradientn(colours = c("darkred",  "red", "white","lightblue", "darkblue"),
                        values = c(0, 0.5, 0.6, 0.7, 1)) +
  ylab("CO2 flux (umol/m2/s)")+
  scale_x_discrete(breaks = function(x) x[seq(1, length(x), by = 30)]) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)  )


#sort fluxes by R2 
fluxes_good <- as.data.frame(subset(allflux, lin_r.squared >= .90, select = Plot_ID:TIMESTAMP_max))
fluxes_lowr2 <- as.data.frame(subset(allflux, lin_r.squared < .90 & lin_flux.estimate < 1, select = Plot_ID:TIMESTAMP_max))
fluxes_weird <- as.data.frame(subset(allflux, lin_r.squared < .90 & lin_flux.estimate > 1, select = Plot_ID:TIMESTAMP_max))

#find fluxes that represent the first and fourth quartile of the data 
low <- quantile(allflux$lin_flux.estimate, 0.05) 
high <- quantile(allflux$lin_flux.estimate, 0.95) 

fluxes_low <- as.data.frame(subset(allflux, lin_flux.estimate <= low, select = Plot_ID:TIMESTAMP_max))
fluxes_high <- as.data.frame(subset(allflux, lin_flux.estimate >= high, select = Plot_ID:TIMESTAMP_max))

tails <- rbind(fluxes_low, fluxes_high)

```

## check the fluxes for which I changed the start time due to lags or ebullition (recorded in notes of metadata)
```{r}

#### If something is in a V shape should we take the second part to make a positive linear curve?

# Subset data and create relative time (seconds since flux start)
r_all <- alldat %>%
  filter(Plot_ID %in% deadband_overrides$Plot_ID) %>%
  group_by(Plot_ID) %>%
  mutate(
    time_sec = as.numeric(
      difftime(TIMESTAMP, min(TIMESTAMP), units = "secs")
    )
  ) %>%
  ungroup()


# Random sample of plots for visualization
plots_sample <- sample(unique(r_all$Plot_ID), 2,replace = FALSE)

r_plot <- r_all %>%
  filter(Plot_ID %in% plots_sample)

# Plot
ggplot(
  r_plot,
  aes(
    x = TIMESTAMP,
    y = co2,
    color = time_sec >= DeadBand_used
  )
) +
  geom_point(size = 1.8) +
  facet_wrap(~ Plot_ID, scales = "free", ncol = 2) +
  scale_color_manual(
    values = c("grey60", "red"),
    labels = c("Discarded (deadband)", "Used after deadband")
  ) +
  theme_classic() +
  ylab(expression(paste(CO[2], " (ppm)"))) +
  xlab("Time")

```


## Spot check a few fluxes with R2 > 0.9  
```{r, echo=FALSE}

### pull a random sample of good fluxes to check and plot 
good_plots <- unique(fluxes_good$Plot_ID)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
good_sample <- sample(good_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in good_sample){ 
  r1 <- subset(alldat, Plot_ID == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(as.POSIXct(TIMESTAMP), co2)) + geom_point() + 
  scale_x_datetime(breaks = "15 sec") + 
  theme_classic() +
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

####################
```

## Take a look at the  fluxes <0.9 and a flux estimate below 1
```{r, echo=FALSE}

####Do we just leave these alone because the fluxes are so low?####

### pull out fluxes with R2 < 0.9 and put them in a list 
low_plots <- unique(fluxes_lowr2$Plot_ID)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
low_sample <- sample(low_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in low_sample){ 
  r1 <- subset(alldat, Plot_ID == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, co2)) + geom_point() + theme_classic() +
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

####################

```

## Take a look at the bad fluxes <0.9 and have a high flux estimate 
# See if we want to remove these fluxes or if we want to try and recalculate these ones? 
```{r, echo=FALSE}

### pull out fluxes with R2 < 0.9 and put them in a list 
bad_plots <- unique(fluxes_weird$Plot_ID)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
bad_sample <- sample(bad_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in bad_sample){ 
  r1 <- subset(alldat, Plot_ID == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, co2)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth()
  print(p1)
  }

####################


```

## Look at the really high or really low ones to see if there is a timing issue or something
```{r, echo=FALSE}

### pull out fluxes in the 5% and 95% tails to look at really high and really low 
tail_plots <- unique(tails$Plot_ID)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
tail_sample <- sample(tail_plots, 5, replace=FALSE)

#Loop to run five random plots subset above 
for(P in tail_sample){ 
  r1 <- subset(alldat, Plot_ID == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, co2)) + geom_point() + theme_classic() + 
  scale_x_datetime(breaks = "15 sec") + 
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) + 
  geom_smooth()
  print(p1)
  }

#would be nice to put on the plot the flux estimate and if it is a high or low or something 
####################


```

## Check all fluxes
```{r, echo=FALSE}

##Put all fluxes together
fluxes_all <- as.data.frame(subset(allflux, select = Plot_ID:TIMESTAMP_max))

all_plots <- unique(fluxes_all$Plot_ID)

# take a random sample (FALSE so we don't get the same plot twice)
# set.seed(...) to guarantee reproducibility
all_sample <- sample(all_plots, 5, replace=FALSE)

#Loop to run five random plots subset above
for(P in all_sample){
  r1 <- subset(alldat, Plot_ID == P & !is.na(metadat_row))
  p1 <- ggplot(r1, aes(TIMESTAMP, co2)) + geom_point() + theme_classic() +
  scale_x_datetime(breaks = "15 sec") +
  ylab(expression(paste( CO [2], " (ppm)"))) +
  xlab("Time") + labs(title = P) +
  geom_smooth(color="red")
  print(p1)
  }

####################
```


## Remove any fluxes that do not fit criteria 
```{r, echo=FALSE}

# Create a new dataframe by removing rows that were present in fluxes_weird (8 obs)
#allflux_filtered <- allflux[!(allflux$Plot_ID %in% fluxes_weird$Plot_ID), ] # not all of them are bad, so I am removing below the ones we visually confirmed that were bad

#Any others that are not in the weird category that need to be removed: 
  #also removing ones that are crazy off but were not in the fluxes_weird dataframe 

# allflux_filtered <- allflux %>%
#   filter(!Plot_ID %in% c("", #leak 
#                          ""  #plateau + sharp increase - soil disturbed during measurement?
#                       ))


# View the result
# head(allflux_filtered)


```


## Do some visualization of QAQC plots to assess fluxes  
```{r, echo=FALSE}

p1 <- ggplot(allflux, aes(lin_flux.estimate, rob_flux.estimate, color = Plot_ID)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  labs(title = "A") + 
  ylab(expression(paste( CO [2], " Flux (umol m"^-2* " s"^-1*") Robust Method"))) +
  xlab(expression(paste( CO [2], " Flux (umol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot_ID")) +
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p1

p2 <- ggplot(allflux, aes(lin_r.squared, poly_r.squared, color = Plot_ID)) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  xlim(0,1) + ylim(0,1) +
  labs(title = "B") + 
  ylab(expression(paste("R"^2* " Second-Order Polynomial"))) +
  xlab(expression(paste("R"^2* " Linear Model"))) +
  guides(color = guide_legend(title = "Plot_ID")) + 
  theme(legend.position="none") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)
p2

p3 <- ggplot(allflux, aes(lin_flux.estimate, HM81_flux.estimate, color = lin_r.squared )) +
  geom_point(size=4) + geom_abline() + 
  theme_classic() + 
  #xlim(0,50) + #ylim(0,50) +
  labs(title = "C") + 
  ylab(expression(paste( CO [2], " Flux (umol m"^-2* " s"^-1*") HM1981 Method"))) +
  xlab(expression(paste( CO [2], " Flux (umol m"^-2* " s"^-1*") Linear Method"))) +
  guides(color = guide_legend(title = "Plot_ID")) + 
  theme(legend.position="right") +
  theme(axis.title.x = element_text(size=12), axis.text = element_text(size=12),
        axis.title.y = element_text(size=12), legend.text=element_text(size=12),
        panel.border = element_rect(colour = "black", fill=NA, linewidth =1), 
        aspect.ratio = 1)

p3

#ggarrange(p1, p2, p3, ncol=3, nrow=1, common.legend = TRUE, legend="none")
```

## Read out final dataframe and export for use elsewhere 
```{r, echo=FALSE}

allflux_filtered_renamed <- allflux %>%
    separate(
    Plot_ID,
    into = c("Year", "Month_num", "Day", "Site", "Zone", "Rest"),
    sep = "_",
    fill = "right",
    extra = "merge",
    remove = FALSE
  ) %>%
  
  mutate(
    Month = month.abb[as.integer(Month_num)],
        # RepNumber is always the final number
    RepNumber = as.integer(str_extract(Plot_ID, "[0-9]+$")),
        # Replicate = numeric collar ID only (removes letters like a, b)
    Replicate = as.integer(str_extract(Rest, "^[0-9]+"))
  ) %>%
      
  rename(Plot = Plot_ID) %>%
  
  select(
    RepNumber, Month, Site, Zone, Replicate, Plot,
    everything(),
    -Year, -Day, -Month_num, -Rest, -base_id
  )


#export as a csv to the folder that you want 
write.csv(allflux_filtered_renamed, file="Processed Data/COMPASS_Synoptic_LE_SmartChamber_Fluxes_Final_2023_CO2.csv", row.names = FALSE) 
#Change file name


```



